---
title: "Significance grid - Full dataset, per-pot variables - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-08-17"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# READ ME FIRST

This is the **per-pot** counterpart of `../PerPlant/Significance_Grid_DRAFT.Rmd`: same grid
layout (rows = ANOVA terms, columns = response variables, one grid for raw-gram "mass"
variables and one for "percent" variables), same **full dataset** scope (Inter + Mono
pots together, Polyculture as a 3-way factor), same everything - except the mass-family
variables are fit on the **pot-total** biomass columns (`W_NPP`, `W_ANPP_g`, `W_BNPP_g`,
`W_Berries_g`, `Pot_NPP_g`) instead of the density-adjusted, "per plant" (`_pp`) columns
(`W_NPP_pp`, `W_ANPP_g_pp`, `W_BNPP_g_pp`, `W_Berries_g_pp`, `Pot_NPP_g_pp`) that the
original doc uses. Percent/allocation and AMF-colonization variables have no `_pp`
counterpart in the first place (see `Data/Data_key_column_definitions.csv`), so those
columns and results are unchanged from the original.

As with the original:

- **cell label** (`Yes` / `Kinda (<.1)` / `No` / blank-`NA`) based on the **BH-adjusted**
  p-value (`p_adj`) from `term_summary`, not a raw per-test p-value.
- **direction text** (e.g. `Inter > Mono`, or per-slice text for interactions like
  `Inter: Inoc > Strl<br>Mono: Strl > Inoc`) computed automatically from group means.
- **a small thumbnail plot** embedded directly in every cell (`No` included), generated
  from the same data used to fit the model, with the p-value and significance stars
  (`***`/`**`/`*`/`.`/`ns`) printed as a subtitle directly on the plot - bolded and
  darkened when p_adj < .05.

A blank/`NA` cell here means the term literally isn't in that variable's model, same as
the original - this happens automatically from the model structure in `battery_core.R`,
nothing is being special-cased here to produce it.

This doc sources this folder's own `battery_core.R` and `grid_helpers.R` (not the ones in
`CleanCode/`) - see the header comments in each for exactly what differs (only
`battery_core.R`'s mass-family response columns; `grid_helpers.R` is an unchanged copy).

```{r, message=FALSE}
source("battery_core.R")   # combined/inoc_only/inter_only/mono_only, battery_spec,
                             # spec_by_label, results_list, term_summary (scope = "full",
                             # per-pot mass variables)
source("grid_helpers.R")    # plot_cell(), direction_label(), build_cell_data(), render_grid()
```

```{r}

plot_dir <- "plots/significance_grid_plots"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. Build the cell data (one row per response variable x ANOVA term)

```{r}
cell_data <- build_cell_data(term_summary, spec_by_label, results_list, plot_dir)
```

# 2. Render the two grids (mass / percent), matching the old table's layout

```{r}
term_labels <- term_labels_for_scope("full")
```

## 2.1 Mass variables (grams, per pot)

```{r}
render_grid("mass", cell_data, battery_spec, term_labels)
```

## 2.2 Percent / allocation variables

```{r}
render_grid("percent", cell_data, battery_spec, term_labels)
```

# 3. Focused question: at Control irrigation, how does inoculation affect berry allocation?

Pulls one specific slice out of the "Berry allocation (%)" x `Inoculation` grid cell
above (Section 2.2) instead of reading it off the 3-way thumbnail: restricted to
**Control-water pots only**, does AMF inoculation change the percent of wheat biomass
allocated to berries? Reuses the exact same fitted model as the grid
(`results_list[["Berry allocation (%)"]]$model`, fit on `W_pcent_berries_logit ~
Inoculation * Polyculture * Water`) - nothing is refit here.

```{r, message=FALSE}
library(scales)   # percent-format display below

berry_spec  <- spec_by_label[["Berry allocation (%)"]]
berry_model <- results_list[["Berry allocation (%)"]]$model
```

## 3.1 The data: every Control-water pot

```{r}
berry_control_data <- berry_spec$data %>%
  dplyr::filter(Water == "Control") %>%
  dplyr::select(Pot, Inoculation, Polyculture, W_pcent_berries) %>%
  dplyr::arrange(Inoculation, Polyculture, Pot) %>%
  dplyr::mutate(`Berry allocation` = scales::percent(W_pcent_berries, accuracy = 0.1)) %>%
  dplyr::select(-W_pcent_berries)

berry_control_data %>%
  knitr::kable(caption = "Control-water pots: berry allocation (%), by pot")
```

## 3.2 Group means (raw % scale)

```{r}
berry_control_summary <- berry_spec$data %>%
  dplyr::filter(Water == "Control") %>%
  dplyr::group_by(Inoculation) %>%
  dplyr::summarise(
    n        = dplyr::n(),
    mean_pct = mean(W_pcent_berries, na.rm = TRUE),
    se_pct   = sd(W_pcent_berries, na.rm = TRUE) / sqrt(n),
    .groups  = "drop"
  ) %>%
  dplyr::mutate(
    `Mean berry allocation` = scales::percent(mean_pct, accuracy = 0.1),
    `± SE`                  = scales::percent(se_pct, accuracy = 0.1)
  ) %>%
  dplyr::select(Inoculation, n, `Mean berry allocation`, `± SE`)

berry_control_summary %>%
  knitr::kable(caption = "Control-water pots: berry allocation (%) by Inoculation, pooled over Polyculture")
```

## 3.3 Formal test: Inoculation effect, sliced to Water = Control

Slices the fitted 3-way model's `Inoculation` comparison at `Water = "Control"` (averaged
over Polyculture) via `emmeans` - this is the model-based version of the raw means above,
i.e. it uses all 60 pots to estimate residual variance, not just the ~20 Control-water
ones. Test statistic/p-value are on the logit scale the model was fit on (valid for
significance); estimated marginal means are back-transformed to % via `plogis()` for
readability.

```{r, message=FALSE}
# suppressMessages() because emmeans() prints a console NOTE ("Results may be
# misleading due to involvement in interactions") any time the model it's
# slicing has interaction terms - expected and harmless here (that's exactly
# why we're slicing to Water = Control instead of reading the marginal
# Inoculation row), not something to surface as if it were a warning about
# this specific result.
berry_emm <- suppressMessages(emmeans::emmeans(berry_model, pairwise ~ Inoculation | Water))

berry_control_emmeans <- as.data.frame(berry_emm$emmeans) %>%
  dplyr::filter(Water == "Control") %>%
  dplyr::mutate(`Estimated mean` = scales::percent(plogis(emmean), accuracy = 0.1)) %>%
  dplyr::select(Inoculation, Water, `Estimated mean`)

berry_control_contrast <- as.data.frame(berry_emm$contrasts) %>%
  dplyr::filter(Water == "Control") %>%
  dplyr::mutate(
    estimate = round(estimate, 3),
    t.ratio  = round(t.ratio, 3),
    p.value  = format_p(p.value)
  )

knitr::kable(berry_control_emmeans, caption = "Model-estimated mean berry allocation, Control water")
knitr::kable(berry_control_contrast, caption = "Inoc vs. Strl contrast (logit scale), Control water")
```

## 3.4 Plot: raw data, Control-water pots only

```{r}
ggplot(berry_spec$data %>% dplyr::filter(Water == "Control"),
       aes(x = Inoculation, y = W_pcent_berries)) +
  geom_boxplot(outlier.shape = NA, width = 0.5, fill = "grey90") +
  geom_jitter(aes(color = Polyculture), width = 0.08, height = 0, size = 2) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(
    title = "Berry allocation by inoculation, Control-water pots",
    x     = "Inoculation",
    y     = "Berry allocation (% of wheat biomass)",
    color = "Polyculture"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

