---
title: "Significance grid - Monoculture only, per-pot variables - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-09-16"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# READ ME FIRST

This is the **per-pot** counterpart of `../PerPlant/Significance_Grid_Mono_DRAFT.Rmd`:
same grid as `Significance_Grid_DRAFT.Rmd` (this folder's full-dataset, per-pot grid),
restricted to **Mono (monoculture) pots only** - both Inoculation levels (Strl and Inoc)
stay in, unlike the Sterile/Inoculated grids, which each drop down to one Inoculation
level but keep both Polyculture levels. Everything about how a cell is built - the
`Yes`/`Kinda (<.1)`/`No` label off BH-adjusted `p_adj`, the direction text, the
thumbnail with its p-value/stars baked in - is identical; only the underlying models
change, because of what dropping down to one polyculture does to the design:

- **Polyculture is no longer a predictor.** Every pot here is Mono, so there's no
  variation left for Polyculture to explain - the 3-way `Inoculation*Polyculture*Water`
  models from the full-dataset grid become 2-way `Inoculation*Water` models. That means
  there's no "Polyculture", "Poly:Inoc", "Poly:Water", or 3-way row in this grid at all -
  they're not blank/`NA`, they simply don't exist as term rows here (`term_labels_for_scope
  ("mono")` only has 3 possible terms: `AMF inoculation`, `Water (PPT)`, `Inoc:Water`).
- **AMF colonization** is fit on Inoc-only Mono pots, with `Water` as the only predictor
  (Polyculture is fixed by this doc's scope; Inoculation is fixed by the Inoc-only
  subsetting - neither has any variation left to test).
- **Fava columns are dropped entirely** - `Fava_g` and `Pot_pcent_fava` are exactly 0 for
  every Mono pot by definition (no fava plant present), so there's zero variance for an
  ANOVA to test. They don't appear as `NA` columns here; they're just not part of this
  scope's `battery_spec` at all.
- **"Total pot productivity" is dropped too**, along with the Fava columns - it's
  numerically identical to "Wheat total NPP" for this scope (max difference ~5e-5 g,
  i.e. rounding), since a Mono pot's total biomass IS its wheat biomass with no fava
  plant to add. Per your call, it's excluded here rather than kept as a redundant
  column.
- Mass-family variables are fit on the same **pot-total** biomass columns as
  `Significance_Grid_DRAFT.Rmd` (`W_NPP`, `W_ANPP_g`, `W_BNPP_g`, `W_Berries_g`), not
  the density-adjusted `_pp` columns.

All of this comes from `build_battery_spec("mono")` in this folder's `battery_core.R` -
nothing here is hand-adjusted per variable.

```{r, message=FALSE}
source("battery_core.R")   # combined/inoc_only/sterile_only/inter_only/mono_only,
                             # build_battery_spec(), run_battery()
source("grid_helpers.R")    # plot_cell(), direction_label(), build_cell_data(), render_grid()
```

```{r}
# This scope's own spec + fitted battery - NOT the top-level battery_spec/
# term_summary that battery_core.R also builds (those are scope = "full").
battery_spec_mono <- build_battery_spec("mono")
batt_mono          <- run_battery(battery_spec_mono)
```

```{r}
# Own plot directory, separate from the other grids', so identically-named
# cells (e.g. a "Water" main-effect thumbnail exists in every grid) don't
# overwrite each other on disk.
plot_dir <- "plots/significance_grid_plots_mono"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. Build the cell data (one row per response variable x ANOVA term)

```{r}
cell_data <- build_cell_data(batt_mono$term_summary, batt_mono$spec_by_label, batt_mono$results_list, plot_dir)
```

# 2. Render the two grids (mass / percent)

```{r}
term_labels <- term_labels_for_scope("mono")
```

## 2.1 Mass variables (grams, per pot)

```{r}
render_grid("mass", cell_data, battery_spec_mono, term_labels)
```

## 2.2 Percent / allocation variables

```{r}
render_grid("percent", cell_data, battery_spec_mono, term_labels)
```
