---
title: "Significance grid - Inoculated pots only, per-pot variables - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-09-16"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# READ ME FIRST

Same grid as `Significance_Grid_DRAFT.Rmd` (this folder's full-dataset, per-pot grid),
restricted to **Inoculated pots only** - the counterpart of
`Significance_Grid_Sterile_DRAFT.Rmd`, which restricts to Sterile pots instead.
Everything about how a cell is built - the `Yes`/`Kinda (<.1)`/`No` label off
BH-adjusted `p_adj`, the direction text, the thumbnail with its p-value/stars baked in -
is identical; only the underlying models change, because of what dropping down to one
inoculation level does to the design:

- **Inoculation is no longer a predictor.** Every pot here is Inoc, so there's no
  variation left for Inoculation to explain - the full dataset's 3-way
  `Inoculation*Polyculture*Water` models become 2-way `Polyculture*Water` models. That
  means there's no "AMF inoculation", "Poly:Inoc", "Inoc:Water", or 3-way row in this
  grid at all - they're not blank/`NA`, they simply don't exist as term rows here
  (`term_labels_for_scope("inoc")` only has 3 possible terms: `Polyculture`,
  `Water (PPT)`, `Poly:Water`).
- **AMF colonization (Hyphal / Arbuscular-vesicular) is kept, unchanged in what it's
  fit on** - unlike the Sterile grid, which drops these two variables entirely, they
  were already Inoc-only in every other scope (a structural zero for Strl pots), so
  restricting to Inoc pots here doesn't lose any data for them - `Polyculture*Water`
  is exactly the design they've always been fit on.
- **Fava columns are still included** - every Inter pot has a fava plant regardless of
  inoculation, so `Fava biomass` and `Fava % of pot biomass` are still meaningful here,
  modeled on Inoc Inter pots as `~ Water` (both Polyculture and Inoculation are fixed
  by this doc's scope + the Inter-only subsetting).
- Mass-family variables are fit on the same **pot-total** biomass columns as
  `Significance_Grid_DRAFT.Rmd` (`W_NPP`, `W_ANPP_g`, `W_BNPP_g`, `W_Berries_g`,
  `Pot_NPP_g`), not the density-adjusted `_pp` columns.

All of this comes from `build_battery_spec("inoc")` in this folder's `battery_core.R` -
nothing here is hand-adjusted per variable.

```{r, message=FALSE}
source("battery_core.R")   # combined/inoc_only/sterile_only/inter_only/mono_only,
                             # build_battery_spec(), run_battery()
source("grid_helpers.R")    # plot_cell(), direction_label(), build_cell_data(), render_grid()
```

```{r}
# This scope's own spec + fitted battery - NOT the top-level battery_spec/
# term_summary that battery_core.R also builds (those are scope = "full").
battery_spec_inoc <- build_battery_spec("inoc")
batt_inoc          <- run_battery(battery_spec_inoc)
```

```{r}
# Own plot directory, separate from the other grids', so identically-named
# cells (e.g. a "Water" main-effect thumbnail exists in every grid) don't
# overwrite each other on disk.
plot_dir <- "plots/significance_grid_plots_inoculated"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. Build the cell data (one row per response variable x ANOVA term)

```{r}
cell_data <- build_cell_data(batt_inoc$term_summary, batt_inoc$spec_by_label, batt_inoc$results_list, plot_dir)
```

# 2. Render the two grids (mass / percent)

```{r}
term_labels <- term_labels_for_scope("inoc")
```

## 2.1 Mass variables (grams, per pot)

```{r}
render_grid("mass", cell_data, battery_spec_inoc, term_labels)
```

## 2.2 Percent / allocation variables

```{r}
render_grid("percent", cell_data, battery_spec_inoc, term_labels)
```

# 3. Focused figure: Fava biomass & Fava % of pot biomass, by Water

Pulls the "Fava biomass" and "Fava % of pot biomass" x `Water` grid cells above (Section
2) into one combined, labeled, full-size figure instead of reading the two tiny
thumbnails separately - same underlying data and `~ Water` model as the grid (Inoc Inter
pots only), nothing refit here, just replotted bigger and side by side for direct
comparison.

```{r, message=FALSE}
library(patchwork)   # combines the two ggplots into one labeled figure
library(scales)      # percent-format y-axis for the second panel
```

```{r}
# One mean +/- SE bar plot by Water, full-size (real axis titles, not the
# blank ones the grid's compact thumbnails use) - same p-value/stars
# convention as the grid (format_p()/p_to_stars() from grid_helpers.R),
# keyed off this variable's BH-adjusted p_adj for the Water term specifically.
fava_water_plot <- function(label, y_var, y_lab, percent_axis = FALSE) {
  spec  <- batt_inoc$spec_by_label[[label]]
  p_val <- batt_inoc$term_summary %>%
    dplyr::filter(label == !!label, term == "Water") %>%
    dplyr::pull(p_adj)

  summary_df <- spec$data %>%
    dplyr::group_by(Water) %>%
    dplyr::summarise(
      mean_y = mean(.data[[y_var]], na.rm = TRUE),
      se_y   = sd(.data[[y_var]], na.rm = TRUE) / sqrt(sum(!is.na(.data[[y_var]]))),
      .groups = "drop"
    )

  p <- ggplot(summary_df, aes(x = Water, y = mean_y)) +
    theme_minimal(base_size = 12) +
    geom_col(fill = "grey30", width = 0.6) +
    geom_errorbar(aes(ymin = mean_y - se_y, ymax = mean_y + se_y), width = 0.15) +
    expand_limits(y = 0) +
    labs(
      title    = label,
      subtitle = paste0(format_p(p_val), "  ", p_to_stars(p_val)),
      x        = "Water (PPT)",
      y        = y_lab
    ) +
    theme(
      plot.title    = element_text(face = "bold", size = 13),
      plot.subtitle = element_text(
        colour = ifelse(p_val < 0.05, "black", "grey40"),
        face   = ifelse(p_val < 0.05, "bold", "plain")
      )
    )

  if (percent_axis) p <- p + scale_y_continuous(labels = scales::percent)
  p
}

fava_biomass_plot <- fava_water_plot("Fava biomass", "Fava_g", "Fava biomass (g)")
fava_pcent_plot   <- fava_water_plot("Fava % of pot biomass", "Pot_pcent_fava",
                                      "Fava % of pot biomass", percent_axis = TRUE)

fava_biomass_plot + fava_pcent_plot +
  patchwork::plot_annotation(
    title    = "Inoculated pots only, per-pot variables",
    tag_levels = "A"
  )
```

# 4. Focused figure: Wheat berry biomass & Berry allocation (%), by Polyculture:Water

Pulls the "Wheat berry biomass" and "Berry allocation (%)" x `Polyculture:Water` grid
cells above (Section 2) into one combined, labeled, full-size figure - same underlying
data and interaction model as the grid (Inoc pots only), nothing refit. Each panel adds
a **pairwise Inter-vs-Mono comparison within each Water level**, shown as letters next
to each point: the same letter on both lines at a given Water level means Inter and
Mono are *not* significantly different there (p >= .05); different letters mean they
are (p < .05). These within-Water contrasts come from slicing the same fitted
`Polyculture*Water` model via `emmeans::emmeans(model, pairwise ~ Polyculture | Water)`
- the model-based equivalent of the grid's own `Polyculture:Water` term, just broken out
one Water level at a time instead of read off the omnibus interaction p-value (which is
what the panel subtitle still shows).

```{r}
# Full-size Polyculture x Water interaction plot (real axis titles + legend,
# not the grid's blank-axis thumbnail), with Inter-vs-Mono significance
# letters layered on top at each Water level.
#   - label:         battery_spec label, e.g. "Wheat berry biomass"
#   - plot_response: which column to actually PLOT (raw scale) - matches
#     spec$plot_response in battery_core.R, e.g. the untransformed
#     proportion for Berry allocation (%), whose model is fit on the logit
#     column instead.
poly_water_panel <- function(label, plot_response, y_lab, percent_axis = FALSE) {
  spec  <- batt_inoc$spec_by_label[[label]]
  model <- batt_inoc$results_list[[label]]$model
  data  <- spec$data

  p_val <- batt_inoc$term_summary %>%
    dplyr::filter(label == !!label, term == "Polyculture:Water") %>%
    dplyr::pull(p_adj)

  summary_df <- data %>%
    dplyr::group_by(Water, Polyculture) %>%
    dplyr::summarise(
      mean_y = mean(.data[[plot_response]], na.rm = TRUE),
      se_y   = sd(.data[[plot_response]], na.rm = TRUE) / sqrt(sum(!is.na(.data[[plot_response]]))),
      .groups = "drop"
    )

  # Inter-vs-Mono contrast, sliced to each Water level, off the model itself
  # (whatever scale it was fit on - logit for percent-family, raw grams for
  # biomass) - suppressMessages() for the same harmless emmeans NOTE noted
  # elsewhere in this project ("Results may be misleading due to involvement
  # in interactions"), expected here since that's exactly why we're slicing.
  emm_contrasts <- suppressMessages(
    emmeans::emmeans(model, pairwise ~ Polyculture | Water)
  )$contrasts %>%
    as.data.frame()

  # With exactly two Polyculture levels, a compact letter display collapses
  # to: same letter if that Water level's contrast isn't significant,
  # different letters (higher raw mean = "a") if it is.
  letters_df <- summary_df %>%
    dplyr::group_by(Water) %>%
    dplyr::group_modify(function(sub, key) {
      sig <- emm_contrasts$p.value[emm_contrasts$Water == key$Water] < 0.05
      sub <- sub %>% dplyr::arrange(dplyr::desc(mean_y))
      sub$letter <- if (isTRUE(sig)) c("a", "b") else c("a", "a")
      sub
    }) %>%
    dplyr::ungroup()

  # Letters sit just above each point's own error bar - padding is a fixed
  # fraction of this panel's y-range so it scales sensibly whether the axis
  # is raw grams or a 0-1 proportion.
  y_span <- diff(range(c(summary_df$mean_y - summary_df$se_y, summary_df$mean_y + summary_df$se_y)))
  pad    <- y_span * 0.08

  # Small horizontal dodge so the two Polyculture series (and their letters)
  # don't sit exactly on top of each other at a given Water level - without
  # this, a letter for one line can visually collide with the OTHER line's
  # error bar whenever the two means are close together.
  dodge <- ggplot2::position_dodge(width = 0.15)

  p <- ggplot(summary_df, aes(x = Water, y = mean_y, color = Polyculture, group = Polyculture)) +
    theme_minimal(base_size = 12) +
    geom_line(position = dodge, linewidth = 0.8) +
    geom_point(position = dodge, size = 2.5) +
    geom_errorbar(aes(ymin = mean_y - se_y, ymax = mean_y + se_y), position = dodge, width = 0.1) +
    geom_text(
      data        = letters_df,
      aes(x = Water, y = mean_y + se_y + pad, label = letter, color = Polyculture),
      position    = dodge,
      fontface    = "bold",
      size        = 4,
      show.legend = FALSE
    ) +
    expand_limits(y = c(0, max(summary_df$mean_y + summary_df$se_y) + 2 * pad)) +
    labs(
      title    = label,
      subtitle = paste0(format_p(p_val), "  ", p_to_stars(p_val)),
      x        = "Water (PPT)",
      y        = y_lab,
      color    = "Polyculture"
    ) +
    theme(
      legend.position = "bottom",
      plot.title      = element_text(face = "bold", size = 13),
      plot.subtitle   = element_text(
        colour = ifelse(p_val < 0.05, "black", "grey40"),
        face   = ifelse(p_val < 0.05, "bold", "plain")
      )
    )

  if (percent_axis) p <- p + scale_y_continuous(labels = scales::percent)
  p
}

berry_biomass_panel    <- poly_water_panel("Wheat berry biomass", "W_Berries_g",
                                            "Wheat berry biomass (g)")
berry_allocation_panel <- poly_water_panel("Berry allocation (%)", "W_pcent_berries",
                                            "Berry allocation (% of wheat biomass)",
                                            percent_axis = TRUE)

berry_biomass_panel + berry_allocation_panel +
  patchwork::plot_annotation(
    title      = "Inoculated pots only, per-pot variables",
    tag_levels = "A"
  )
```
