---
title: "Biomass_Basic"
author: "Celeste Basken"
date: "2026-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}

library(tidyverse)

```


## Berries 

Clustering without regard to PPT:

```{r pressure, echo=FALSE}
# Without sorting by ppt at all

berries <- read.csv("Biomass/Berry_Weights.csv")

berries <- berries %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Group = interaction(Inoculation, Polyculture, sep = "_")
  )

ggplot(berries, aes(x = Group, y = Biomass_g, fill = Group)) + 
  geom_col()


```

# With every replicate individually
This shows the variation by ppt if you just look at inoc and inter status

```{r, echo=FALSE}
# With each bar individual; as you can see there is a lot of variation, hopefully due to ppt

berries <- read.csv("Biomass/Berry_Weights.csv")

berries <- berries %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Group = interaction(Inoculation, Polyculture, sep = "_"),
    Rep = factor(row_number())
  )

ggplot(berries, aes(x = Group, y = Biomass_g, fill = Group)) +
  geom_col(aes(group = Rep), position = position_dodge(width = 0.8))

```
# Clustered with ppt included
There appears to be different behavior in the water status between inter and mono.

```{r, echo=FALSE}
# With each bar individual and ppt

berries <- read.csv("Biomass/Berry_Weights.csv")

berries <- berries %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Water = factor(Water, levels = c("Control", "Moderate", "Severe")),
    Group = interaction(Inoculation, Polyculture, sep = "_")
  )

ggplot(berries, aes(x = Group, y = Biomass_g, fill = Water)) +
  geom_col(position = "dodge")

```

The just water graph is a little concerning because there appears to be no difference between control and moderate...

```{r, echo=FALSE}
# Just water

water_means <- berries %>%
  mutate(
    Water = factor(Water, levels = c("Control", "Moderate", "Severe"))
  ) %>%
  group_by(Water) %>%
  summarise(mean_biomass = mean(Biomass_g, na.rm = TRUE),
            .groups = "drop")

ggplot(water_means, aes(x = Water, y = mean_biomass)) +
  geom_col()

```

