---
title: "Berries_initialstats"
author: "Celeste Basken"
date: "2026-02-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}

library(tidyverse)
library(broom)
library(knitr)
library(dplyr)
```

# Berries 

```{r}
# Loading in data to explore
berries <- read.csv("../Biomass/Berry_Weights.csv")

berries <- berries %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Group = interaction(Inoculation, Polyculture, sep = "_")
  )

# Adding a column for continuous water weights: based on the assumption that the emitters are 2 gal/hr 

berries <- berries %>%
  mutate(
    Water_gal_per_wk = case_when(
      Water == "Control"  ~ 48,
      Water == "Moderate" ~ 32,
      Water == "Severe"   ~ 16,
      TRUE ~ NA_real_
    )
  )

```


## Berry statistics

To conclude my berry work, let's generate some statistics

Before I begin I want to check some assumptions of a t-test!!

The boxes shouldn't have overly long tails or outliers. The residuals should be straight. These assumptions seem to be met (??). 

* I worry that that the residuals look too S-shaped

```{r}
# Check of distributions
ggplot(berries, aes(x = Inoculation, y = Biomass_g_pp)) +
  geom_boxplot() +
  geom_jitter(width = 0.15, alpha = 0.6) +
  facet_wrap(~ Water) +
  theme_bw() +
  ggtitle("Berry biomass by inoculation status")

# Q–Q plots of residuals
ggplot(berries, aes(sample = Biomass_g_pp)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ Inoculation + Water) +
  theme_bw() +
  ggtitle("QQ plot of residuals for inoc status")

```

This next section is generating a variety of summary statistics about the data, and then saving them as CSVs. 

```{r}

# Load + clean
berries <- berries %>%
  mutate(
    Inoculation = factor(Inoculation, levels = c("Strl", "Inoc")),
    Polyculture = factor(Polyculture, levels = c("Mono", "Inter")),
    Water = factor(Water, levels = c("Control", "Moderate", "Severe"))
  ) %>%
  filter(!is.na(Biomass_g_pp))

# Descriptive statistics
desc_stats <- function(x) {
  tibble(
    n      = sum(!is.na(x)),
    mean   = mean(x, na.rm = TRUE),
    sd     = sd(x, na.rm = TRUE),
    se     = sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x))),
    median = median(x, na.rm = TRUE),
    q25    = quantile(x, 0.25, na.rm = TRUE, names = FALSE),
    q75    = quantile(x, 0.75, na.rm = TRUE, names = FALSE),
    iqr    = IQR(x, na.rm = TRUE),
    min    = min(x, na.rm = TRUE),
    max    = max(x, na.rm = TRUE)
  )
}

# Each of the 4 treatments within each Water level
summary_4treat_by_water <- berries %>%
  group_by(Water, Inoculation, Polyculture) %>%
  summarise(desc_stats(Biomass_g_pp), .groups = "drop")

# Inoculation only (fold inter/mono together), within each Water
summary_inoc_by_water <- berries %>%
  group_by(Water, Inoculation) %>%
  summarise(desc_stats(Biomass_g_pp), .groups = "drop")

# Polyculture only (fold inoc/strl together), within each Water
summary_poly_by_water <- berries %>%
  group_by(Water, Polyculture) %>%
  summarise(desc_stats(Biomass_g_pp), .groups = "drop")

# Oerall summaries across Water (all data)
summary_4treat_overall <- berries %>%
  group_by(Inoculation, Polyculture) %>%
  summarise(desc_stats(Biomass_g_pp), .groups = "drop")

summary_inoc_overall <- berries %>%
  group_by(Inoculation) %>%
  summarise(desc_stats(Biomass_g_pp), .groups = "drop")

summary_poly_overall <- berries %>%
  group_by(Polyculture) %>%
  summarise(desc_stats(Biomass_g_pp), .groups = "drop")


# Looking at them as tables
summary_4treat_by_water
summary_inoc_by_water
summary_poly_by_water

# Saving these as csvs

write.csv(summary_4treat_by_water,
          file.path(("../Biomass/Berries"), "summary_4treat_by_water.csv"),
          row.names = FALSE)

write.csv(summary_inoc_by_water,
          file.path(("../Biomass/Berries"), "summary_inoc_by_water.csv"),
          row.names = FALSE)

write.csv(summary_poly_by_water,
          file.path(("../Biomass/Berries"), "summary_poly_by_water.csv"),
          row.names = FALSE)

write.csv(summary_4treat_overall,
          file.path(("../Biomass/Berries"), "summary_4treat_overall.csv"),
          row.names = FALSE)

write.csv(summary_inoc_overall,
          file.path(("../Biomass/Berries"), "summary_inoc_overall.csv"),
          row.names = FALSE)

write.csv(summary_poly_overall,
          file.path(("../Biomass/Berries"), "summary_poly_overall.csv"),
          row.names = FALSE)


```

## Berry statistical analysis

I used AI to generate stats. This should be verified for sure!! Especially to make sure I meet the assumptions to even be using these stats.

```{r}
# These statistics were Chat GPT and should be verified

# P-VALUES: Simple comparisons within each Water ----
# Welch t-tests are robust to unequal variances (default in t.test)

# a) Inoculation effect within each Water (folding Polyculture)
ttest_inoc_within_water <- berries %>%
  group_by(Water) %>%
  do(tidy(t.test(Biomass_g_pp ~ Inoculation, data = .))) %>%
  ungroup() %>%
  select(Water, estimate1, estimate2, statistic, p.value, conf.low, conf.high, method)

# b) Polyculture effect within each Water (folding Inoculation)
ttest_poly_within_water <- berries %>%
  group_by(Water) %>%
  do(tidy(t.test(Biomass_g_pp ~ Polyculture, data = .))) %>%
  ungroup() %>%
  select(Water, estimate1, estimate2, statistic, p.value, conf.low, conf.high, method)

# ---- 3-way factorial model (lets you test interactions) ----
# If your data are fairly normal-ish, this is a strong default.
fit_aov <- aov(Biomass_g_pp ~ Inoculation * Polyculture * Water, data = berries)
aov_table <- broom::tidy(fit_aov)

# ---- Optional: Nonparametric alternatives (in case assumptions are shaky) ----
# Wilcoxon rank-sum within each Water for Inoculation / Polyculture
wilcox_inoc_within_water <- berries %>%
  group_by(Water) %>%
  do(tidy(wilcox.test(Biomass_g_pp ~ Inoculation, data = ., exact = FALSE))) %>%
  ungroup() %>%
  select(Water, statistic, p.value, method)

wilcox_poly_within_water <- berries %>%
  group_by(Water) %>%
  do(tidy(wilcox.test(Biomass_g_pp ~ Polyculture, data = ., exact = FALSE))) %>%
  ungroup() %>%
  select(Water, statistic, p.value, method)

# Print these

ttest_inoc_within_water
ttest_poly_within_water

aov_table

# Saving as CSV

write.csv(ttest_inoc_within_water,
          file.path(("../Biomass/Berries/Stats"), "ttest_inoc_within_water.csv"),
          row.names = FALSE)

write.csv(ttest_poly_within_water,
          file.path(("../Biomass/Berries/Stats"), "ttest_poly_within_water.csv"),
          row.names = FALSE)

write.csv(aov_table,
          file.path(("../Biomass/Berries/Stats"), "anova_3way_table.csv"),
          row.names = FALSE)

write.csv(wilcox_inoc_within_water,
          file.path(("../Biomass/Berries/Stats"), "wilcox_inoc_within_water.csv"),
          row.names = FALSE)

write.csv(wilcox_poly_within_water,
          file.path(("../Biomass/Berries/Stats"), "wilcox_poly_within_water.csv"),
          row.names = FALSE)


```


```{r}
an1 <- aov(Biomass_g_pp~Polyculture+Inoculation+Water, data= berries)
summary(an1)

an2 <- aov(Biomass_g_pp~Polyculture*Inoculation*Water, data= berries)
summary(an2)

lm2 <- lm(Biomass_g_pp~Polyculture*Inoculation*Water, data= berries)
summary(lm2)
```


