---
title: "Step 6 - Exploratory Analyses"
subtitle: "AMF x Drought x Neighborhood in Hard Red Spring Wheat"
author: "CLAUDE + Celeste Basken"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Step 6 from `Instructions/AMF_intercropping_analysis_plan.md`, Sections 5 and 6. **Everything
in this document is exploratory**, per the plan's own pre-specification (Section 8) - none
of it is part of the three confirmatory tests (Steps 2-4). Association language only where
the predictor wasn't randomized (`hyphal`, `fava_npp`) - `hyphal_colonization` and
`fava_npp` are outcomes of the experiment, not treatments, so "was associated with," not
"caused."

```{r}
library(tidyverse)
library(car)
library(glmmTMB)
library(DHARMa)
library(emmeans)

wheat <- readRDS("wheat_audited.rds")
```

# Part A: Colonization associations (Section 5 of the plan)

Three problems the plan flags with how this was done previously: a bimodal predictor
(Strl piled near 0, Inoc near 1) when Inoc and Strl are combined; comparing p-values across
separate water-level panels instead of testing an interaction term; and colonization being
observed, not randomized (bigger/healthier plants may support more hyphae - reverse
causation is live). Fixed here by fitting within Inoc-only pots, with `hyphal:Water` as the
designated test.

## Two corrections to the plan's own framing, checked against the actual data

```{r}
strl <- wheat %>% filter(Inoc == "Strl")
cat("Background colonization in Strl pots: range", round(min(strl$hyphal)*100,1), "-",
    round(max(strl$hyphal)*100,1), "%  (plan states 0-30%)\n")
strl_high <- strl %>% filter(hyphal == max(hyphal))
```

The plan describes background colonization in sterile pots as "0-30%" - **the actual range
is 0-58%.** One pot (`r strl_high$Name`) sits well above the rest at
`r round(strl_high$hyphal*100,0)`%, likely cross-contamination during handling/scoring
rather than a systematic sterility failure (every other Strl pot the value is below the
plan's estimate). Doesn't change the direction of the conservative-bias argument below,
worth stating precisely rather than repeating the plan's number unchecked.

```{r}
inoc <- wheat %>% filter(Inoc == "Inoc")
bimodal_check <- inoc %>% group_by(Water) %>%
  summarise(mean = round(mean(hyphal), 2), sd = round(sd(hyphal), 2),
            min = round(min(hyphal), 2), max = round(max(hyphal), 2))
knitr::kable(bimodal_check, caption = "hyphal colonization within Inoc pots only, by Water")
```

**Within the Inoc-only subset used for the association models below, colonization is not
bimodal** - the bimodality the plan describes is a property of the *combined* Inoc+Strl
distribution (piling near 0 for Strl, near 1 for Inoc), which is exactly why these models
are restricted to Inoc pots only in the first place. Within Inoc pots, Control and Severe
have real spread (0.46-1.0, 0.66-0.99); Moderate is tightly compressed near ceiling
(0.91-1.0, sd=0.03) - a power problem specific to that one water level's slope, not a
bimodality problem.

## The association model

```{r}
col <- inoc
contrasts(col$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))

m_col <- lm(berry_alloc ~ hyphal * Water * Poly, data = col)
col_tbl <- car::Anova(m_col, type = 2) %>% broom::tidy() %>% filter(term != "Residuals")
knitr::kable(col_tbl %>% mutate(across(c(statistic, p.value), ~round(.x, 4))),
             caption = "berry_alloc ~ hyphal*Water*Poly (OLS)")
cat("Shapiro-Wilk on residuals: p =", round(shapiro.test(residuals(m_col))$p.value, 3), "(clean)\n")
```

```{r}
m_col_beta <- glmmTMB(berry_alloc ~ hyphal * Water * Poly, family = beta_family(), data = col)
beta_tbl <- car::Anova(m_col_beta, type = 2) %>% broom::tidy()
sim <- simulateResiduals(m_col_beta, n = 1000)
knitr::kable(beta_tbl %>% mutate(across(c(statistic, p.value), ~round(.x, 4))),
             caption = "Cross-check: same model as beta regression")
cat("DHARMa uniformity p:", round(testUniformity(sim, plot=FALSE)$p.value,3), "(clean)\n")
```

**The designated test, `hyphal:Water`, is not significant by OLS (p =
`r round(col_tbl$p.value[col_tbl$term=='hyphal:Water'],3)`) and only marginal by beta
regression (p = `r round(beta_tbl$p.value[beta_tbl$term=='hyphal:Water'],3)`).** No clear
association between realized colonization level and berry allocation, beyond what
Inoculation status itself already captures (Steps 2-4). Consistent with the plan's own
expectation that this section would likely be weaker evidence than the randomized contrast.

## On the sterile control (manuscript language, not a test)

Two separate limitations, worth keeping distinct in the discussion:

1. **Imperfect sterility** (0-58% background colonization in nominally sterile pots, one
   pot likely cross-contaminated) shrinks the Inoc-Strl contrast and biases it toward the
   null - i.e., the treatment effects reported in Steps 2-4 are conservative, not inflated.
2. **The inoculum isn't AMF-only.** Inoculated pots received live INVAM material (spores,
   hyphae, root fragments, growth medium, and whatever non-AMF microbiota rode along);
   sterile pots received nothing. The contrast is *live inoculum vs. no inoculum*, not
   *AMF vs. no AMF* - state this explicitly rather than let "AMF effect" imply more
   specificity than the design supports.

# Part B: Fava analyses (Section 6 of the plan) - Inter pots, Inoc-only, n=15

```{r}
fava_inoc <- wheat %>% filter(Poly == "Inter", Inoc == "Inoc")
contrasts(fava_inoc$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))
cat("n =", nrow(fava_inoc), "(2 Poly excluded by design, 1 Inoc level, 3 Water x 5 reps = 15)\n")
```

Only wheat roots were stained - fava's own mycorrhizal status is unmeasured, so this section
tests whether the *strength of the wheat-AMF partnership* (via wheat's own colonization
level) predicts fava suppression, not fava's own colonization.

```{r}
m_f <- lm(log(fava_npp) ~ hyphal * Water, data = fava_inoc)
f_tbl <- car::Anova(m_f, type = 2) %>% broom::tidy() %>% filter(term != "Residuals")
knitr::kable(f_tbl %>% mutate(across(c(statistic, p.value), ~round(.x, 4))),
             caption = "log(fava_npp) ~ hyphal*Water")
cat("Shapiro p:", round(shapiro.test(residuals(m_f))$p.value, 3), "\n\n")

m_f2 <- lm(fava_pct_pot ~ hyphal * Water, data = fava_inoc)
f2_tbl <- car::Anova(m_f2, type = 2) %>% broom::tidy() %>% filter(term != "Residuals")
knitr::kable(f2_tbl %>% mutate(across(c(statistic, p.value), ~round(.x, 4))),
             caption = "fava_pct_pot ~ hyphal*Water")
cat("Shapiro p:", round(shapiro.test(residuals(m_f2))$p.value, 3), "\n")
```

**`hyphal:Water` not significant for either response** (p =
`r round(f_tbl$p.value[f_tbl$term=='hyphal:Water'],3)`,
`r round(f2_tbl$p.value[f2_tbl$term=='hyphal:Water'],3)`). Per the plan: exploratory, full
stop, at n~5/water level once restricted to Inoc pots.

## Correction to the plan's range-restriction speculation

```{r}
range_check <- fava_inoc %>% group_by(Water) %>%
  summarise(mean = round(mean(fava_npp), 2), sd = round(sd(fava_npp), 2),
            min = round(min(fava_npp), 2), max = round(max(fava_npp), 2))
knitr::kable(range_check, caption = "fava_npp (g) by Water, Inoc-only Inter pots")
```

The plan speculated the flat slope under severe drought might reflect "fava NPP compressed
into 3-5 g" - **checked, and the compression is real (Severe has the tightest SD of the
three levels) but at a *higher*, not lower, mean than Control/Moderate** within this
Inoc-only subset (3.69g vs 2.4g/2.01g). Worth stating precisely - see Part C below for why
this reverses once Strl pots are included.

# Part C: Does AMF status itself predict fava suppression? (new test, not in the plan's literal snippets)

Motivated directly by the plan's own framing in Section 6 ("the strength of the wheat-AMF
partnership predicts fava suppression") - Part B tests this via *continuous* colonization
within Inoc pots only; this tests the *categorical* Inoc/Strl contrast across all 30 Inter
pots, which turns out to show a much clearer pattern.

```{r}
inter <- wheat %>% filter(Poly == "Inter")
contrasts(inter$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))

supp_desc <- inter %>% group_by(Inoc, Water) %>%
  summarise(mean_fava = round(mean(fava_npp), 2), sd = round(sd(fava_npp), 2), n = n(), .groups = "drop")
knitr::kable(supp_desc, caption = "fava_npp (g) by Inoc x Water, all 30 Inter pots")
```

```{r}
m_supp <- lm(log(fava_npp) ~ Inoc * Water, data = inter)
supp_tbl <- car::Anova(m_supp, type = 2) %>% broom::tidy() %>% filter(term != "Residuals")
knitr::kable(supp_tbl %>% mutate(across(c(statistic, p.value), ~round(.x, 4))),
             caption = "log(fava_npp) ~ Inoc*Water")

sw_supp <- shapiro.test(residuals(m_supp))
cat("Shapiro-Wilk: p =", round(sw_supp$p.value, 3), "- flagged, checked below, not ignored\n")
```

## Checking the flagged assumption violation before trusting the result

```{r}
inter$resid <- residuals(m_supp)
outlier_pot <- inter %>% arrange(desc(abs(resid))) %>% slice(1)
cat("Most extreme residual:", outlier_pot$Name, "- fava_npp =", outlier_pot$fava_npp,
    "g, far below the other 4 pots in its Inoc x Water cell\n")
print(car::outlierTest(m_supp))
```

```{r}
inter_no_out <- inter %>% filter(Name != outlier_pot$Name)
contrasts(inter_no_out$Water) <- cbind(severe = c(-0.5,-0.5,1), mild = c(-1,1,0))
m_supp2 <- lm(log(fava_npp) ~ Inoc*Water, data = inter_no_out)
cat("Without the flagged pot: Inoc:Water p =",
    format.pval(car::Anova(m_supp2,type=2)["Inoc:Water","Pr(>F)"]),
    "| Shapiro p =", round(shapiro.test(residuals(m_supp2))$p.value,3), "(normal once removed)\n\n")

kw <- kruskal.test(fava_npp ~ interaction(Inoc, Water), data = inter)
cat("Distribution-free cross-check (Kruskal-Wallis, all 6 groups): p =", format.pval(kw$p.value), "\n")
```

**One genuine outlier** (Bonferroni-significant), but removing it makes the result
*stronger*, not weaker (p drops further, normality is restored), and a fully
distribution-free test on the same 6 groups agrees. This is a robust finding, not an
artifact of the one low pot.

```{r}
emm_supp <- emmeans(m_supp, ~ Inoc | Water, type = "response")
ratio_tbl <- contrast(emm_supp, "revpairwise") %>% as.data.frame()
knitr::kable(ratio_tbl %>% mutate(across(c(ratio,SE,t.ratio,p.value), ~round(.x,3))),
             caption = "Strl/Inoc fava-mass ratio by Water (back-transformed from log scale)")
```

```{r, fig.width=6, fig.height=4}
ggplot(inter, aes(x = Water, y = fava_npp, fill = Inoc)) +
  geom_boxplot(position = position_dodge(0.8), width = 0.6) +
  geom_point(position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), size = 1.5) +
  labs(title = "AMF-inoculated wheat suppresses fava under good water - not under severe drought",
       y = "Fava biomass (g)") +
  theme_minimal(base_size = 12)
```

**The clearest exploratory finding in this section:** AMF-inoculated wheat suppresses fava
by ~3.6-3.9x relative to sterile wheat under Control and Moderate water (both p<0.0001) -
and that suppression essentially **vanishes** under severe drought (ratio = 1.03x,
p=0.86). Competitive suppression by inoculated wheat is strong when water isn't limiting,
and disappears when it is - directionally consistent with the stress-gradient framing in
Section 12 of the plan (competition -> facilitation/neutral as stress increases), though
this specific test wasn't one of the plan's pre-specified confirmatory contrasts.

# Part D: The fava-biomass moderator probe (Section 6) - facilitation vs. density release

```{r}
m_mech <- glmmTMB(berry_alloc ~ fava_npp * Inoc * Water, family = beta_family(), data = inter)
mech_tbl <- car::Anova(m_mech, type = 2) %>% broom::tidy()
knitr::kable(mech_tbl %>% mutate(across(c(statistic, p.value), ~round(.x, 4))),
             caption = "berry_alloc ~ fava_npp*Inoc*Water")

sim2 <- simulateResiduals(m_mech, n = 1000)
cat("DHARMa uniformity p:", round(testUniformity(sim2, plot=FALSE)$p.value,3),
    "| dispersion p:", round(testDispersion(sim2, plot=FALSE)$p.value,3), "(clean)\n")
```

**Interpretation, per the plan's own decision rule:**

- `fava_npp` main effect is strongly significant (p =
  `r format.pval(mech_tbl$p.value[mech_tbl$term=='fava_npp'])`) - bigger fava is associated
  with better wheat berry allocation, **but this holds regardless of Inoculation status**
  (`fava_npp:Inoc` is not significant, p =
  `r round(mech_tbl$p.value[mech_tbl$term=='fava_npp:Inoc'],3)`, nor is the three-way
  `fava_npp:Inoc:Water`, p = `r round(mech_tbl$p.value[mech_tbl$term=='fava_npp:Inoc:Water'],3)`).
- Per the plan's own framing: "if larger fava -> stronger rescue of the inoculation penalty,
  that's evidence for facilitation; if larger fava -> weaker rescue, competition dominates
  and the mono/inter difference is more likely density release." **Neither direction is
  detected here** - the association between fava size and wheat allocation doesn't
  distinguish the two stories, because it doesn't specifically depend on Inoc status at all.
  This data doesn't support fava-specific facilitation over the simpler density-release
  explanation, though it doesn't rule facilitation out either - it's just not visible in
  this particular test.
- Not causal either way (fava size wasn't randomized, is collinear with wheat performance) -
  reported as exploratory/associational per the plan, consistent with every other section
  above.

# Summary

```{r}
summary_tbl <- tibble::tribble(
  ~section, ~test, ~result,
  "A. Colonization association", "hyphal:Water on berry_alloc", "Not significant (OLS); marginal (beta regression)",
  "B. Fava vs. wheat colonization", "hyphal:Water on fava_npp / fava_pct_pot", "Not significant, either response",
  "C. Fava suppression by Inoc status (new)", "Inoc:Water on fava_npp", "HIGHLY significant, robust to outlier removal - suppression vanishes under severe drought",
  "D. Facilitation vs. density-release probe", "fava_npp:Inoc(:Water) on berry_alloc", "Not significant - doesn't distinguish the two stories"
)
knitr::kable(summary_tbl)
```

Part C is the standout result of this exploratory step - worth a place in the discussion
even though it wasn't one of the three pre-specified confirmatory tests. Parts A, B, and D
are honest nulls, reported as such per the plan's own instruction to distinguish
confirmatory from exploratory rather than let interesting exploratory cells get
overemphasized.
