---
title: "Step 8.5 — Research Direction"
subtitle: "AMF × Drought × Neighborhood in Hard Red Spring Wheat: what we asked, what the data show"
author: "Celeste Basken"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.align = "center")
```

```{r libs}
library(tidyverse)
library(car)
library(glmmTMB)
library(nlme)
library(emmeans)

wheat <- readRDS("wheat_audited.rds") %>%
  mutate(log_NPP = log(NPP), log_ANPP = log(ANPP), log_BNPP = log(BNPP),
         log_berry = log(berry), log_veg = log(veg))

wheat_c <- wheat
contrasts(wheat_c$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))

COL_INOC <- "#2c7fb8"
COL_STRL <- "#d95f02"

# --- helpers: pull p-values robustly across model families ---
p_of <- function(m, term, type = 2, ...) {
  a <- car::Anova(m, type = type, ...)
  pcol <- grep("^Pr\\(", colnames(a), value = TRUE)[1]
  if (!term %in% rownames(a)) return(NA_real_)
  a[term, pcol]
}
fmt_p <- function(p) {
  if (is.na(p)) return("NA")
  if (p < 0.0001) return("< 0.0001")
  formatC(p, format = "f", digits = 4)
}
get_cl <- function(d) {
  lo <- intersect(c("asymp.LCL", "lower.CL"), names(d))[1]
  hi <- intersect(c("asymp.UCL", "upper.CL"), names(d))[1]
  c(d[[lo]], d[[hi]])
}
```

```{r key_numbers}
# Primary beta regression on berry allocation
m1 <- glmmTMB(berry_alloc ~ Poly * Inoc * Water, dispformula = ~ Poly,
              family = beta_family(), data = wheat_c)
p_3way_beta <- p_of(m1, "Poly:Inoc:Water")
p_2way_beta <- p_of(m1, "Inoc:Water")

# Unadjusted vs size-adjusted allometric models
m_base   <- lm(berry_alloc ~ Poly * Inoc * Water, data = wheat_c)
m_allo2  <- lm(log_berry ~ log_veg + Poly * Inoc * Water, data = wheat_c)
p_3way_unadj <- p_of(m_base, "Poly:Inoc:Water")
p_3way_adj   <- p_of(m_allo2, "Poly:Inoc:Water")
p_2way_unadj <- p_of(m_base, "Inoc:Water")
p_2way_adj   <- p_of(m_allo2, "Inoc:Water")

# Confirmatory #3: productivity
lev_npp <- car::leveneTest(log_NPP ~ Poly, data = wheat_c)[1, "Pr(>F)"]
m_npp <- if (lev_npp >= 0.05) lm(log_NPP ~ Poly * Inoc * Water, data = wheat_c) else
  gls(log_NPP ~ Poly * Inoc * Water, weights = varIdent(form = ~ 1 | Poly), data = wheat_c)
p_npp_iw <- p_of(m_npp, "Inoc:Water")

# Difference-of-differences at Severe, response scale
emm_link <- emmeans(m1, ~ Inoc | Poly * Water)
dd <- contrast(regrid(emm_link), interaction = c("revpairwise", "revpairwise"), by = "Water")
dd_ci <- as.data.frame(confint(dd))
sev <- dd_ci %>% filter(Water == "Severe")
sev_est <- sev$estimate[1]; sev_cl <- get_cl(sev)

# Fava suppression
inter_c <- wheat %>% filter(Poly == "Inter")
contrasts(inter_c$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))
m_supp <- lm(log(fava_npp) ~ Inoc * Water, data = inter_c)
p_supp_iw <- p_of(m_supp, "Inoc:Water")

# Colonization dose-response (inoculated pots only)
col <- wheat %>% filter(Inoc == "Inoc")
m_col <- lm(berry_alloc ~ hyphal * Water * Poly, data = col)
p_col_hw <- p_of(m_col, "hyphal:Water")

# Mechanism: arbuscule:vesicle
inoc <- wheat %>% filter(Inoc == "Inoc")
contrasts(inoc$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))
m_av <- lm(AV ~ Poly * Water, data = inoc)
p_av_pw <- p_of(m_av, "Poly:Water")
p_av_w  <- p_of(m_av, "Water")
m_hyph_q <- glm(cbind(hyph_n, 100 - hyph_n) ~ Poly * Water,
                family = quasibinomial, data = inoc)
p_hyph_w <- p_of(m_hyph_q, "Water", test = "F")
```

---

# Part 0 — The paper in brief

## Central question

**Does the plant neighborhood determine whether arbuscular mycorrhizal symbiosis is a benefit or a cost to wheat under drought?**

Mycorrhizal outcomes are famously context-dependent, ranging from strongly beneficial to net
costly along a mutualism–parasitism continuum. Nearly all of that literature varies *abiotic*
context — water, phosphorus, light. This experiment asks whether the *biotic* neighborhood does
the same work: whether having a legume neighbor changes where wheat falls on that continuum when
water becomes limiting.

The experiment is a fully crossed greenhouse factorial: stand composition (wheat monoculture vs.
wheat + fava) × AMF inoculation (live inoculum vs. none) × water (80%, 60%, 40% of field
capacity), n = 5 pots per cell, 60 pots, harvested at Zadoks 92.

## Sub-questions and what the data show

**Q1 — Under drought, does AMF change how wheat divides biomass between growth and grain?**

Yes, and sharply. Reproductive allocation collapses in inoculated monoculture wheat under severe
drought while the other three treatment combinations hold roughly steady. The three-way
Poly × Inoc × Water interaction on berry allocation is significant in the primary beta regression
(*p* = `r fmt_p(p_3way_beta)`). The effect is *confined* to one cell — monoculture under severe
drought — which is not a weakness but exactly the localization a three-way interaction predicts.

**Q2 — Is that a genuine allocation shift, or an artifact of AMF making plants bigger?**

A genuine allocation shift. Adjusting for plant size does not weaken the result; it *strengthens*
it (three-way interaction *p* = `r fmt_p(p_3way_unadj)` unadjusted → `r fmt_p(p_3way_adj)`
size-adjusted). The reason is that inoculated plants under severe drought were the *larger* ones,
and larger plants normally produce more grain — so size was masking the penalty rather than
creating it. This was the pre-registered gate for the whole project, and it cleared in the
favorable direction.

**Q3 — Does total productivity change too, or only its division?**

Only its division, and this is the cleanest statement the paper can make. The Inoc × Water
interaction on total wheat NPP is null (*p* = `r fmt_p(p_npp_iw)`). Drought did not change how
inoculation affected the *amount* of biomass wheat produced — it changed *where that biomass
went*. Total growth is preserved; grain is not.

**Q4 — Does a legume neighbor change the outcome?**

Yes. The penalty is essentially absent in mixed stands. The pattern has the shape the stress
gradient hypothesis predicts: monoculture and mixed stands are indistinguishable under control and
moderate water and diverge only under severe drought. The novel twist is that the divergence
appears **only in inoculated plants** — the mutualism itself generates the cost that the neighbor
then offsets. At severe drought the difference-of-differences (Mono − Inter, of the Inoc − Sterile
gap) is `r round(sev_est, 3)` allocation points, 95% CI [`r round(sev_cl[1], 3)`,
`r round(sev_cl[2], 3)`].

**Q5 — What does the symbiosis cost the neighbor?**

Inoculation suppressed fava biomass in mixed stands, and that suppression *vanished* under severe
drought (Inoc × Water on fava biomass, *p* = `r fmt_p(p_supp_iw)`). This is mutualist-mediated
competition: the fungal partnership tilts the competitive hierarchy toward wheat when water is
adequate, and stops doing so when it is not. Exploratory, but well powered relative to the other
exploratory analyses and the strongest non-confirmatory result in the dataset.

**Q6 — What is the mechanism?**

Unresolved, and reported as such. The pre-specified structural test — whether the fungus shifted
from exchange (arbuscules) toward storage (vesicles) under drought — returned no support
(Poly × Water on the arbuscule:vesicle ratio, *p* = `r fmt_p(p_av_pw)`). That null is informative
rather than empty: it rules out the most obvious structural explanation and constrains what the
mechanism can be. See Figure 5 for what remains on the table.

## The synthesis

Reading Q3, Q4 and Q5 together gives a coherent account:

> **AMF's competitive benefit to wheat is conditional on water; its cost is not.** Under adequate
> water the partnership pays — visible as fava suppression in mixed stands. Under severe drought
> the benefit disappears (fava suppression vanishes) while the carbon cost persists. Total growth
> is unaffected either way; what changes is where the biomass is placed. In monoculture, with no
> legume neighbor to offset the cost, wheat pays it out of grain. With fava present, it does not.

This is a statement about **the biotic neighborhood setting position on the mutualism–parasitism
continuum**, which is the paper's contribution. The agronomic implication — that inoculation
decisions made at planting carry yield risk under drought in monoculture, and that legume
intercropping may hedge it — follows from this but should be framed as a hypothesis for field
testing, not a recommendation.

## What the paper claims, and what it does not

| Claims | Does **not** claim |
|---|---|
| The AMF effect on wheat reproductive allocation under severe drought depends on stand composition | That fava specifically, via N fixation, rescues wheat yield |
| The effect is an allocation shift, not a size artifact | That the mechanism is known |
| Total productivity is unaffected; only allocation shifts | That this predicts field yields |
| AMF suppresses fava except under severe drought (exploratory) | That AMF alone, isolated from co-introduced microbiota, causes these effects |

---

# Part 1 — The figures, explained

## Figure 1. The central result: reproductive allocation

```{r fig1, fig.width=8.5, fig.height=5.4}
emm <- as.data.frame(emmeans(m1, ~ Inoc | Poly * Water, type = "response"))

ann <- tibble(
  Poly = factor("Mono", levels = levels(emm$Poly)),
  label = paste0("Diff-of-diffs (Mono - Inter) at Severe:\n",
                 round(sev_est, 3), " pts, 95% CI [",
                 round(sev_cl[1], 3), ", ", round(sev_cl[2], 3), "]")
)

ggplot(emm, aes(x = Water, y = response, color = Inoc, group = Inoc)) +
  geom_line(position = position_dodge(0.3), linewidth = 0.8) +
  geom_point(position = position_dodge(0.3), size = 3) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.15,
                position = position_dodge(0.3)) +
  facet_wrap(~Poly) +
  scale_color_manual(values = c(Inoc = COL_INOC, Strl = COL_STRL)) +
  scale_y_continuous(limits = c(0, 0.45)) +
  geom_text(data = ann, aes(x = 2, y = 0.43, label = label), inherit.aes = FALSE,
            size = 3.2, hjust = 0.5, vjust = 1, color = "grey25") +
  labs(y = "Berry allocation (proportion of total NPP)", x = NULL,
       title = "Reproductive allocation by water and inoculation, within each stand type",
       subtitle = "Estimated marginal means from the primary beta regression; bars = 95% CI") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

**What you are looking at.** The proportion of each wheat plant's total biomass that ended up as
grain, across the three water treatments, split by whether the pot was inoculated (blue) or not
(orange). The left panel is mixed stands; the right is monoculture.

**What it shows.** Three of the four treatment trajectories are unremarkable — allocation drifts
modestly downward as water declines. The fourth, inoculated monoculture, falls off a cliff at
severe drought. Everything else in this document exists to establish that this drop is real, that
it is not a size artifact, and that it is specific to the absence of a legume neighbor.

**The test.** Beta regression with a logit link on the proportion, dispersion allowed to vary by
stand composition, water partitioned into two planned orthogonal contrasts (severe vs. the mean of
control and moderate; control vs. moderate) rather than left as an unstructured 3-level factor.
The designated test is the Poly × Inoc × Water interaction, *p* = `r fmt_p(p_3way_beta)`; the
Inoc × Water interaction is *p* = `r fmt_p(p_2way_beta)`.

**Why the annotated interval, not letters.** A compact letter display across all 12 cells means 66
pairwise comparisons and near-zero power, and it would not actually state the claim. The
difference-of-differences does: it is the amount by which the inoculation penalty is *larger in
monoculture than in mixed stands*, under severe drought. That single interval is the result.

**Caveats.** Beta regression is appropriate here because the response is a proportion bounded by 0
and 1 and because allocation variance is not constant across the range; no values were at the
exact bounds, so no boundary transform was needed. The three allocation fractions (above-ground,
below-ground, grain) necessarily sum to one, so they are not independent — grain allocation is
treated as the single primary response and the other two as descriptive, with a Dirichlet
regression in the Step 8 supplement as a compositional robustness check.

## Figure 2. The gate: is this just plant size?

```{r fig2, fig.width=8.5, fig.height=5.5}
ggplot(wheat, aes(x = log_veg, y = log_berry, color = Inoc)) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, linewidth = 0.7) +
  facet_grid(Poly ~ Water) +
  scale_color_manual(values = c(Inoc = COL_INOC, Strl = COL_STRL)) +
  labs(title = "Grain output against plant size (log-log)",
       subtitle = "Bottom-right panel carries the paper's central contrast",
       x = "log(vegetative mass per plant)", y = "log(berry mass per plant)") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

**Why this figure has to exist.** Allocation measured as a ratio is confounded with plant size,
because plants of different sizes allocate differently regardless of strategy — bigger plants are
further along the same developmental trajectory, not necessarily following a different one. Both
treatments here changed plant size, so "AMF reduced grain fraction" could in principle mean nothing
more than "AMF made plants bigger, and bigger plants have different fractions." The fix is to stop
comparing ratios and instead ask whether treatments shift the relationship between grain output and
vegetative size.

**What it shows.** In the well-watered mixed-stand panels the relationship tilts upward: bigger
plant, more grain, normal biology. In the bottom-right panel — monoculture under severe drought —
the inoculated points (blue) sit noticeably to the *right* of the sterile points, meaning larger
plants, while sitting well *below* them, meaning less grain. Bigger bodies, less wheat. That is the
opposite of the allometric expectation, which is why size adjustment strengthens rather than
weakens the result (`r fmt_p(p_3way_unadj)` → `r fmt_p(p_3way_adj)` for the three-way;
`r fmt_p(p_2way_unadj)` → `r fmt_p(p_2way_adj)` for Inoc × Water).

**The important caveat.** In that bottom-right panel the two groups occupy **non-overlapping ranges
on the x-axis** — sterile plants cluster small, inoculated plants cluster large, with no size at
which both are represented. A size-adjusted contrast there is therefore an *extrapolation*: the
estimated vertical gap depends on the assumed common slope, and the data cannot pin that slope down
within the cell. The direction is stable across every plausible slope, and under the positive slope
actually observed in the well-watered panels the effect gets larger — so the estimate is
conservative. But the manuscript should report a sensitivity range rather than quoting a single
fold-difference as though it were precisely identified.

**What rescues it.** In that cell the two groups appear completely separated on grain mass — every
inoculated pot below every sterile pot. Complete separation means the finding survives without any
model at all, and an exact distribution-free test confirms it (see Step 8 supplementary table,
rows tagged Step 4b). An assumption-free test agreeing with the model is worth more here than a
smaller *p*-value.

**Secondary observation worth reporting.** The fitted slopes are positive under adequate water and
flat-to-negative in all three severe-drought panels. Severe drought appears to break the normal
scaling of reproductive output with plant size. That is a publishable observation on its own, and
it makes the main finding coherent: once extra growth stops converting into grain, a fungal partner
that pushes carbon toward vegetative tissue becomes pure cost. Treat it as suggestive — with n = 5
and compressed x-ranges, individual panel slopes are unreliable, so any claim should rest on the
pooled interaction term, not on the six fitted lines.

## Figure 3. Allocation moves, productivity does not

```{r fig3, fig.width=8, fig.height=7.5}
prod_long <- wheat %>%
  select(Poly, Inoc, Water, NPP, berry, pot_productivity) %>%
  pivot_longer(c(NPP, berry, pot_productivity), names_to = "metric", values_to = "value") %>%
  mutate(metric = dplyr::recode(metric,
           NPP = "Wheat NPP per plant (g)",
           berry = "Berry mass per plant (g)",
           pot_productivity = "Pot-level total productivity (g/plant)"),
         metric = factor(metric, levels = c("Wheat NPP per plant (g)",
                                            "Pot-level total productivity (g/plant)",
                                            "Berry mass per plant (g)")))

summ <- prod_long %>%
  group_by(metric, Poly, Inoc, Water) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            se = sd(value, na.rm = TRUE) / sqrt(sum(!is.na(value))), .groups = "drop")

ggplot(summ, aes(x = Water, y = mean, color = Inoc, group = Inoc)) +
  geom_line(position = position_dodge(0.3)) +
  geom_point(position = position_dodge(0.3), size = 2.5) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.15,
                position = position_dodge(0.3)) +
  facet_grid(metric ~ Poly, scales = "free_y") +
  scale_color_manual(values = c(Inoc = COL_INOC, Strl = COL_STRL)) +
  labs(y = NULL, x = NULL,
       title = "Total productivity vs. grain: only one of them responds",
       subtitle = paste0("Inoc x Water on wheat NPP: p = ", fmt_p(p_npp_iw),
                         " (null) - contrast the top row against the bottom")) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom", legend.title = element_blank(),
        strip.text.y = element_text(size = 8.5))
```

**What it shows.** The top row is total wheat biomass per plant; the bottom row is grain per plant.
Inoculated and sterile lines track each other closely in the top row — including in monoculture
under severe drought, where the allocation effect is largest — while diverging sharply in the
bottom row. The middle row is pot-level total productivity, the closest available system-level
metric.

**Why this matters more than it looks.** This is the null that makes the paper clean. If AMF had
also reduced total productivity under drought, the story would be the familiar and less interesting
"mycorrhizal benefit disappears under stress." Instead, drought did not change how inoculation
affected the *amount* of biomass produced (*p* = `r fmt_p(p_npp_iw)`) — it changed only where the
biomass was placed. The plants grew just as much; they simply did not turn it into grain. That
distinction is the paper's actual contribution and it should lead the abstract.

**Caveat on the middle row.** Pot-level productivity cannot be compared cleanly across stand types,
because the two stand types differ in planting density (see Part 2). It is reported for
completeness and for the within-stand comparison, not as a system-level yield claim. A land
equivalent ratio, which would be the proper agronomic metric, cannot be computed because the design
includes no fava monoculture.

## Figure 4. What the symbiosis costs the neighbor

```{r fig4, fig.width=8, fig.height=4.8}
inter <- wheat %>% filter(Poly == "Inter")

ggplot(inter, aes(x = Water, y = fava_npp, fill = Inoc)) +
  geom_boxplot(position = position_dodge(0.8), width = 0.6, outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8),
             size = 1.4, alpha = 0.85) +
  scale_fill_manual(values = c(Inoc = COL_INOC, Strl = COL_STRL)) +
  labs(x = NULL, y = "Fava biomass (g)",
       title = "Fava biomass by inoculation status, across water treatments",
       subtitle = paste0("All 30 mixed-stand pots. Inoc x Water: p = ", fmt_p(p_supp_iw),
                         " - suppression disappears under severe drought. EXPLORATORY.")) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

**What it shows.** Where water is adequate, fava grown alongside inoculated wheat is markedly
smaller than fava grown alongside sterile wheat. Under severe drought that gap closes.

**Why it belongs in the paper.** This is the other side of the same coin, and it converts a
one-directional study into a bidirectional one. Figure 1 asks what the neighbor does to the
symbiosis; this asks what the symbiosis does to the neighbor. The answer — that a shared mycorrhizal
partner restructures the competitive hierarchy in favor of wheat, and stops doing so under severe
drought — is a recognizably ecological result and it slots directly into the mutualist-mediated
competition literature.

**How it completes the synthesis.** The disappearance of fava suppression under severe drought is
independent evidence that AMF stops delivering a competitive benefit to wheat at that water level.
Combined with Figure 1, where wheat pays a grain cost precisely there, the reading is that the
benefit is conditional on water while the cost is not.

**Caveats.** Exploratory — not among the pre-specified confirmatory tests, and this analysis uses
all 30 mixed-stand pots rather than a randomized fava manipulation. Fava roots were never stained,
so fava's own colonization status is unmeasured and no claim is made about it; the analysis relates
fava *performance* to wheat's inoculation treatment, which was randomized. Note also that the
earlier version of this analysis — regressing fava biomass on continuous wheat hyphal colonization —
was weak and has been dropped, because colonization is bimodal and not randomized. This version
uses the randomized factor instead and is both better powered and better identified.

## Figure 5. The mechanism test that did not pan out

```{r fig5, fig.width=8, fig.height=4.4}
av_summ <- inoc %>%
  group_by(Poly, Water) %>%
  summarise(mean = mean(AV, na.rm = TRUE),
            se = sd(AV, na.rm = TRUE) / sqrt(sum(!is.na(AV))), .groups = "drop")

ggplot(av_summ, aes(x = Water, y = mean, color = Poly, group = Poly)) +
  geom_line(position = position_dodge(0.2), linewidth = 0.8) +
  geom_point(position = position_dodge(0.2), size = 3) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.12,
                position = position_dodge(0.2)) +
  geom_point(data = inoc, aes(x = Water, y = AV, color = Poly), inherit.aes = FALSE,
             position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.2),
             size = 1.3, alpha = 0.35) +
  labs(x = NULL, y = "log arbuscule : vesicle ratio",
       title = "Functional composition of colonization (inoculated pots only)",
       subtitle = paste0("Poly x Water: p = ", fmt_p(p_av_pw),
                         "  |  Water main effect: p = ", fmt_p(p_av_w),
                         "  -  no support for the predicted shift")) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

**The hypothesis being tested.** Arbuscules are where plant and fungus actually exchange resources;
vesicles are fungal lipid storage. The arbuscule:vesicle ratio is therefore an index of whether the
association is *trading* or *hoarding*. The prediction was that under severe drought in monoculture
the ratio would fall — the fungus shifting from exchange to storage — supplying a direct mechanism
for the lost grain.

**What happened.** No support. The Poly × Water interaction is *p* = `r fmt_p(p_av_pw)` and the
water main effect is *p* = `r fmt_p(p_av_w)`. Overall hyphal colonization intensity is likewise flat
across water treatments (*p* = `r fmt_p(p_hyph_w)`), reproducing the earlier finding that
colonization persists under drought — but the hoped-for shift in its functional *composition* is not
there.

**Why this is worth showing rather than burying.** A pre-specified mechanistic test that returns a
clean null constrains the explanation space, and reporting it is what separates a credible paper
from a fishing expedition. What it rules out: the penalty is not explained by a visible structural
reorganization of the symbiosis. Colonization stayed intact in both amount and kind while the
plant's reproductive output collapsed.

**What remains on the table.** Three candidate mechanisms survive:

1. **Phenological delay.** If inoculation delayed flowering or senescence, plants harvested on a
   single common date would have had their grain filling truncated. This is now the leading
   candidate and it is consistent with larger vegetative bodies and less grain. Per-pot Zadoks-92
   timing was not recorded, so this cannot be tested in the existing data — it should be named as
   the priority for follow-up.
2. **Whole-plant carbon drain not visible in structure ratios.** Carbon flux to the fungus need not
   change the arbuscule:vesicle ratio; a constant-composition symbiosis can still be expensive.
   Testing this needs isotopic labelling or tissue chemistry.
3. **Altered water relations.** AMF-mediated changes in stomatal behavior could sustain vegetative
   tissue longer under stress, delaying or displacing reproductive investment.

**Framing consequence.** Without a demonstrated mechanism this is a pattern paper with an unusually
well-defended central pattern, not a mechanistic paper. That is publishable in the target journals,
but the discussion must present mechanism as hypothesis, and the abstract must not imply one.

## Figure 6. Dose–response within inoculated pots

```{r fig6, fig.width=8, fig.height=4.6}
ggplot(col, aes(x = hyphal, y = berry_alloc, color = Water)) +
  geom_point(size = 2.2, alpha = 0.85) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, linewidth = 0.7) +
  facet_wrap(~Poly) +
  labs(x = "Hyphal colonization (proportion)", y = "Berry allocation",
       title = "Does more colonization mean a bigger penalty?",
       subtitle = paste0("Inoculated pots only. One designated test - hyphal x Water: p = ",
                         fmt_p(p_col_hw), ". EXPLORATORY, association only.")) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

**What this asks.** If AMF causes the allocation penalty, more colonization should mean a larger
penalty. A dose–response gradient would strengthen the causal reading considerably.

**Why it is restricted to inoculated pots.** Across the full dataset, colonization is bimodal —
values pile near zero in sterile pots and near one in inoculated pots — so a regression on the full
range would be an inoculation contrast wearing a continuous-predictor costume, with the fit driven
by leverage rather than by any gradient. Restricting to inoculated pots gives a genuinely continuous
predictor.

**How to read it, carefully.** One interaction test is reported (*p* = `r fmt_p(p_col_hw)`), not
three separate per-panel *p*-values. Comparing "significant in this panel, not in that one" is the
same inferential error as comparing separate subset analyses, and it is what an earlier version of
this figure did.

**The unavoidable caveat.** Colonization is a plant *response*, not a randomized treatment. Larger
or healthier root systems can support more hyphae, so reverse causation is live and this figure can
only ever support association language. The randomized inoculation factor carries the causal claim;
this figure is supporting texture.

---

# Part 2 — Statistical approach and standing caveats

## Design and inference

Fully crossed 2 × 2 × 3 factorial, n = 5, balanced with no pot losses. The pot is the experimental
unit throughout — all measurements are per-plant means within a pot, never individual plants, so
there is no pseudoreplication at the plant level. Pots were re-randomized across benches monthly and
reconnected to irrigation, and each irrigation line was present on every bench, so water treatment
is not confounded with bench position.

## Model families

Proportions are modelled by beta regression with a logit link; masses are modelled on the log scale;
colonization counts out of 100 grid intersects are modelled as quasibinomial or beta-binomial with
overdispersion checked. Where Levene's test indicates unequal variance across stand types, the model
allows separate variances (`varIdent` in GLS, `dispformula` in glmmTMB) rather than pooling.

## Water as two planned contrasts

Water is not treated as a linear covariate, because several mass variables show a non-monotonic
pattern with moderate drought exceeding the control — most plausibly mild hypoxia in the wettest
treatment at 80% of field capacity. Instead its 2 df are partitioned into two orthogonal planned
contrasts: severe vs. the mean of control and moderate, and control vs. moderate. This recovers
focused power without assuming monotonicity and keeps the severe-drought effect cleanly separated
from any artifact affecting only the wetter treatments.

## Confirmatory vs. exploratory

Four tests were designated confirmatory in advance; roughly 150 tests exist across the full
analysis. The confirmatory set is the size-adjusted three-way interaction on grain output, the
Inoc × Water interaction on berry allocation in the primary beta regression, and the Inoc × Water
interaction on total NPP. Everything else in this document — Figures 4, 5 and 6 in their entirety —
is exploratory and labelled as such in every subtitle. The complete test record, generated from code
rather than transcribed, is in the Step 8 supplement and `supplementary_test_table.csv`.

## The density confound

Monoculture pots held 12 wheat plants; mixed pots held 6 wheat plus 1 fava, which works out to
roughly 60% of monoculture wheat-equivalent density. Stand composition therefore changes two things
at once — density and neighbor identity — and **the main effect of stand composition is not
interpretable as an intercropping effect**. Higher per-plant biomass in mixed stands is the expected
consequence of halving intraspecific competition.

The interaction terms are far more defensible. For density alone to explain the three-way
interaction, density would have to modulate the AMF × drought interaction specifically — a much more
specific claim, and one that is itself interesting. It cannot be excluded with this design, because
there is no 6-wheat-without-fava control, and no further pots can be run. The manuscript should
name the factor "stand composition," define it by plant counts in the abstract, state the
wheat-equivalent density figure, and never write "the effect of intercropping" about a main effect.

Note also that grain allocation — the primary response — is a *within-plant ratio* and is therefore
the response least mechanically distorted by density, which is fortunate.

## Other standing limitations

- **The inoculum is not a pure AMF treatment.** Inoculated pots received live inoculum containing
  spores, hyphae, root fragments, carrier medium and co-introduced microbiota; sterile pots received
  nothing. The contrast is live inoculum vs. none. A killed-inoculum-plus-filtrate control would be
  needed to isolate AMF, and was not run.
- **Sterile pots showed background colonization.** This shrinks the contrast and makes the tests
  conservative, not liberal.
- **Fava nodulation was observed but not quantified.** A rhizobial inoculant was applied at planting
  and nodules were seen at harvest; no counts or masses exist. The N-fixation mechanism is therefore
  plausible but undemonstrated.
- **Senescence timing was not recorded per pot.** This now matters more than anticipated, because
  phenological delay is the leading surviving mechanistic hypothesis (Figure 5).
- **The central result rests on ten pots.** The effect is localized to monoculture under severe
  drought: five inoculated and five sterile. This is why the leave-one-out, permutation and
  distribution-free checks in Step 4b carry disproportionate weight, and why the manuscript should
  report them rather than only the parametric result.
- **Greenhouse, pots, one cultivar, one inoculum mixture, one season.** Claims are about mechanism
  and context-dependency, not about field yield prediction.

## Framing consequence

The data support an ecological paper with an agronomic implications section, not an agronomic paper.
An agronomy readership will object to pots, n = 5, a non-standard density, and no field validation;
an ecology readership takes greenhouse mesocosms at face value because the question is
context-dependency rather than yield. Lead with the mutualism–parasitism continuum and the stress
gradient framing, keep the agronomic implications to a short section, and present them as hypotheses
for field testing.

---

```{r confirmatory_summary}
conf <- tibble(
  `Confirmatory test` = c(
    "#1  Poly x Inoc x Water, size-adjusted grain output",
    "#2  Inoc x Water, berry allocation (primary beta regression)",
    "#3  Inoc x Water, total wheat NPP"),
  `p-value` = c(fmt_p(p_3way_adj), fmt_p(p_2way_beta), fmt_p(p_npp_iw)),
  Reading = c("Effect is real and not a size artifact",
              "Inoculation penalty is drought-dependent",
              "NULL - productivity unaffected; only allocation shifts")
)
knitr::kable(conf, caption = "The pre-specified confirmatory tests. Everything else is exploratory.")
```

*Full test record: Step 8 supplement and `supplementary_test_table.csv`. Sensitivity, permutation
and leave-one-out checks: Step 4b addendum.*
