---
title: "Step 8 - Figures and Supplementary Test Table"
subtitle: "AMF x Drought x Neighborhood in Hard Red Spring Wheat"
author: "CLAUDE + Celeste Basken"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.align = "center")
```

Step 8 from `Instructions/AMF_intercropping_analysis_plan.md` (Sections 8 and 9). Built only
now that Steps 2-6 hold up, per the plan's own instruction not to build figures/text before
the gates clear. **Step 7 (grain counts from stored berries) was skipped per your call** -
nothing here depends on it.

Every model behind every figure and table row is refit fresh in this document from
`wheat_audited.rds` (not transcribed by hand, not copy-pasted from earlier steps' printed
output) - per the plan's own instruction that the supplementary table "is generated from
code rather than transcribed by hand." The numbers below were spot-checked against each
step's own document during development and match.

```{r}
library(tidyverse)
library(car)
library(glmmTMB)
library(nlme)
library(emmeans)
library(DirichletReg)
library(patchwork)

wheat <- readRDS("wheat_audited.rds") %>%
  mutate(log_NPP = log(NPP), log_ANPP = log(ANPP), log_BNPP = log(BNPP),
         log_berry = log(berry), log_veg = log(veg))
wheat_c <- wheat
contrasts(wheat_c$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))

COL_INOC <- "#2c7fb8"
COL_STRL <- "#d95f02"
```

# Part 1: Figures

## Figure 1. Design schematic

Given the density confound (Section 1 of the plan), showing the design plainly is a
credibility move rather than an optional extra.

```{r fig1, fig.width=8.5, fig.height=7}
pot_plants <- bind_rows(
  tibble(system = "Monoculture (24 planted -> 12 thinned)",
         x = rep(0:3, 3)[1:12], y = rep(0:2, each = 4)[1:12], type = "Wheat"),
  tibble(system = "Intercrop (12 wheat -> 6 thinned;\n3 fava -> 1 thinned)",
         x = c(rep(0:2, 2)[1:6], 3.5), y = c(rep(0:1, each = 3)[1:6], 0.5),
         type = c(rep("Wheat", 6), "Fava"))
) %>% mutate(system = factor(system, levels = unique(system)))

p1 <- ggplot(pot_plants, aes(x = x, y = y, color = type)) +
  geom_point(size = 4) +
  facet_wrap(~system, ncol = 2) +
  scale_color_manual(values = c(Wheat = COL_INOC, Fava = "#238b45")) +
  coord_fixed(clip = "off", xlim = c(-0.5, 4), ylim = c(-0.5, 2.5)) +
  labs(title = "A. Pot composition by stand type", color = NULL) +
  theme_void(base_size = 12) +
  theme(strip.text = element_text(size = 10.5, face = "bold"),
        strip.background = element_rect(fill = "grey92", color = NA),
        panel.background = element_rect(fill = "grey98", color = "grey80"),
        panel.spacing = unit(1.2, "lines"), legend.position = "bottom",
        plot.title = element_text(hjust = 0, face = "plain", size = 12))

timeline <- tibble(stage = c("Planting\n(Z0)", "Drought onset\n(Z39, floret dev.)",
                              "Harvest\n(Z92, ~90% pots)"), z = c(0, 39, 92))
water_labels <- tibble(label = c("Control: 80% FC", "Moderate: 60% FC", "Severe: 40% FC"),
                        y = c(-0.55, -0.75, -0.95))

p2 <- ggplot() +
  annotate("rect", xmin = 39, xmax = 92, ymin = -1.1, ymax = -0.4, fill = COL_STRL, alpha = 0.12) +
  geom_segment(aes(x = 0, xend = 92, y = 0, yend = 0), color = "grey40", linewidth = 1) +
  geom_point(data = timeline, aes(x = z, y = 0), size = 3, color = "grey20") +
  geom_text(data = timeline, aes(x = z, y = 0.15, label = stage), size = 3.2, vjust = 0) +
  geom_text(data = water_labels, aes(x = 65, y = y, label = label), size = 3.1, color = "grey30", hjust = 0) +
  labs(title = "B. Drought timing (Zadoks growth stage) - shaded = drought period, all 3 Water levels applied") +
  xlim(-5, 100) + ylim(-1.2, 0.5) +
  theme_void(base_size = 12) +
  theme(plot.title = element_text(hjust = 0, size = 11))

p1 / p2 + plot_layout(heights = c(2, 1))
```

*Illustrative schematic, not a data plot - built to the counts/timing stated in the plan and
methods. Every pot was re-randomized across benches monthly and reconnected, so Water is not
confounded with bench (the usual fatal flaw in greenhouse drought studies) - worth one
sentence in the methods, per the plan.*

## Figure 2. The key figure - berry allocation, Water x Inoc, faceted by stand composition

```{r fig2, fig.width=8.5, fig.height=5.4}
m1 <- glmmTMB(berry_alloc ~ Poly * Inoc * Water, dispformula = ~ Poly, family = beta_family(), data = wheat_c)
emm <- as.data.frame(emmeans(m1, ~ Inoc | Poly * Water, type = "response"))

emm_link <- emmeans(m1, ~ Inoc | Poly * Water)
dd_resp <- contrast(regrid(emm_link), interaction = c("revpairwise", "revpairwise"), by = "Water")
dd_ci <- as.data.frame(confint(dd_resp))
sev_dd <- dd_ci %>% filter(Water == "Severe")

p_key <- ggplot(emm, aes(x = Water, y = response, color = Inoc, group = Inoc)) +
  geom_line(position = position_dodge(0.3), linewidth = 0.8) +
  geom_point(position = position_dodge(0.3), size = 3) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.15, position = position_dodge(0.3)) +
  facet_wrap(~Poly) +
  scale_color_manual(values = c(Inoc = COL_INOC, Strl = COL_STRL)) +
  scale_y_continuous(limits = c(0, 0.45), expand = expansion(mult = c(0.02, 0.02))) +
  labs(y = "Berry allocation (proportion of total NPP)", x = NULL,
       title = "Berry allocation by Water x Inoculation, faceted by stand composition",
       subtitle = "Points = estimated marginal means (primary beta regression), bars = 95% CI") +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom", legend.title = element_blank())

ann <- tibble(Poly = factor("Mono", levels = levels(emm$Poly)),
              label = paste0("Diff-of-diffs (Mono - Inter), Severe:\n", round(sev_dd$estimate, 3),
                              " pts, 95% CI [", round(sev_dd$asymp.LCL, 3), ", ", round(sev_dd$asymp.UCL, 3), "]"))
p_key + geom_text(data = ann, aes(x = 2, y = 0.42, label = label), inherit.aes = FALSE,
                   size = 3.3, hjust = 0.5, vjust = 1, color = "grey20")
```

*This is the primary result (Confirmatory tests #1-2, Steps 2-3). The annotated
difference-of-differences is the single interval the plan designates as "the result."*

## Figure 3. Allometry - does size explain the allocation shift?

```{r fig3, fig.width=8.5, fig.height=5.5}
m_allo2 <- lm(log_berry ~ log_veg + Poly * Inoc * Water, data = wheat_c)

ggplot(wheat, aes(x = log_veg, y = log_berry, color = Inoc)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, linewidth = 0.7) +
  facet_grid(Poly ~ Water) +
  scale_color_manual(values = c(Inoc = COL_INOC, Strl = COL_STRL)) +
  labs(title = "Allometric relationship: log(berry) vs. log(vegetative mass per plant)",
       subtitle = "Mono x Severe: no overlap in x-range between groups - a size-adjusted contrast there is an extrapolation, \n not an interpolation (see Step 4b sensitivity analysis)",
       x = "log(vegetative mass per plant)", y = "log(berry mass per plant)") +
  theme_minimal(base_size = 12) + theme(legend.position = "bottom", legend.title = element_blank())
```

*Makes Section 2 (the allometry gate) visible. The Mono x Severe panel is the one the whole
paper's contrast leans on - see the addendum follow-up (Step 4b) for the extrapolation
caveat and sensitivity range before quoting a single fold-difference from it.*

## Figure 4. Productivity - does overall plant size shift the way allocation does?

```{r fig4, fig.width=8, fig.height=7.5}
prod_long <- wheat %>%
  select(Poly, Inoc, Water, NPP, berry, pot_productivity) %>%
  pivot_longer(c(NPP, berry, pot_productivity), names_to = "metric", values_to = "value") %>%
  mutate(metric = dplyr::recode(metric, NPP = "Wheat NPP per plant (g)", berry = "Berry mass per plant (g)",
                          pot_productivity = "Pot-level total productivity (g/plant)"),
         metric = factor(metric, levels = c("Wheat NPP per plant (g)", "Pot-level total productivity (g/plant)",
                                             "Berry mass per plant (g)")))

summ <- prod_long %>% group_by(metric, Poly, Inoc, Water) %>%
  summarise(mean = mean(value, na.rm = TRUE), se = sd(value, na.rm = TRUE) / sqrt(sum(!is.na(value))), .groups = "drop")

ggplot(summ, aes(x = Water, y = mean, color = Inoc, group = Inoc)) +
  geom_line(position = position_dodge(0.3)) +
  geom_point(position = position_dodge(0.3), size = 2.5) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.15, position = position_dodge(0.3)) +
  facet_grid(metric ~ Poly, scales = "free_y") +
  scale_color_manual(values = c(Inoc = COL_INOC, Strl = COL_STRL)) +
  labs(y = NULL, x = NULL, title = "Productivity metrics by Water x Inoculation x Stand composition",
       subtitle = "Confirmatory test #3: Inoc:Water on NPP is NOT significant (p=0.625, top row) - contrast against berry mass \n (bottom row)") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom", legend.title = element_blank(), strip.text.y = element_text(size = 8.5))
```

*Visual counterpart to the headline finding of Step 3: NPP tracks closely between Inoc/Strl
even in Mono x Severe (top row), while berry mass alone diverges sharply (bottom row) -
allocation shifts, overall productivity mostly doesn't.*

## Figure 5. Colonization associations (exploratory, Inoc pots only)

```{r fig5, fig.width=8, fig.height=4.8}
col <- wheat %>% filter(Inoc == "Inoc")
m_col <- lm(berry_alloc ~ hyphal * Water * Poly, data = col)
col_hw_p <- car::Anova(m_col, type = 2)["hyphal:Water", "Pr(>F)"]

ggplot(col, aes(x = hyphal, y = berry_alloc, color = Water)) +
  geom_point(size = 2.2, alpha = 0.8) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, linewidth = 0.7) +
  facet_wrap(~Poly) +
  labs(x = "Hyphal colonization (proportion)", y = "Berry allocation",
       title = "Colonization association with berry allocation (Inoc pots only, exploratory)",
       subtitle = paste0("hyphal:Water interaction test (the ONE designated test - not per-panel p-values): p = ",
                          round(col_hw_p, 3))) +
  theme_minimal(base_size = 12) + theme(legend.position = "bottom")
```

*Association language only - `hyphal` was not randomized. One interaction test reported, not
three separate panel p-values (that comparison error is exactly what the plan flagged with
the previous version of this figure).*

## Figure 6. Fava suppression (exploratory)

```{r fig6, fig.width=8, fig.height=8.5}
fava_inoc <- wheat %>% filter(Poly == "Inter", Inoc == "Inoc")
m_f <- lm(log(fava_npp) ~ hyphal * Water, data = fava_inoc)
f_hw_p <- car::Anova(m_f, type = 2)["hyphal:Water", "Pr(>F)"]

p6a <- ggplot(fava_inoc, aes(x = hyphal, y = fava_npp, color = Water)) +
  geom_point(size = 2.5) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, linewidth = 0.7) +
  labs(x = "Wheat hyphal colonization", y = "Fava biomass (g)",
       title = "A. Fava biomass vs. wheat hyphal colonization (Inoc-only, n=15)",
       subtitle = paste0("hyphal:Water: p = ", round(f_hw_p, 3), " - exploratory, no clear association")) +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom")

inter <- wheat %>% filter(Poly == "Inter")
inter_c <- inter
contrasts(inter_c$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))
m_supp <- lm(log(fava_npp) ~ Inoc * Water, data = inter_c)
supp_iw_p <- car::Anova(m_supp, type = 2)["Inoc:Water", "Pr(>F)"]

p6b <- ggplot(inter, aes(x = Water, y = fava_npp, fill = Inoc)) +
  geom_boxplot(position = position_dodge(0.8), width = 0.6, outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.8), size = 1.3) +
  scale_fill_manual(values = c(Inoc = COL_INOC, Strl = COL_STRL)) +
  labs(x = NULL, y = "Fava biomass (g)",
       title = "B. Fava suppression by Inoculation status (all 30 Inter pots)",
       subtitle = paste0("Inoc:Water: p ", ifelse(supp_iw_p < 0.0001, "< 0.0001", paste("=", round(supp_iw_p,4))),
                          " - suppression vanishes under severe drought")) +
  theme_minimal(base_size = 11) + theme(legend.position = "bottom", legend.title = element_blank())

p6a / p6b
```

*Panel A is the plan's literally-requested figure (weak/null). Panel B is the stronger,
better-powered exploratory finding from Step 6 Part C - both clearly labeled exploratory.*

---

# Part 2: Supplementary test table

Per Section 8: "the existing significance tables represent roughly 150 tests... report the
full test table in a supplement." Every row below comes from a model fit fresh in this
document - none are hand-transcribed. **Only 4 rows are the pre-specified confirmatory
tests** (flagged `CONFIRMATORY #1/#2/#3`); everything else is exploratory or a robustness
check on one of those three, labeled as such in the `status` column, not just in prose.

Effect sizes with CIs for the confirmatory results are reported in their own step documents
(Step 2's size-adjusted contrasts, Step 3's emmeans/difference-of-differences) rather than
duplicated as a column here - a single partial-eta-squared-style column isn't
well-defined or comparable across the mix of model families below (OLS, GLS, beta
regression, quasibinomial, beta-binomial, Dirichlet regression), so this table's job is the
exhaustive p-value/test-statistic record specifically, not a second effect-size compendium.

```{r build_table}
tidy_test <- function(model, type = 2, test = NULL) {
  a <- if (!is.null(test)) car::Anova(model, type = type, test = test) else car::Anova(model, type = type)
  broom::tidy(a) %>% filter(term != "Residuals")
}

rows <- list()

# ---- Step 2 ----
m_base <- lm(berry_alloc ~ Poly * Inoc * Water, data = wheat_c)
rows$s2_base <- tidy_test(m_base) %>% mutate(step = "2", status = "Diagnostic", model = "berry_alloc ~ Poly*Inoc*Water (unadjusted baseline)")

rows$s2_allo2 <- tidy_test(m_allo2) %>% mutate(step = "2", status = "CONFIRMATORY #1", model = "log(berry) ~ log(veg)+Poly*Inoc*Water (size-adjusted elevation)")

wheat_c$PolyWater <- interaction(wheat_c$Poly, wheat_c$Water)
m_allo2_gls <- gls(log_berry ~ log_veg + Poly * Inoc * Water, weights = varIdent(form = ~1 | PolyWater), method = "ML", data = wheat_c)
rows$s2_gls <- tidy_test(m_allo2_gls) %>% mutate(step = "2", status = "CONFIRMATORY #1 (robustness)", model = "Same, heteroscedasticity-robust GLS")

# ---- Step 3 ----
rows$s3_m1 <- tidy_test(m1) %>% mutate(step = "3", status = "CONFIRMATORY #2", model = "berry_alloc ~ Poly*Inoc*Water (primary beta regression)")

secondary_specs <- list(
  list(v = "log_NPP", lbl = "log(NPP per plant)"), list(v = "log_ANPP", lbl = "log(ANPP per plant)"),
  list(v = "log_BNPP", lbl = "log(BNPP per plant)"), list(v = "log_berry", lbl = "log(berry mass per plant)"),
  list(v = "W_pcent_ANPP", lbl = "ANPP % of wheat biomass"), list(v = "W_pcent_BNPP", lbl = "BNPP % of wheat biomass"),
  list(v = "pot_productivity", lbl = "Pot-level total productivity")
)
for (s in secondary_specs) {
  lev_p <- car::leveneTest(as.formula(paste(s$v, "~ Poly")), data = wheat_c)[1, "Pr(>F)"]
  m <- if (lev_p >= 0.05) lm(as.formula(paste(s$v, "~ Poly*Inoc*Water")), data = wheat_c) else
    gls(as.formula(paste(s$v, "~ Poly*Inoc*Water")), weights = varIdent(form = ~1 | Poly), data = wheat_c)
  status <- if (s$v == "log_NPP") "CONFIRMATORY #3" else "Exploratory"
  rows[[paste0("s3_", s$v)]] <- tidy_test(m) %>% mutate(step = "3", status = status, model = paste0(s$lbl, " ~ Poly*Inoc*Water"))
}

wheat_dr <- wheat_c %>% filter(!is.na(W_pcent_ANPP))
wheat_dr$Y <- DR_data(wheat_dr[, c("W_pcent_ANPP", "W_pcent_BNPP", "W_pcent_berries")])
m_dir_full <- DirichReg(Y ~ Poly * Inoc * Water | Poly * Inoc * Water | Poly * Inoc * Water, data = wheat_dr)
m_dir_red  <- DirichReg(Y ~ Poly * Inoc + Inoc * Water + Poly * Water | Poly * Inoc + Inoc * Water + Poly * Water | Poly * Inoc + Inoc * Water + Poly * Water, data = wheat_dr)
dir_lrt <- anova(m_dir_red, m_dir_full)
rows$s3_dir <- tibble(term = "Poly:Inoc:Water (joint, all 3 parts)", statistic = dir_lrt$Difference[2], df = dir_lrt$df[2],
                       p.value = dir_lrt[["Pr(>Chi)"]][2], step = "3", status = "Exploratory (robustness)",
                       model = "Dirichlet regression, compositional joint test")

# ---- Step 4b (addendum) ----
cell <- wheat %>% filter(Poly == "Mono", Water == "Severe")
wt <- coin::wilcox_test(berry ~ factor(Inoc), data = cell, distribution = "exact")
rows$s4b_wilcox <- tibble(term = "Inoc (Mono x Severe cell only)", statistic = as.numeric(coin::statistic(wt)), df = NA_real_,
                           p.value = as.numeric(coin::pvalue(wt)), step = "4b", status = "Robustness (distribution-free)",
                           model = "Exact Wilcoxon, berry ~ Inoc, Mono x Severe cell")

perm_null <- readRDS("loo_perm_null_dist.rds")
obs_chisq <- car::Anova(m1, type = 2)["Poly:Inoc:Water", "Chisq"]
rows$s4b_perm <- tibble(term = "Poly:Inoc:Water (permutation)", statistic = obs_chisq, df = NA_real_,
                         p.value = mean(perm_null >= obs_chisq), step = "4b", status = "Robustness (distribution-free)",
                         model = "10,000-draw permutation test, labels shuffled within Water")

wt2 <- t.test(berry_alloc ~ Inoc, data = cell)
rows$s4b_welch <- tibble(term = "Inoc (Mono x Severe cell only)", statistic = unname(wt2$statistic), df = unname(wt2$parameter),
                          p.value = wt2$p.value, step = "4b", status = "Robustness (unequal variance)",
                          model = "Welch t-test, berry_alloc ~ Inoc, Mono x Severe cell")

m_slope_simple <- lm(log_berry ~ log_veg * Water, data = wheat)
rows$s4c_slope <- tidy_test(m_slope_simple) %>% mutate(step = "4c", status = "Exploratory (suggestive)",
                                                          model = "log(berry) ~ log(veg)*Water (pooled, best-powered)")

# ---- Step 5 ----
inoc <- wheat %>% filter(Inoc == "Inoc")
contrasts(inoc$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))
inoc$PolyWater <- interaction(inoc$Poly, inoc$Water)

m_av <- lm(AV ~ Poly * Water, data = inoc)
rows$s5_av <- tidy_test(m_av) %>% mutate(step = "5", status = "Exploratory", model = "AV ~ Poly*Water (OLS)")

m_av_gls <- gls(AV ~ Poly * Water, weights = varIdent(form = ~1 | PolyWater), method = "ML", data = inoc)
rows$s5_av_gls <- tidy_test(m_av_gls) %>% mutate(step = "5", status = "Exploratory (robustness)", model = "AV ~ Poly*Water (heteroscedasticity-robust GLS)")

structures <- list(list(resp = "cbind(arb_n, 100-arb_n)", lbl = "arbuscules (arb_n)"),
                    list(resp = "cbind(ves_n, 100-ves_n)", lbl = "vesicles (ves_n)"),
                    list(resp = "cbind(hyph_n, 100-hyph_n)", lbl = "overall hyphal presence (hyph_n)"))
for (s in structures) {
  form <- as.formula(paste(s$resp, "~ Poly*Water"))
  m_q <- glm(form, family = quasibinomial, data = inoc)
  rows[[paste0("s5_", s$lbl, "_q")]] <- tidy_test(m_q, test = "F") %>% mutate(step = "5", status = "Exploratory", model = paste0(s$lbl, " ~ Poly*Water (quasibinomial)"))
  m_bb <- glmmTMB(form, family = betabinomial, data = inoc)
  rows[[paste0("s5_", s$lbl, "_bb")]] <- tidy_test(m_bb) %>% mutate(step = "5", status = "Exploratory (robustness)", model = paste0(s$lbl, " ~ Poly*Water (beta-binomial)"))
}

m_link <- glmmTMB(berry_alloc ~ AV * Water * Poly, family = beta_family(), data = inoc)
rows$s5_link <- tidy_test(m_link) %>% mutate(step = "5", status = "Exploratory", model = "berry_alloc ~ AV*Water*Poly")

# ---- Step 6 ----
rows$s6_col <- tidy_test(m_col) %>% mutate(step = "6", status = "Exploratory", model = "berry_alloc ~ hyphal*Water*Poly (OLS)")
m_col_beta <- glmmTMB(berry_alloc ~ hyphal * Water * Poly, family = beta_family(), data = col)
rows$s6_col_beta <- tidy_test(m_col_beta) %>% mutate(step = "6", status = "Exploratory (robustness)", model = "Same, beta regression")

rows$s6_f <- tidy_test(m_f) %>% mutate(step = "6", status = "Exploratory", model = "log(fava_npp) ~ hyphal*Water")
m_f2 <- lm(fava_pct_pot ~ hyphal * Water, data = fava_inoc)
rows$s6_f2 <- tidy_test(m_f2) %>% mutate(step = "6", status = "Exploratory", model = "fava_pct_pot ~ hyphal*Water")

rows$s6_supp <- tidy_test(m_supp) %>% mutate(step = "6", status = "Exploratory (standout finding)", model = "log(fava_npp) ~ Inoc*Water")

m_mech <- glmmTMB(berry_alloc ~ fava_npp * Inoc * Water, family = beta_family(), data = inter_c)
rows$s6_mech <- tidy_test(m_mech) %>% mutate(step = "6", status = "Exploratory", model = "berry_alloc ~ fava_npp*Inoc*Water")

final_table <- bind_rows(rows) %>%
  select(step, status, model, term, statistic, df, p.value) %>%
  mutate(statistic = round(statistic, 3), p.value = round(p.value, 5)) %>%
  mutate(status = case_when(
    status == "CONFIRMATORY #1" & term == "Poly:Inoc:Water" ~ "CONFIRMATORY #1",
    status == "CONFIRMATORY #1" ~ "Diagnostic (same model as Confirmatory #1)",
    status == "CONFIRMATORY #1 (robustness)" & term == "Poly:Inoc:Water" ~ "CONFIRMATORY #1 (robustness)",
    status == "CONFIRMATORY #1 (robustness)" ~ "Diagnostic (same model as Confirmatory #1 robustness)",
    status == "CONFIRMATORY #2" & term == "Inoc:Water" ~ "CONFIRMATORY #2",
    status == "CONFIRMATORY #2" ~ "Diagnostic (same model as Confirmatory #2)",
    status == "CONFIRMATORY #3" & term == "Inoc:Water" ~ "CONFIRMATORY #3",
    status == "CONFIRMATORY #3" ~ "Diagnostic (same model as Confirmatory #3)",
    TRUE ~ status
  ))

write_csv(final_table, "supplementary_test_table.csv")
```

```{r}
cat("Total rows:", nrow(final_table), " (plan's own estimate: 'roughly 150 tests')\n")
cat("Confirmatory rows (the ONLY 4 that matter for the paper's central claims):\n")
print(final_table %>% filter(grepl("^CONFIRMATORY", status)) %>% select(step, model, term, statistic, df, p.value))
```

```{r}
knitr::kable(final_table, caption = "Full supplementary test table - every model, every term, every step. Also saved as supplementary_test_table.csv.")
```

Saved to `supplementary_test_table.csv` alongside this document for direct inclusion as a
manuscript supplementary file.
