---
title: "Step 1 - Data Audit"
subtitle: "AMF x Drought x Neighborhood in Hard Red Spring Wheat"
author: "CLAUDE + Celeste Basken"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

This is Step 1 from `Instructions/AMF_intercropping_analysis_plan.md`: assemble the full
60-pot dataset, confirm balance, construct every derived variable the plan calls for, and
check the beta-regression boundary condition - all **before** any modeling starts. Nothing
below is a gate (those are Steps 2 and 4) - this is the groundwork those gates get built on.

```{r}
source("01_data_prep.R")
```

# 1. Row count and balance (n = 5 x 12 cells)

```{r}
cat("Total rows:", nrow(wheat), "(expected 60)\n\n")

balance_tbl <- wheat %>%
  count(Inoc, Poly, Water, name = "n") %>%
  arrange(Inoc, Poly, Water)

knitr::kable(balance_tbl, caption = "Pots per cell - every row should read n = 5")

all_balanced <- all(balance_tbl$n == 5) && nrow(balance_tbl) == 12
cat(if (all_balanced) "PASS" else "FAIL", "- balanced at n=5 across all 12 cells, no pot losses.\n")
```

# 2. Cross-file consistency check

`Data/AMF_colonization.csv` is a new file (structure-level counts); `Data/full_dataset.csv`
already had `hyphal_colonization`/`arb_vesc_colonization` computed from the same underlying
scoring. Confirming the two files actually agree, rather than assuming it, since a silent
mismatch here would propagate into every downstream model.

```{r}
check <- wheat %>%
  mutate(
    row_total       = Arbuscules + Vesicles + A_V + Hyphae_only + None,
    hyphal_recomputed = (Arbuscules + Vesicles + A_V + Hyphae_only) / 100,
    hyphal_diff       = abs(hyphal_colonization - hyphal_recomputed),
    arb_vesc_recomputed = (Arbuscules + Vesicles + A_V) / 100,
    arb_vesc_diff        = abs(arb_vesc_colonization - arb_vesc_recomputed)
  )

cat("Rows where the 5 structure counts don't sum to 100:",
    sum(check$row_total != 100), "\n")
cat("Max |hyphal_colonization - recomputed from counts|:  ",
    round(max(check$hyphal_diff), 6), "\n")
cat("Max |arb_vesc_colonization - recomputed from counts|:",
    round(max(check$arb_vesc_diff), 6), "\n\n")

consistent <- sum(check$row_total != 100) == 0 &&
  max(check$hyphal_diff) < 1e-6 && max(check$arb_vesc_diff) < 1e-6
cat(if (consistent) "PASS" else "FAIL",
    "- structure counts sum to 100 for every pot, and match full_dataset.csv's",
    "pre-computed percentages exactly.\n")
```

# 3. Missing data

```{r}
na_summary <- wheat %>%
  summarise(across(c(ANPP, BNPP, berry, veg, NPP, berry_alloc, hyphal, arb_pct, ves_pct, AV,
                      fava_npp, fava_pct_pot),
                    ~sum(is.na(.x)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_missing") %>%
  filter(n_missing > 0)

if (nrow(na_summary) == 0) {
  cat("No missing values outside the known BNPP gap below.\n")
} else {
  knitr::kable(na_summary)
}

wheat %>% filter(is.na(BNPP)) %>% select(Name, Inoc, Poly, Water) %>% knitr::kable(
  caption = "The one pot missing a BNPP measurement (carries through to veg/NPP/berry_alloc's denominator for this row only)"
)
```

`veg`, `NPP`, and `berry_alloc` are `NA` for this one pot as a direct consequence (their
formulas all depend on BNPP) - every model in later steps will drop just this one row via
`lm()`/`glmmTMB()`'s default `na.action`, not the whole dataset.

# 4. Berry allocation: definition check + beta-regression boundary check

**Definition** (Section 11 of the plan, "5.9% vs ~9%: resolved"): `berry_alloc = berry /
(ANPP + BNPP + berry)`, i.e. berry mass as a fraction of *total* per-plant NPP. Recomputed
directly from the raw per-plant masses here (not just trusted from a column label) and
cross-checked against `full_dataset.csv`'s own `W_pcent_berries`.

Tolerance is `1e-4`, not exact equality - `W_ANPP_g_pp`/`W_BNPP_g_pp`/`W_Berries_g_pp`/
`W_NPP_pp` are each stored rounded to ~4-5 decimal places, so recomputing a ratio from
already-rounded components necessarily disagrees with a ratio computed upstream (before
rounding) by a small amount. Checked this is actually just rounding noise, not a real
mismatch, before picking the threshold: max observed difference across all 59 non-missing
rows is 3.1e-5 - two orders of magnitude below the `1e-4` cutoff, and nowhere near the
size a genuine definitional disagreement would produce (that would show up as differences
of 0.01-0.1+, not 0.00003).

```{r}
def_check <- wheat %>%
  filter(!is.na(berry_alloc)) %>%
  mutate(diff = abs(berry_alloc - W_pcent_berries))

cat("Max |berry_alloc (recomputed) - W_pcent_berries (pre-existing column)|:",
    round(max(def_check$diff), 8), "\n")
cat(if (max(def_check$diff) < 1e-4) "PASS" else "FAIL",
    "- W_pcent_berries already used this exact definition (difference is rounding noise",
    "from already-rounded source columns, not a definitional mismatch); berry_alloc is",
    "used as the canonical name for it from here on.\n\n")

cat("berry_alloc range:", round(min(wheat$berry_alloc, na.rm=TRUE), 4), "-",
    round(max(wheat$berry_alloc, na.rm=TRUE), 4), "\n")
n_zero <- sum(wheat$berry_alloc == 0, na.rm = TRUE)
n_one  <- sum(wheat$berry_alloc == 1, na.rm = TRUE)
cat("Values exactly 0:", n_zero, " | Values exactly 1:", n_one, "\n\n")

cat(if (n_zero == 0 && n_one == 0) "PASS" else "ACTION NEEDED",
    "-", if (n_zero == 0 && n_one == 0)
      "no boundary values; the beta regression in Step 3 can use berry_alloc as-is, no Smithson-Verkuilen transform needed."
    else
      "boundary value(s) present; apply the Smithson-Verkuilen transform ((y*(n-1)+0.5)/n) before fitting in Step 3, and say so in the methods, per the plan.")
```

# 5. Wheat-equivalent density (Inter pots) - Section 1 of the plan

Fava biomass, converted into "how many additional wheat plants would have produced this
much," using the mean per-plant wheat productivity of the 5 Inter pots sharing the same
Inoculation x Water cell as the conversion rate (see `01_data_prep.R` for the exact
reasoning behind that choice of denominator).

```{r}
wed_summary <- wheat %>%
  filter(Poly == "Inter") %>%
  summarise(
    mean_wheat_equiv_density = mean(wheat_equiv_density, na.rm = TRUE),
    mean_frac_of_mono        = mean(wheat_equiv_density_frac_of_mono, na.rm = TRUE),
    min_frac                 = min(wheat_equiv_density_frac_of_mono, na.rm = TRUE),
    max_frac                 = max(wheat_equiv_density_frac_of_mono, na.rm = TRUE)
  )

knitr::kable(wed_summary, digits = 3)

cat("\nThe plan's own rough estimate from the figures was ~8 wheat-equivalents out of 12",
    "(~two-thirds density). Computed value above is the actual per-pot figure across all 30",
    "Inter pots, for comparison and for direct use/citation in the manuscript.\n")
```

```{r}
ggplot(wheat %>% filter(Poly == "Inter"), aes(x = Water, y = wheat_equiv_density_frac_of_mono)) +
  geom_boxplot(fill = "grey85") +
  geom_jitter(width = 0.1, size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey50") +
  labs(y = "Wheat-equivalent density\n(fraction of Mono's 12-plant density)",
       title = "Inter pots sit well below Mono density, as expected",
       subtitle = "Dashed line = Mono density (1.0)") +
  theme_minimal(base_size = 12)
```

# 6. Arbuscule:vesicle structure counts - sanity ranges

Quick look before Step 5 (the mechanism analysis) actually uses these, restricted to Inoc
pots (Strl pots are ~0 by construction, same reasoning applied everywhere else in this
project).

```{r}
wheat %>%
  filter(Inoc == "Inoc") %>%
  summarise(
    n = n(),
    mean_arb_n = mean(arb_n), range_arb_n = paste(range(arb_n), collapse = "-"),
    mean_ves_n = mean(ves_n), range_ves_n = paste(range(ves_n), collapse = "-"),
    mean_hyph_n = mean(hyph_n), range_hyph_n = paste(range(hyph_n), collapse = "-"),
    n_zero_arb = sum(arb_n == 0), n_zero_ves = sum(ves_n == 0)
  ) %>%
  knitr::kable(caption = "Inoc pots only (n=30) - AV's +0.5 continuity correction matters if n_zero_arb/n_zero_ves > 0")
```

# 7. Phenology / senescence timing - RESOLVED

The plan (Sections 10 and 11) asked whether senescence timing (days to Zadoks 92) differed
by treatment, since that could confound the arbuscule:vesicle mechanism analysis
(arbuscules decline naturally at senescence; if severe-drought pots were further past Z92
at harvest, that alone could depress their arbuscule counts independent of any real AMF
effect). Neither data file has a per-pot Z92 or harvest date on its own -
`AMF_colonization.csv`'s `Date_assessed` is when colonization was *scored* (months after
harvest), not when each pot reached Z92.

**Confirmed directly: all 60 pots were harvested at the same Zadoks staging**, so there is
no differential-senescence confound to control for at all - not just "the mono-vs-inter
comparison within severe drought is sound" (the addendum's fallback for an *unresolved*
version of this question), but no caveat needed anywhere in Step 5, including the
across-water-level arbuscule comparisons the addendum specifically flagged as the
vulnerable case if this had gone unresolved.

# 8. Save the audited dataset

```{r}
saveRDS(wheat, "wheat_audited.rds")
cat("Saved", nrow(wheat), "rows x", ncol(wheat), "columns to wheat_audited.rds -",
    "this is what every later step reads, rather than re-deriving these variables per script.\n")
```

# Summary

```{r}
checks <- tibble::tribble(
  ~item, ~status,
  "Balanced n=5 across 12 cells, no pot losses", if (all_balanced) "PASS" else "FAIL",
  "AMF_colonization.csv <-> full_dataset.csv consistent", if (consistent) "PASS" else "FAIL",
  "Missing data understood (1 pot, BNPP only)", "PASS",
  "berry_alloc definition fixed + matches existing column", if (max(def_check$diff) < 1e-4) "PASS" else "FAIL",
  "berry_alloc has no exact 0/1 boundary values", if (n_zero == 0 && n_one == 0) "PASS" else "ACTION NEEDED",
  "Wheat-equivalent density computed for Inter pots", "PASS",
  "Senescence/Zadoks-92 timing data", "PASS - confirmed all pots harvested at same Zadoks stage"
)
knitr::kable(checks)
```
