---
title: "Step 4b/4c - Addendum follow-ups"
subtitle: "AMF x Drought x Neighborhood in Hard Red Spring Wheat"
author: "CLAUDE + Celeste Basken"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

From `Instructions/addendum_after_step2.md`, inserted at the position the addendum itself
specifies: after Step 4 (fragility gate passed), before Step 5 (mechanism). Three follow-up
checks on the Step 2/3 allometry and primary-model results, plus one new secondary test.

```{r}
library(tidyverse)
library(car)
library(emmeans)
library(coin)

wheat <- readRDS("wheat_audited.rds") %>% mutate(log_berry = log(berry), log_veg = log(veg))
cell  <- wheat %>% filter(Poly == "Mono", Water == "Severe")
```

# 4b-i. Non-overlapping size ranges in the key cell

```{r}
overlap_check <- cell %>% select(Name, Inoc, log_veg) %>% arrange(Inoc, log_veg)
knitr::kable(overlap_check, digits = 3, caption = "log(veg) by pot, Mono x Severe")

inoc_range <- range(cell$log_veg[cell$Inoc == "Inoc"])
strl_range <- range(cell$log_veg[cell$Inoc == "Strl"])
```

**Correction to the addendum's own framing, worth stating plainly:** it isn't quite complete
non-overlap. Inoc pots span [`r round(inoc_range[1],2)`, `r round(inoc_range[2],2)`]; Strl
pots span [`r round(strl_range[1],2)`, `r round(strl_range[2],2)`] - four of five Strl pots
sit clearly below all five Inoc pots, but one Strl pot (57, log(veg) = 0.74) falls inside the
Inoc range. Close to complete separation, not exact - the underlying concern (a size-adjusted
contrast here leans heavily on the assumed common slope) still applies and is addressed
below regardless.

```{r}
d_berry <- mean(cell$log_berry[cell$Inoc == "Inoc"]) - mean(cell$log_berry[cell$Inoc == "Strl"])
d_veg   <- mean(cell$log_veg[cell$Inoc == "Inoc"])   - mean(cell$log_veg[cell$Inoc == "Strl"])
cat("d_berry (Inoc - Strl):", round(d_berry, 3), "\n")
cat("d_veg   (Inoc - Strl):", round(d_veg, 3), "\n")
```

## Sensitivity range across assumed common slopes

```{r}
sens <- tibble(assumed_slope = seq(-1, 1, by = 0.25)) %>%
  mutate(fold_difference_Strl_over_Inoc = exp(-(d_berry - assumed_slope * d_veg)))
knitr::kable(sens, digits = 2, caption = "Fold-difference (Strl/Inoc) at assumed common slope b")
```

```{r, fig.width=6, fig.height=4}
ggplot(sens, aes(x = assumed_slope, y = fold_difference_Strl_over_Inoc)) +
  geom_line() + geom_point(size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, linetype = "dotted", color = "grey60") +
  annotate("rect", xmin = 0.75, xmax = 0.95, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "steelblue") +
  labs(title = "Sensitivity of the Mono x Severe fold-difference to the assumed common slope",
       subtitle = "Shaded band = slope range actually observed in Inter/Control and Inter/Moderate panels",
       x = "assumed common slope (b)", y = "fold-difference (Strl / Inoc)") +
  theme_minimal(base_size = 12)
```

Under the biologically normal **positive** slope (the one actually observed in the
well-watered panels, ~0.8-0.9 per Section 3's `m_slope` results below), the fold-difference
is **larger** than the raw estimate, not smaller - a strongly negative slope is needed to
shrink it much, and even at slope = -1 the effect stays a substantial ~1.8x.

**Manuscript language, per the addendum:** report this range rather than a single point
estimate, and note the estimate is conservative under the slope observed in well-watered
panels - do not quote a single fold-difference as precisely identified from this
particular cell alone.

# 4b-ii. Distribution-free confirmation

```{r}
sorted_berry <- cell %>% select(Name, Inoc, berry) %>% arrange(berry)
knitr::kable(sorted_berry, digits = 4, caption = "Raw berry mass, sorted - Mono x Severe")

max_inoc <- max(cell$berry[cell$Inoc == "Inoc"])
min_strl <- min(cell$berry[cell$Inoc == "Strl"])
complete_sep <- max_inoc < min_strl
```

**Same correction as above, on the raw scale this time:** not quite complete separation on
`berry` either - max(Inoc) = `r round(max_inoc,3)`, min(Strl) = `r round(min_strl,3)`, so one
Inoc pot (30) sits above one Strl pot (57). Four of five rank cleanly; one pair crosses.

```{r}
wt <- coin::wilcox_test(berry ~ Inoc, data = cell, distribution = "exact")
wt
```

Still a strong, assumption-free result (p = `r round(pvalue(wt),4)`, not the addendum's
~0.008 floor for perfect separation, but nowhere near null either) - an exact test that
doesn't depend on normality, variance homogeneity, or the allometric model at all agrees
with the parametric primary model.

## Permutation test of the three-way interaction itself

Shuffles `Poly` and `Inoc` labels independently within each `Water` level (preserving the
Water marginal structure, matching what's actually held fixed by design), refits the
primary beta-regression model on each shuffle, and asks how extreme the observed
`Poly:Inoc:Water` chi-sq is against that null distribution.

```{r}
perm_null <- readRDS("loo_perm_null_dist.rds")

wheat_c <- wheat
contrasts(wheat_c$Water) <- cbind(severe = c(-0.5, -0.5, 1), mild = c(-1, 1, 0))
m1 <- glmmTMB::glmmTMB(berry_alloc ~ Poly * Inoc * Water, dispformula = ~ Poly,
                        family = glmmTMB::beta_family(), data = wheat_c)
obs_chisq <- car::Anova(m1, type = 2)["Poly:Inoc:Water", "Chisq"]
perm_p <- mean(perm_null >= obs_chisq)

cat("Observed Poly:Inoc:Water chi-sq:", round(obs_chisq, 2), "\n")
cat("Permutation draws:", length(perm_null), "\n")
cat("Permutation p-value:", perm_p, "\n")
```

```{r, fig.width=6, fig.height=4}
ggplot(tibble(chisq = perm_null), aes(x = chisq)) +
  geom_histogram(bins = 40, fill = "grey70") +
  geom_vline(xintercept = obs_chisq, color = "firebrick", linewidth = 1) +
  labs(title = "Permutation null distribution vs. observed Poly:Inoc:Water chi-sq",
       subtitle = paste0("Observed value (red line) - permutation p = ", round(perm_p, 4)),
       x = "chi-sq", y = "count") +
  theme_minimal(base_size = 12)
```

A fully assumption-free permutation test still finds this real - the observed value sits
well into the tail of the shuffled-label null distribution.

# 4b-iii. Variance heterogeneity

```{r}
cat("=== Welch's t-test (unequal variance), berry_alloc, Mono x Severe ===\n")
wt2 <- t.test(berry_alloc ~ Inoc, data = cell)
print(wt2)

var_tbl <- cell %>% group_by(Inoc) %>%
  summarise(sd_berry_alloc = sd(berry_alloc), sd_berry_raw = sd(berry), n = n())
knitr::kable(var_tbl, digits = 4)
```

**A nuance worth flagging, not glossing over:** the addendum describes Inoc pots as "far
more scattered... roughly a sevenfold spread" - true on **raw berry mass**
(`r round(max(cell$berry[cell$Inoc=='Inoc'])/min(cell$berry[cell$Inoc=='Inoc']),1)`x spread
within Inoc vs `r round(max(cell$berry[cell$Inoc=='Strl'])/min(cell$berry[cell$Inoc=='Strl']),1)`x
within Strl), but **reversed on `berry_alloc`** (the SD is
`r round(var_tbl$sd_berry_alloc[var_tbl$Inoc=='Strl']/var_tbl$sd_berry_alloc[var_tbl$Inoc=='Inoc'],1)`x
higher in Strl than Inoc) - dividing by total NPP changes which group looks more variable,
since Inoc plants in this cell are also the larger ones. `berry_alloc` is the paper's primary
response, so this is the more decision-relevant direction, and it's the opposite of the raw-
mass framing. Either way: Welch's test (which doesn't assume equal variance) still finds the
group difference clearly significant (p = `r format.pval(wt2$p.value)`), so the substantive
conclusion is unaffected - just correcting which group is "more scattered" on which scale.

## Does scatter track colonization intensity? (dose-response probe)

```{r}
inoc_cell <- cell %>% filter(Inoc == "Inoc") %>% arrange(hyphal)
knitr::kable(inoc_cell %>% select(Name, hyphal, berry_alloc, berry), digits = 4,
             caption = "The 5 Inoc/Mono/Severe pots, ranked by colonization")

ct <- cor.test(inoc_cell$hyphal, inoc_cell$berry_alloc)
cat("Pearson r =", round(ct$estimate, 3), ", p =", round(ct$p.value, 3),
    "(n = 5 - essentially no power; 95% CI [", round(ct$conf.int[1],2), ",",
    round(ct$conf.int[2],2), "])\n")
```

**No detectable dose-response signal at this sample size**, and what pattern exists doesn't
match the "least-colonized = best performer" story the addendum flagged as the interesting
case: the single least-colonized pot (hyphal = 0.66) is actually the *worst* performer
(berry_alloc = 0.025), and the best performer (0.107) sits at intermediate colonization
(0.89), not the low end. With n=5 this is not strong evidence against a dose-response
relationship either - just genuinely inconclusive. Noted for Step 5, not treated as a
finding either way.

# 4c. New secondary result: does drought decouple size from reproduction?

```{r}
m_slope_full <- lm(log_berry ~ log_veg * Water * Poly * Inoc, data = wheat)
full_tbl <- car::Anova(m_slope_full, type = 2) %>% broom::tidy() %>%
  filter(grepl("log_veg", term))
knitr::kable(full_tbl %>% mutate(across(c(statistic,p.value), ~round(.x,4))),
             caption = "Fully-saturated model - log_veg terms only")
```

```{r}
# Better-powered version, per the addendum's own caution: individual per-panel
# slopes are unreliable at n=5 with compressed x-ranges - the claim should come
# from the interaction term itself, in the best-powered specification available,
# not from eyeballing six separate fitted lines.
m_slope_simple <- lm(log_berry ~ log_veg * Water, data = wheat)
simple_tbl <- car::Anova(m_slope_simple, type = 2) %>% broom::tidy() %>% filter(term != "Residuals")
knitr::kable(simple_tbl %>% mutate(across(c(statistic,p.value), ~round(.x,4))),
             caption = "log(berry) ~ log(veg) * Water, pooled across Poly/Inoc (better powered: 54 residual df vs 36)")

trends <- emmeans::emtrends(m_slope_simple, ~ Water, var = "log_veg")
print(trends)
```

```{r, fig.width=7, fig.height=4}
ggplot(wheat, aes(x = log_veg, y = log_berry, color = Water)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE) +
  labs(title = "Allometric slope by Water level, pooled across Poly/Inoc",
       x = "log(vegetative mass per plant)", y = "log(berry mass per plant)") +
  theme_minimal(base_size = 12)
```

**Reported as suggestive, not confirmed - not oversold just because the pattern is visually
clean.** The slope estimates move exactly the way the addendum's visual read predicted
(Control +0.91, Moderate +0.78, Severe -0.12 - a clear flattening/reversal), and this is the
better-powered of the two specifications tested. But the formal test for `log_veg:Water`
lands at p = `r round(simple_tbl$p.value[simple_tbl$term=='log_veg:Water'],3)` here (and
p = `r round(full_tbl$p.value[full_tbl$term=='log_veg:Water'],3)` in the fully-saturated
model) - marginal, not clean. **This is not being added to the confirmatory set** (Section
8's three tests are unchanged); it's a real, pre-specified-by-the-addendum pattern worth
describing in the discussion as suggestive, with the actual p-value stated rather than
rounded up to "significant."

Proceeding to Step 5 (mechanism - arbuscule:vesicle analysis).
