---
title: "Significance grid - Monoculture only - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-08-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# READ ME FIRST

Same grid as `Significance_Grid_DRAFT.Rmd`, restricted to **Mono (monoculture) pots
only**. Everything about how a cell is built - the `Yes`/`Kinda (<.1)`/`No` label off
BH-adjusted `p_adj`, the direction text, the thumbnail with its p-value/stars baked in -
is identical; only the underlying models change, because of what dropping down to one
polyculture does to the design:

- **Polyculture is no longer a predictor.** Every pot here is Mono, so there's no
  variation left for Polyculture to explain - the 3-way `Inoculation*Polyculture*Water`
  models from the full-dataset grid become 2-way `Inoculation*Water` models. That means
  there's no "Polyculture", "Poly:Inoc", "Poly:Water", or 3-way row in this grid at all -
  they're not blank/`NA`, they simply don't exist as term rows here (`term_labels_for_scope
  ("mono")` only has 3 possible terms: `AMF inoculation`, `Water (PPT)`, `Inoc:Water`).
- **AMF colonization** is fit on Inoc-only Mono pots, with `Water` as the only predictor
  (Polyculture is fixed by this doc's scope; Inoculation is fixed by the Inoc-only
  subsetting - neither has any variation left to test).
- **Fava columns are dropped entirely** - `Fava_g` and `Pot_pcent_fava` are exactly 0 for
  every Mono pot by definition (no fava plant present), so there's zero variance for an
  ANOVA to test. They don't appear as `NA` columns here; they're just not part of this
  scope's `battery_spec` at all.
- **"Total pot productivity" is dropped too**, along with the Fava columns - it's
  numerically identical to "Wheat total NPP per plant" for this scope (max difference
  ~5e-5 g, i.e. rounding), since a Mono pot's total biomass IS its wheat biomass with no
  fava plant to add. Per your call, it's excluded here rather than kept as a redundant
  column.

All of this comes from `build_battery_spec("mono")` in `battery_core.R` - nothing here is
hand-adjusted per variable.

```{r, message=FALSE}
source("battery_core.R")   # combined/inoc_only/inter_only/mono_only, run_anova_check(),
                             # build_battery_spec(), run_battery()
source("grid_helpers.R")    # plot_cell(), direction_label(), build_cell_data(), render_grid()
```

```{r}
# This scope's own spec + fitted battery - NOT the top-level battery_spec/
# term_summary that battery_core.R also builds (those are scope = "full").
battery_spec_mono <- build_battery_spec("mono")
batt_mono          <- run_battery(battery_spec_mono)
```

```{r}
# Own plot directory, separate from the other two grids', so identically-
# named cells (e.g. a "Water" main-effect thumbnail exists in all three
# grids) don't overwrite each other on disk.
plot_dir <- "significance_grid_plots_mono"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. Build the cell data (one row per response variable x ANOVA term)

```{r}
cell_data <- build_cell_data(batt_mono$term_summary, batt_mono$spec_by_label, batt_mono$results_list, plot_dir)
```

# 2. Render the two grids (mass / percent)

```{r}
term_labels <- term_labels_for_scope("mono")
```

## 2.1 Mass variables (grams, per plant)

```{r}
render_grid("mass", cell_data, battery_spec_mono, term_labels)
```

## 2.2 Percent / allocation variables

```{r}
render_grid("percent", cell_data, battery_spec_mono, term_labels)
```
