---
title: "big_takeaways"
output: html_document
---

This doc is just to look at figures that could help me with my big takeaways (or what I think they are so far..?)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}

library(tidyverse)
library(broom)
library(knitr)
library(dplyr)
library(ggplot2)
```

## Load data

```{r}
# Combined data
combined <- read.csv("../Cross_compare/NPP_AMF_combined.csv")

# need to remove the fourth row bc I'm missing data (the bnpp)
combined <- combined[-4,]

combined <- combined %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Group = interaction(Inoculation, Polyculture, sep = "_")
  )

# Speciality subsetted ones
mono_only <- combined[combined$Polyculture != "Inter", ]

inter_only <- combined[combined$Polyculture != "Mono", ]

```

## Part 1: Monoculture story

For NPP, inoculation and water are both significant

```{r}
# NPP

# Check stat significance
an_biomass2_mono_only <- aov(NPP_g_pp~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)

# Water: Inoc interaction
ggplot(mono_only, aes(x = Water, y = NPP_g_pp, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Mono Only: Water:Inoculation Interaction Impacts on system NPP", x = "", y = "NPP")

# Berries

##  Total berries
# Stats
an_berries2_mono_only <- aov(Berries_g_pp~Inoculation*Water, data= mono_only)
summary(an_berries2_mono_only)

# Water: Inoc interaction
ggplot(mono_only, aes(x = Water, y = Berries_g_pp, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Mono Only: Water:Inoculation Interaction Impacts on Berry Productivity", x = "", y = "NPP")

## Percent berries
# Stats
an_berries_percent2_mono_only <- aov(percent_berries_of_wheat~Inoculation*Water,
                                     data= mono_only)
summary(an_berries_percent2_mono_only)

## BERRIES AS PERCENT OF WHEAT (which is the same lmfao since its mono but whateva)
# Water: Inoc interaction
ggplot(mono_only, aes(x = Water, y = percent_berries_of_wheat, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoc Interaction Impacts on Allocation to Berries Within Wheat", x = "", y = "% berries of wheat")

```

```{r}
## BNPP
# Water: Inoc interaction
ggplot(mono_only, aes(x = Water, y = percent_BNPP_of_wheat, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoc Interaction Impacts on Allocation to BNPP Within Wheat", x = "", y = "% BNPP")

## ANPP
# Water: Inoc interaction
ggplot(mono_only, aes(x = Water, y = percent_ANPP_of_wheat, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoc Interaction Impacts on Allocation to ANPP Within Wheat",
       x = "", y = "% ANPP")
```


```{r}
# AMF hyphal_colonization 

# Check stat significance
an_hyphal2_mono_only <- aov(hyphal_colonization~Water, data= mono_only)
summary(an_hyphal2_mono_only)

# Water
ggplot(mono_only, aes(x = Water, y = hyphal_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Hyphal Colonization", x = "", y = "% roots with hyphal structures")

# AMF arb_vesc_colonization 

# Check stat significance
an_arb_vesc2_mono_only <- aov(arb_vesc_colonization~Water, data= mono_only)
summary(an_arb_vesc2_mono_only)

# Water
ggplot(mono_only, aes(x = Water, y = arb_vesc_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Arbuscular Vesicular Colonization", x = "", y = "% roots with Arbuscular or Vesicular AMF structures")

```



## Part 2: Intercropping only, what is fava doing?

```{r}

# NPP including fava
# Check stat significance
an_biomass2_mono_only <- aov(NPP_g_pp~Inoculation*Water, data= inter_only)
summary(an_biomass2_mono_only)

# Water: Inoc interaction
ggplot(inter_only, aes(x = Water, y = NPP_g_pp, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Inter Only: Water:Inoculation Interaction Impacts on system NPP", x = "", y = "NPP")

#FAVA productivity
# Water: Inoc interaction
ggplot(inter_only, aes(x = Water, y = Fava_g, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on fava productivity", x = "", y = "Fava (g)")

#FAVA percent
# Water: Inoc interaction
ggplot(inter_only, aes(x = Water, y = percent_fava, color = Inoculation, 
                       group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on allocation to fava", x = "", y = "Fava %")

```

```{r}
# Wheat NPP combined
# Water: Inoc interaction
ggplot(inter_only, aes(x = Water, y = Wheat_NPP, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on wheat (combined) productivity", x = "", y = "Wheat NPP (g)")

an_wheatNPP2_inter_only <- aov(Wheat_NPP~Inoculation*Water, data= mono_only)
summary(an_wheatNPP2_inter_only)

# Wheat NPP combined %
# Water: Inoc interaction
ggplot(inter_only, aes(x = Water, y = percent_wheat_NPP, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on wheat (combined) productivity", x = "", y = "Wheat NPP (g)")

an_wheatNPP_percent_inter_only <- aov(percent_wheat_NPP~Inoculation*Water, data= mono_only)
summary(an_wheatNPP_percent_inter_only)
```

How is this happening? Which specific parts of wheat are losing out

```{r}
# BERRIES
# Percent
ggplot(inter_only, aes(x = Water, y = percent_berries, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoc Interaction Impacts % berries of system productivity", x = "", y = "Berry %")

an_berries_percent_inter_only <- aov(percent_berries~Inoculation*Water, data= mono_only)
summary(an_berries_percent_inter_only)

# total
  # not interesting
ggplot(inter_only, aes(x = Water, y = Berries_g_pp, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on Berry (g)", x = "", y = "Berry g")

## BERRIES WITHIN WHEAT
# Percent
ggplot(inter_only, aes(x = Water, y = percent_berries_of_wheat, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoc Interaction Impacts on Allocation to Berries Within Wheat", x = "", y = "Berry % of Wheat")

# is this stat significant?
an_berries_percent_inter_only <- aov(percent_berries_of_wheat~Inoculation*Water, data= mono_only)
summary(an_berries_percent_inter_only)


```

```{r}
# ANPP
# Percent
ggplot(inter_only, aes(x = Water, y = percent_ANPP, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on allocation to ANPP", x = "", y = "ANPP %")

an_ANPP_percent_inter_only <- aov(percent_ANPP~Inoculation*Water, data= mono_only)
summary(an_ANPP_percent_inter_only)

# total
ggplot(inter_only, aes(x = Water, y = ANPP_g_pp, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on ANPP (g)", x = "", y = "ANPP g")

# BNPP
# Percent
ggplot(inter_only, aes(x = Water, y = percent_BNPP, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on allocation to BNPP", x = "", y = "BNPP %")

# total
  # not interesting
ggplot(inter_only, aes(x = Water, y = BNPP_g_pp, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on BNPP (g)", x = "", y = "BNPP g")

an_BNPP_percent_inter_only <- aov(percent_BNPP~Inoculation*Water, data= mono_only)
summary(an_BNPP_percent_inter_only)
```


## Part 3. Fava can offset the losses during drought of berry % incurred by inoculation!

```{r}
ggplot(combined, aes(x = Water, 
                     y = percent_berries_of_wheat, 
                     color = Inoculation, 
                     shape = Polyculture,
                     group = interaction(Inoculation, Polyculture))) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  stat_summary(fun = mean, geom = "line", aes(linetype = Polyculture)) +
  labs(
    title = "Fava Intercropping Ameliorates Decline in Berry Allocation
    Incurred by Inoculation During Drought",
    x = "",
    y = "Berry % of Wheat"
  )
```

