---
title: "Direction_of_significance"
author: "Celeste Basken"
date: "2026-03-11"
output: html_document
---

This doc is to look at interactions that were significant in our table of significance and then create a rough figure for em.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}

library(tidyverse)
library(broom)
library(knitr)
library(dplyr)
library(ggplot2)
```

## Load data

```{r}
# Combined data
combined <- read.csv("../Cross_compare/NPP_AMF_combined.csv")

# need to remove the fourth row bc I'm missing data (the bnpp)
combined <- combined[-4,]

combined <- combined %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Group = interaction(Inoculation, Polyculture, sep = "_")
  )
```

# 1. Simple biomass interactions

## ANPP

```{r}
# ANPP x polyculture
ggplot(combined, aes(x = Polyculture, y = ANPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on ANPP (treatment means)", x = "", y = "ANPP per plant (g)")

# ANPP x Inoculation
ggplot(combined, aes(x = Inoculation, y = ANPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on ANPP (treatment means)", x = "", y = "ANPP per plant (g)")

# ANPP x PPT
ggplot(combined, aes(x = Water, y = ANPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on ANPP", x = "", y = "ANPP per plant (g)")

```
## BNPP

```{r}
# polyculture
ggplot(combined, aes(x = Polyculture, y = BNPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on BNPP", x = "", y = "BNPP per plant (g)")

# not stat significant but curious
ggplot(combined, aes(x = Water, y = BNPP_g_pp,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture) +
  labs(title = "BNPP productivity (g) with 3-way interaction Poly: Inoc: Water")

```

## Berry

```{r}
# polyculture
ggplot(combined, aes(x = Polyculture, y = Berries_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on Berries", x = "", y = "Berries g per plant")

# Water
ggplot(combined, aes(x = Water, y = Berries_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Berries", x = "", y = "Berries g per plant")

# Berry 3-way
ggplot(combined, aes(x = Water, y = Berries_g_pp,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture) +
  labs(title = "Berry productivity (g) with 3-way interaction Poly: Inoc: Water")


```

## Fava

```{r}
# Inoculation
ggplot(combined, aes(x = Inoculation, y = Fava_g)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on Fava NPP", x = "", y = "Fava g")

# Water
ggplot(combined, aes(x = Water, y = Fava_g)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Fava NPP", x = "", y = "Fava g")

# Water: Inoc interaction
ggplot(combined, aes(x = Water, y = Fava_g, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Intaction for Fava NPP", x = "", y = "Fava g")

```

## NPP

```{r}
# Polyculture
ggplot(combined, aes(x = Polyculture, y = NPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on Total NPP", x = "", y = "NPP g")

# Water
ggplot(combined, aes(x = Water, y = NPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Total NPP", x = "", y = "NPP g")

# Poly: Inoc interaction
ggplot(combined, aes(x = Inoculation, y = NPP_g_pp, color = Polyculture, group = Polyculture)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Polyculture:Inoculation Intaction for Total NPP", x = "", y = "NPP g")

ggplot(combined, aes(x = Polyculture, y = NPP_g_pp, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Polyculture:Inoculation Intaction for Total NPP", x = "Intercropping", y = "NPP g")

# !! not statistically significant but I'm curious!!
# Water: Inoc interaction
ggplot(combined, aes(x = Water, y = NPP_g_pp, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on system NPP", x = "", y = "NPP")

# not stat significant but 3-way
ggplot(combined, aes(x = Water, y = NPP_g_pp,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture) +
  labs(title = "NPP of system with 3-way interaction Poly: Inoc: Water")

```


## Wheat productivity combined in intercropped treatments

What if we looked at wheat success combined

ANSWER: there is no real impact. Drought is affecting both wheat & fava "equally". This fits bc water has no impact on % fava.

```{r}

# total npp vs water, intercropped only
ggplot(data = filter(combined, Polyculture %in% c("Inter")), aes(x = Water, y = NPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on NPP", x = "", y = "NPP g per plant")

# Wheat NPP vs water
ggplot(data = filter(combined, Polyculture %in% c("Inter")), aes(x = Water, y = Wheat_NPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Wheat NPP", x = "", y = "Wheat NPP g per plant")

# % Wheat NPP vs water
ggplot(data = filter(combined, Polyculture %in% c("Inter")), aes(x = Water, y = percent_wheat_NPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on % Wheat NPP", x = "", y = "Wheat NPP g per plant")


# !! not statistically significant but I'm curious!!
# Water: Inoc interaction
ggplot(combined, aes(x = Water, y = percent_wheat_NPP, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on wheat NPP %", x = "", y = "Wheat percent of pot mass (basically wheat vs fava)")

# this is only within intercropping
an_NPP_wheat_percent <- aov(percent_wheat_NPP~Water_gal*Inoculation, 
                            data = filter(combined, Polyculture %in% c("Inter")))
summary(an_NPP_wheat_percent)
# within intercropped system, inoculation has statistically significant impact on allocation to wheat vs fava


```

## Fava and AMF root colonization

These statistically are maybe sus but I'm exploring okay

```{r}
# AMF colonization impact on fava NPP Inter only
an_fava_amf_hyphal <- aov(Fava_g~hyphal_colonization*Water, data= inter_only)
summary(an_fava_amf_hyphal)

an_fava_amf_arb_vesc <- aov(Fava_g~arb_vesc_colonization*Water, data= inter_only)
summary(an_fava_amf_arb_vesc)


# Fava x hyphal
ggplot(inter_only, aes(x = hyphal_colonization, y = Fava_g)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "AMF hyphal colonization impact on fava NPP",
    x = "AMF hyphal colonization (%)",
    y = "Fava NPP"
  ) 

# Fava x arb vesc
ggplot(inter_only, aes(x = arb_vesc_colonization, y = Fava_g)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "AMF Arbuscular-Vesicular colonization impact on fava NPP",
    x = "AMF arb vesc colonization (%)",
    y = "Fava NPP"
  ) 

ggplot(inter_only, aes(x = hyphal_colonization, y = Fava_g)) +
  geom_point(size = 2.5, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "#2C7BB6") +
  facet_wrap(~Water) +
  labs(
    title = "AMF colonization–fava NPP relationship varies with water availability",
    x = "AMF hyphal colonization (%)",
    y = "Fava NPP"
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )


```

```{r}
# fooling around

# HYPHAL

an_fava_percent_amf_hyph <- aov(percent_fava~hyphal_colonization*Water, data= inter_only)
summary(an_fava_percent_amf_hyph)

ggplot(inter_only, aes(x = hyphal_colonization, y = percent_fava)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "AMF hyphal colonization impact on fava competitiveness",
    x = "AMF hyphal colonization (%)",
    y = "Fava percent of pot"
  ) 

ggplot(inter_only, aes(x = hyphal_colonization, y = percent_fava)) +
  geom_point(size = 2.5, alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE, color = "#2C7BB6") +
  facet_wrap(~Water) +
  labs(
    title = "AMF impact on fava competitiveness",
    x = "AMF hyphal colonization (%)",
    y = "Fava percent of pot biomass"
  ) +
  theme_bw(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

# ARB VESC

an_fava_percent_amf_arb_vesc <- aov(percent_fava~arb_vesc_colonization*Water, data= inter_only)
summary(an_fava_percent_amf_arb_vesc)

ggplot(inter_only, aes(x = arb_vesc_colonization, y = percent_fava)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "AMF arbuscular-vesicular colonization impact on fava competitiveness",
    x = "AMF arb-vesc colonization (%)",
    y = "Fava percent of pot"
  ) 


```



# 2. Stoichiometry

## ANPP %

```{r}
# ANPP x polyculture
ggplot(combined, aes(x = Polyculture, y = percent_ANPP_of_wheat)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on ANPP %", x = "", y = "ANPP% of NPP")

# ANPP x Inoculation
ggplot(combined, aes(x = Inoculation, y = percent_ANPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on ANPP %", x = "", y = "ANPP% of NPP")

# ANPP x PPT
ggplot(combined, aes(x = Water, y = percent_ANPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on ANPP %", x = "", y = "ANPP% of NPP")

# Poly: Inoc: Water 3-way with ANPP
ggplot(combined, aes(x = Water, y = percent_ANPP,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture)

# Poly: Inoc: Water 3-way with ANPP
ggplot(combined, aes(x = Water, y = percent_ANPP,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture) +
  labs(title = "Allocation to ANPP (% of NPP) with 3-way interaction Poly: Inoc: Water")

```
## BNPP %

```{r}
# polyculture
ggplot(combined, aes(x = Polyculture, y = percent_BNPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on BNPP %", x = "", y = "BNPP per plant (g)")

# water
ggplot(combined, aes(x = Water, y = percent_BNPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Allocation to BNPP based on PPT", x = "", y = "BNPP per plant (g)")

# Water: Inoc interaction
ggplot(combined, aes(x = Water, y = percent_BNPP, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on Allocation to BNPP", x = "", y = "BNPP %")

# Poly: Inoc: Water 3-way with berries %
ggplot(combined, aes(x = Water, y = percent_BNPP,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture) +
  labs(title = "Allocation to BNPP (% of NPP) with 3-way interaction Poly: Inoc: Water")

```


## Berry %

```{r}

# Water
ggplot(combined, aes(x = Water, y = percent_berries)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Biomass allocation to Berries", x = "", y = "Berries g per plant")

# Water
ggplot(combined, aes(x = Water, y = percent_berries)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Biomass allocation to Berries", x = "", y = "Berries g per plant")

# Water
ggplot(combined, aes(x = Polyculture, y = percent_berries_of_wheat)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Intercropping impact on Biomass allocation to Berries", x = "", y = "Berries g per plant")

# AMF
ggplot(combined, aes(x = Inoculation, y = percent_berries)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on Biomass allocation to Berries", x = "", y = "Berries g per plant")

# poly inoc
ggplot(combined, aes(x = Polyculture, y = percent_berries, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Impacting of Polyculture and Inoculation on Resource Allocation to berries", x = "Intercropping", y = "% Berry")

    # other way poly inoc
ggplot(combined, aes(x = Inoculation, y = percent_berries, color = Polyculture, group = Polyculture)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Impacting of Polyculture and Inoculation on Resource Allocation to berries", x = "Intercropping", y = "% Berry")

# Water: Inoc interaction
ggplot(combined, aes(x = Water, y = percent_berries, 
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction for Berry %", x = "", y = "Berry %")

# Poly: Inoc: Water 3-way with berries %
ggplot(combined, aes(x = Water, y = percent_berries,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture) +
  labs(title = "Allocation to Berries (% of NPP) with 3-way interaction Poly: Inoc: Water")

```
## Fava %

```{r}
# Inoculation
ggplot(combined, aes(x = Inoculation, y = percent_fava)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on Fava % of Biomass", x = "", y = "Fava % of biomass")

# Water: Inoc interaction
ggplot(combined, aes(x = Water, y = percent_fava, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on Fava % of System", x = "", y = "Fava % of system productivity")

```

## AMF root colonization

```{r}
# Water: Hyphae
ggplot(combined, aes(x = Water, y = hyphal_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Hyphal colonization", x = "", y = "% hyphal colonization")

# poly-water hyphal
ggplot(combined, aes(x = Water, y = hyphal_colonization, 
                     color = Polyculture, group = Polyculture)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Impact of Drought & Polyculture on AMF hyphal colonization", x = "", y = "Hyphal colonization %")

# Water: Arbuscular/ Vesicular
ggplot(combined, aes(x = Water, y = arb_vesc_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Arbuscular/ Vesicular colonization", x = "", y = "% hyphal colonization")

# Water: Arbuscular/ Vesicular
ggplot(combined, aes(x = Polyculture, y = arb_vesc_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on Arbuscular/ Vesicular colonization", x = "", y = "% hyphal colonization")

```


