---
title: "Direction_of_significance"
author: "Celeste Basken"
date: "2026-03-11"
output: html_document
---

This doc is to look at interactions that were significant in our table of significance and then create a rough figure for em.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}

library(tidyverse)
library(broom)
library(knitr)
library(dplyr)
library(ggplot2)
```

## Load data

```{r}
# Combined data
combined <- read.csv("../Cross_compare/NPP_AMF_combined.csv")

# need to remove the fourth row bc I'm missing data (the bnpp)
combined <- combined[-4,]

combined <- combined %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Group = interaction(Inoculation, Polyculture, sep = "_")
  )
```

# 1. Simple biomass interactions

## ANPP

```{r}
# ANPP x polyculture
ggplot(combined, aes(x = Polyculture, y = ANPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on ANPP (treatment means)", x = "", y = "ANPP per plant (g)")

# ANPP x Inoculation
ggplot(combined, aes(x = Inoculation, y = ANPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on ANPP (treatment means)", x = "", y = "ANPP per plant (g)")

# ANPP x PPT
ggplot(combined, aes(x = Water, y = ANPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on ANPP", x = "", y = "ANPP per plant (g)")

```
## BNPP

```{r}
# polyculture
ggplot(combined, aes(x = Polyculture, y = BNPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on BNPP", x = "", y = "BNPP per plant (g)")
```

## Berry

```{r}
# polyculture
ggplot(combined, aes(x = Polyculture, y = Berries_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on Berries", x = "", y = "Berries g per plant")

# Water
ggplot(combined, aes(x = Water, y = Berries_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Berries", x = "", y = "Berries g per plant")


```

## Fava

```{r}
# Inoculation
ggplot(combined, aes(x = Inoculation, y = Fava_g)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on Fava NPP", x = "", y = "Fava g")

# Water
ggplot(combined, aes(x = Water, y = Fava_g)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Fava NPP", x = "", y = "Fava g")

# Water: Inoc interaction
ggplot(combined, aes(x = Water, y = Fava_g, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Intaction for Fava NPP", x = "", y = "Fava g")

```

## NPP

```{r}
# Polyculture
ggplot(combined, aes(x = Polyculture, y = NPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on Total NPP", x = "", y = "NPP g")

# Water
ggplot(combined, aes(x = Water, y = NPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Total NPP", x = "", y = "NPP g")

# Poly: Inoc interaction
ggplot(combined, aes(x = Inoculation, y = NPP_g_pp, color = Polyculture, group = Polyculture)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Polyculture:Inoculation Intaction for Total NPP", x = "", y = "NPP g")

ggplot(combined, aes(x = Polyculture, y = NPP_g_pp, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Polyculture:Inoculation Intaction for Total NPP", x = "Intercropping", y = "NPP g")
```
## Wheat productivity combined in intercropped treatments

What if we looked at wheat success combined

ANSWER: there is no real impact. Drought is affecting both wheat & fava "equally". This fits bc water has no impact on % fava.

```{r}
# new column which is wheat NPP
combined <- combined %>%
  rowwise() %>%
  mutate(Wheat_NPP = sum(c(ANPP_g_pp,BNPP_g_pp,Berries_g_pp))) %>%
  ungroup()

# total npp vs water, intercropped only
ggplot(data = filter(combined, Polyculture %in% c("Inter")), aes(x = Water, y = NPP_g_pp)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on NPP", x = "", y = "NPP g per plant")

# Wheat NPP vs water
ggplot(data = filter(combined, Polyculture %in% c("Inter")), aes(x = Water, y = Wheat_NPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Wheat NPP", x = "", y = "Wheat NPP g per plant")

# wheat as percentage of NPP
combined <- combined %>%
  rowwise() %>%
  mutate(percent_wheat_NPP = (Wheat_NPP/NPP_g_pp)) %>%
  ungroup()

# % Wheat NPP vs water
ggplot(data = filter(combined, Polyculture %in% c("Inter")), aes(x = Water, y = percent_wheat_NPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on % Wheat NPP", x = "", y = "Wheat NPP g per plant")


```


# 2. Stoichiometry

## ANPP %

```{r}
# ANPP x polyculture
ggplot(combined, aes(x = Polyculture, y = percent_ANPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on ANPP %", x = "", y = "ANPP% of NPP")

# ANPP x Inoculation
ggplot(combined, aes(x = Inoculation, y = percent_ANPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on ANPP %", x = "", y = "ANPP% of NPP")

# ANPP x PPT
ggplot(combined, aes(x = Water, y = percent_ANPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on ANPP %", x = "", y = "ANPP% of NPP")

# Poly: Inoc: Water 3-way with ANPP
ggplot(combined, aes(x = Water, y = percent_ANPP,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture)

# Poly: Inoc: Water 3-way with ANPP
ggplot(combined, aes(x = Water, y = percent_ANPP,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture) +
  labs(title = "Allocation to ANPP (% of NPP) with 3-way interaction Poly: Inoc: Water")

```
## BNPP %

```{r}
# polyculture
ggplot(combined, aes(x = Polyculture, y = percent_BNPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on BNPP %", x = "", y = "BNPP per plant (g)")

# water
ggplot(combined, aes(x = Water, y = percent_BNPP)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Allocation to BNPP based on PPT", x = "", y = "BNPP per plant (g)")

# Poly: Inoc: Water 3-way with berries %
ggplot(combined, aes(x = Water, y = percent_BNPP,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture) +
  labs(title = "Allocation to BNPP (% of NPP) with 3-way interaction Poly: Inoc: Water")

```


## Berry %

```{r}

# Water
ggplot(combined, aes(x = Water, y = percent_berries)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Biomass allocation to Berries", x = "", y = "Berries g per plant")

# Water
ggplot(combined, aes(x = Water, y = percent_berries)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Biomass allocation to Berries", x = "", y = "Berries g per plant")

# AMF
ggplot(combined, aes(x = Inoculation, y = percent_berries)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on Biomass allocation to Berries", x = "", y = "Berries g per plant")

# poly inoc
ggplot(combined, aes(x = Polyculture, y = percent_berries, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Impacting of Polyculture and Inoculation on Resource Allocation to berries", x = "Intercropping", y = "% Berry")

    # other way poly inoc
ggplot(combined, aes(x = Inoculation, y = percent_berries, color = Polyculture, group = Polyculture)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Impacting of Polyculture and Inoculation on Resource Allocation to berries", x = "Intercropping", y = "% Berry")

# Water: Inoc interaction
ggplot(combined, aes(x = Water, y = percent_berries, 
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction for Berry %", x = "", y = "Berry %")

# Poly: Inoc: Water 3-way with berries %
ggplot(combined, aes(x = Water, y = percent_berries,
                     color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  facet_wrap(~ Polyculture) +
  labs(title = "Allocation to Berries (% of NPP) with 3-way interaction Poly: Inoc: Water")

```
## Fava %

```{r}
# Inoculation
ggplot(combined, aes(x = Inoculation, y = percent_fava)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Inoculation impact on Fava % of Biomass", x = "", y = "Fava % of biomass")

# Water: Inoc interaction
ggplot(combined, aes(x = Water, y = percent_fava, color = Inoculation, group = Inoculation)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Water:Inoculation Interaction Impacts on Fava % of System", x = "", y = "Fava % of system productivity")

```

## AMF root colonization

```{r}
# Water: Hyphae
ggplot(combined, aes(x = Water, y = hyphal_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Hyphal colonization", x = "", y = "% hyphal colonization")

# poly-water hyphal
ggplot(combined, aes(x = Water, y = hyphal_colonization, 
                     color = Polyculture, group = Polyculture)) +
  stat_summary(fun = mean, geom = "point") +
  stat_summary(fun = mean, geom = "line") +
  labs(title = "Impact of Drought & Polyculture on AMF hyphal colonization", x = "", y = "Hyphal colonization %")

# Water: Arbuscular/ Vesicular
ggplot(combined, aes(x = Water, y = arb_vesc_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Arbuscular/ Vesicular colonization", x = "", y = "% hyphal colonization")

# Water: Arbuscular/ Vesicular
ggplot(combined, aes(x = Polyculture, y = arb_vesc_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Polyculture impact on Arbuscular/ Vesicular colonization", x = "", y = "% hyphal colonization")

```

