---
title: "plabeled_figs"
output: html_document
---

Let's do some figure drawin

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(broom)
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(multcompView)
library(multcomp)
library(emmeans)
library(stringr)
```


## Load data

```{r}
# Combined data
combined <- read.csv("../Cross_compare/NPP_AMF_combined.csv")

# need to remove the fourth row bc I'm missing data (the bnpp)
combined <- combined[-4,]

combined <- combined %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Group = interaction(Inoculation, Polyculture, sep = "_")
  )

# percent vegetative of wheat
combined <- combined %>%
  mutate(
    percent_vegetative_of_wheat = percent_ANPP_of_wheat + percent_BNPP_of_wheat
  )

# Speciality subsetted ones
mono_only <- combined[combined$Polyculture != "Inter", ]

inter_only <- combined[combined$Polyculture != "Mono", ]

```




## FIGURE 1: inoc + ppt impact on NPP and yield

```{r}

mono_long <- mono_only %>%
  dplyr::select(Water_gal, Water, Inoculation, Berries_g_pp, NPP_g_pp) %>%
  pivot_longer(
    cols = c(Berries_g_pp, NPP_g_pp),
    names_to = "Response",
    values_to = "Value"
  ) %>%
  mutate(
    Response = case_when(
      Response == "NPP_g_pp" ~ "(A) Total Biomass",
      Response == "Berries_g_pp" ~ "(B) Yield"
    )
  )

letters_df2 <- mono_long %>%
  group_by(Response, Water) %>%
  group_modify(~ {
    mod <- lm(Value ~ Inoculation, data = .x)
    em <- emmeans::emmeans(mod, ~ Inoculation)
    cld <- multcomp::cld(em, adjust = "tukey", Letters = letters)
    as.data.frame(cld)
  }) %>%
  mutate(.group = stringr::str_trim(.group)) %>%
  dplyr::select(Response, Water, Inoculation, .group)

summary_df2 <- mono_long %>%
  group_by(Response, Water, Inoculation) %>%
  summarise(
    mean = mean(Value, na.rm = TRUE),
    se = se(Value),
    ymax = mean + se,
    .groups = "drop"
  )

label_positions2 <- summary_df2 %>%
  left_join(letters_df2, by = c("Response", "Water", "Inoculation")) %>%
  group_by(Response, Water) %>%
  mutate(
    panel_range = max(ymax, na.rm = TRUE) - min(ymax, na.rm = TRUE),

    # base height above error bars (increase this)
    base_y = ymax + 0.3 * panel_range,

    # stagger labels by group (key fix)
    offset = ifelse(Inoculation == levels(Inoculation)[1],
                    0.04 * panel_range,
                    0.10 * panel_range),

    label_y = base_y + offset
  ) %>%
  ungroup()

pd <- position_dodge(width = 0.7)

inoc_ppt_interact.png <- ggplot(
  mono_long,
  aes(x = Water, y = Value, color = Inoculation, group = Inoculation)
) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2, position = pd) +
  stat_summary(fun = mean, geom = "point", size = 3.8, position = pd) +
  stat_summary(fun.data = mean_se, geom = "errorbar", 
             width = 0.12, linewidth = 0.8, position = pd) +
  
  geom_text(
  data = label_positions2,
  aes(x = Water, y = label_y, label = .group, color = Inoculation),
  position = position_dodge(width = 0.3),  # 👈 key fix
  fontface = "bold",
  size = 5
  ) +
  
  facet_wrap(~Response, scales = "free_y") +
  scale_color_manual(values = c("#3B6C8E", "#C2A46D")) +
  scale_y_continuous(expand = expansion(mult = c(0.12, 0.25))) +
  
  labs(
    title = "Interacting Effects of Precipitation and AMF Inoculation",
    x = "Drought treatment",
    y = "Biomass (g per plant)",
    color = "Inoculation",
    caption = "Points show means ± SE. Different letters indicate Sidak-adjusted pairwise differences\n between inoculation treatments within each drought level (p < 0.05)."
  ) +
  
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    strip.text = element_text(face = "bold", size = 13),
    strip.background = element_rect(fill = "#EAF1F5", color = NA),
    
    panel.grid = element_blank(),                 # remove ALL gridlines
    axis.line = element_line(color = "black", linewidth = 0.8),
    axis.ticks = element_line(color = "black"),
    
    panel.spacing = unit(1.5, "lines"),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    plot.caption = element_text(hjust = 0, size = 10),
    plot.margin = margin(10, 20, 15, 10)
  )

ggsave("PresentationFigs/Plabels/inoc_ppt_interact.png",
       plot = inoc_ppt_interact.png,
       width = 9, height = 4.8, dpi = 300)

```



## FIGURE 2: allocation


```{r}
se <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))

format_p <- function(p) {
  case_when(
    is.na(p) ~ "p = NA",
    p < 0.001 ~ "p < 0.001",
    TRUE ~ paste0("p = ", sprintf("%.3f", p))
  )
}

# Colors
inoc_colors <- c(
  "Inoc" = "#3B6C8E",
  "Strl" = "#C2A46D"
)

# Long dataframe: 3 allocation responses
alloc_long <- mono_only %>%
  dplyr::select(
    Water, Inoculation,
    percent_ANPP_of_wheat,
    percent_BNPP_of_wheat,
    percent_berries_of_wheat
  ) %>%
  pivot_longer(
    cols = c(
      percent_ANPP_of_wheat,
      percent_BNPP_of_wheat,
      percent_berries_of_wheat
    ),
    names_to = "Response",
    values_to = "Value"
  ) %>%
  mutate(
    Response = case_when(
      Response == "percent_ANPP_of_wheat" ~ "(A) Aboveground Biomass",
      Response == "percent_BNPP_of_wheat" ~ "(B) Belowground Biomass",
      Response == "percent_berries_of_wheat" ~ "(C) Yield"
    ),
    Response = factor(Response, levels = 
                        c("(A) Aboveground Biomass", "(B) Belowground Biomass", "(C) Yield")),
    Water = factor(Water, levels = c("Control", "Moderate", "Severe")),
    Inoculation = factor(Inoculation, levels = c("Inoc", "Strl"))
  )

# Summary stats
summary_alloc <- alloc_long %>%
  group_by(Response, Water, Inoculation) %>%
  summarise(
    mean = mean(Value, na.rm = TRUE),
    se = se(Value),
    ymin = mean - se,
    ymax = mean + se,
    .groups = "drop"
  )

# Compact letters across all Water x Inoculation combinations within each response
letters_alloc <- alloc_long %>%
  group_by(Response) %>%
  group_modify(~ {
    mod <- lm(Value ~ Water * Inoculation, data = .x)

    em <- emmeans(mod, ~ Water * Inoculation)

    cld_out <- multcomp::cld(
      em,
      adjust = "sidak",
      Letters = letters
    )

    as.data.frame(cld_out)
  }) %>%
  mutate(.group = str_trim(.group)) %>%
  dplyr::select(Response, Water, Inoculation, .group)

# Label positions for significance letters
label_positions_alloc <- summary_alloc %>%
  left_join(letters_alloc, by = c("Response", "Water", "Inoculation")) %>%
  group_by(Response) %>%
  mutate(
    panel_range = max(ymax, na.rm = TRUE) - min(ymin, na.rm = TRUE),
    label_y = ymax + 0.1 * panel_range
  ) %>%
  ungroup()

# Model stats: interaction p-value + model R²
model_stats_alloc <- alloc_long %>%
  group_by(Response) %>%
  group_modify(~ {
    mod <- lm(Value ~ Water * Inoculation, data = .x)
    anova_out <- anova(mod)

    tibble(
      r2 = summary(mod)$r.squared,
      p_interaction = anova_out["Water:Inoculation", "Pr(>F)"]
    )
  }) %>%
  mutate(
    label = paste0(
      "Water × Inoculation: ",
      format_p(p_interaction),
      "\nR² = ",
      sprintf("%.2f", r2)
    )
  )

# Position model-stat labels in each panel
stats_positions_alloc <- summary_alloc %>%
  group_by(Response) %>%
  summarise(
    x = 1,
    y = max(ymax, na.rm = TRUE) + 
      0.28 * (max(ymax, na.rm = TRUE) - min(ymin, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  left_join(model_stats_alloc, by = "Response")

pd <- position_dodge(width = 0.3)

alloc_ppt_interact_3panel.png <- ggplot(
  summary_alloc,
  aes(x = Water, y = mean, color = Inoculation, group = Inoculation)
) +
  geom_line(linewidth = 1.2, position = pd) +
  geom_point(size = 3.8, position = pd) +
  geom_errorbar(
    aes(ymin = ymin, ymax = ymax),
    width = 0.12,
    linewidth = 0.8,
    position = pd
  ) +
  geom_text(
    data = label_positions_alloc,
    aes(
      x = Water,
      y = label_y,
      label = .group,
      color = Inoculation,
      group = Inoculation
    ),
    position = pd,
    inherit.aes = FALSE,
    fontface = "bold",
    size = 5
  ) +
  geom_text(
    data = stats_positions_alloc,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    size = 3.6,
    color = "black"
  ) +
  facet_wrap(~ Response, scales = "free_y", nrow = 1) +
  scale_color_manual(values = inoc_colors) +
  scale_y_continuous(expand = expansion(mult = c(0.12, 0.32))) +
  labs(
    title = "Impacts of Drought and AMF Inoculation on Resource Allocation",
    x = "Drought treatment",
    y = "Allocation (%)",
    color = "Inoculation",
    caption = "Points show means ± SE. Letters indicate Sidak-adjusted pairwise differences among Water × Inoculation treatment combinations within each panel."
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    strip.text = element_text(face = "bold", size = 13),
    strip.background = element_rect(fill = "#EAF1F5", color = NA),
    panel.grid = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.8),
    axis.ticks = element_line(color = "black"),
    panel.spacing = unit(1.5, "lines"),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    plot.caption = element_text(hjust = 0, size = 10),
    plot.margin = margin(10, 20, 15, 10)
  )

ggsave(
  "PresentationFigs/Plabels/alloc_ppt_interact_3panel.png",
  plot = alloc_ppt_interact_3panel.png,
  width = 16,
  height = 7,
  dpi = 300
)
```


## FIGURE 3: Relationships between AMF hyphal colonization and biomass allocation and NPP under drought.

```{r}
# pretty with p values

p_labels <- hyphal_mono_berries %>%
  group_by(Response, Water) %>%
  group_modify(~{
    mod <- lm(Value ~ hyphal_colonization, data = .x)

    tidy_mod <- broom::tidy(mod)
    glance_mod <- broom::glance(mod)

    intercept <- tidy_mod %>%
      filter(term == "(Intercept)") %>%
      pull(estimate)

    slope <- tidy_mod %>%
      filter(term == "hyphal_colonization") %>%
      pull(estimate)

    p_val <- tidy_mod %>%
      filter(term == "hyphal_colonization") %>%
      pull(p.value)

    r2 <- glance_mod$r.squared

    data.frame(
      equation = paste0(
        "y = ", round(intercept, 3),
        ifelse(slope >= 0, " + ", " - "),
        abs(round(slope, 3)), "x"
      ),
      p_label = paste0(
        "p = ", signif(p_val, 2),
        "\nR² = ", signif(r2, 2)
      ),
      x = -Inf,
      y = -Inf
    )
  })

hyphal_berries_NPP.png <- ggplot(hyphal_mono_berries, aes(x = hyphal_colonization, y = Value)) +
  geom_point(size = 2.8, alpha = 0.8, color = "black") +
  geom_smooth(
    method = "lm",
    se = TRUE,
    color = "#3B6C8E",
    fill = "#3B6C8E",
    alpha = 0.2,
    linewidth = 1.2
  ) +
  geom_text(
  data = p_labels,
  aes(x = x, y = y, label = p_label),
  inherit.aes = FALSE,
  hjust = -0.3,
  vjust = -0.0,
  size = 4.2
) +
  facet_grid(Response ~ Water, scales = "free_y", drop = TRUE) +
  labs(
    x = "Root hyphal colonization (%)",
    y = NULL,
    title = "AMF Colonization Impact on Yield Allocation and Total Biomass"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    strip.text = element_text(face = "bold", size = 13),
    strip.background = element_rect(fill = "#EAF1F5", color = NA),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.spacing = unit(1.2, "lines")
  )

ggsave("PresentationFigs/Plabels/hyphal_berries_NPP.png", plot = hyphal_berries_NPP.png, width = 9, height = 6.5, dpi = 300)

caption_eqs <- p_labels %>%
  dplyr::select(Response, Water, equation) %>%
  mutate(
    caption_text = paste0(Response, " under ", Water, ": ", equation)
  )

caption_eqs$caption_text

```

# SUPPLEMENTALS

## FIGURE S1: drought decreases NPP and yield

```{r}
se <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))

format_p <- function(p) {
  case_when(
    is.na(p) ~ "p = NA",
    p < 0.001 ~ "p < 0.001",
    TRUE ~ paste0("p = ", sprintf("%.3f", p))
  )
}

mono_long <- mono_only %>%
  dplyr::select(Water, Berries_g_pp, NPP_g_pp) %>%
  pivot_longer(
    cols = c(Berries_g_pp, NPP_g_pp),
    names_to = "Response",
    values_to = "Value"
  ) %>%
  mutate(
    Response = case_when(
      Response == "NPP_g_pp" ~ "(A) Total Biomass",
      Response == "Berries_g_pp" ~ "(B) Yield"
    ),
    Response = factor(Response, levels = c("(A) Total Biomass", "(B) Yield")),
    Water = factor(Water, levels = c("Control", "Moderate", "Severe"))
  )

summary_mono <- mono_long %>%
  group_by(Response, Water) %>%
  summarise(
    mean = mean(Value, na.rm = TRUE),
    se = se(Value),
    ymin = mean - se,
    ymax = mean + se,
    .groups = "drop"
  )

letters_mono <- mono_long %>%
  group_by(Response) %>%
  group_modify(~ {
    mod <- lm(Value ~ Water, data = .x)
    em <- emmeans(mod, ~ Water)

    cld <- multcomp::cld(
      em,
      adjust = "tukey",
      Letters = letters
    )

    as.data.frame(cld)
  }) %>%
  mutate(.group = stringr::str_trim(.group)) %>%
  dplyr::select(Response, Water, .group)

label_positions <- summary_mono %>%
  left_join(letters_mono, by = c("Response", "Water")) %>%
  group_by(Response) %>%
  mutate(
    panel_range = max(ymax, na.rm = TRUE) - min(ymin, na.rm = TRUE),
    label_y = ymax + 0.10 * panel_range
  ) %>%
  ungroup()

model_stats <- mono_long %>%
  group_by(Response) %>%
  group_modify(~ {
    mod <- lm(Value ~ Water, data = .x)

    tibble(
      p_value = anova(mod)["Water", "Pr(>F)"],
      r2 = summary(mod)$r.squared
    )
  }) %>%
  mutate(
    label = paste0(
      format_p(p_value),
      "\nR² = ",
      sprintf("%.2f", r2)
    )
  )

stats_positions <- summary_mono %>%
  group_by(Response) %>%
  summarise(
    x = 1,
    y = max(ymax, na.rm = TRUE) +
      0.32 * (max(ymax, na.rm = TRUE) - min(ymin, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  left_join(model_stats, by = "Response")

biomass_water_facet.png <- ggplot(
  summary_mono,
  aes(x = Water, y = mean, group = 1)
) +
  geom_line(
    linewidth = 1.3,
    color = "#3B6C8E"
  ) +
  geom_point(
    size = 4,
    color = "#3B6C8E"
  ) +
  geom_errorbar(
    aes(ymin = ymin, ymax = ymax),
    width = 0.12,
    linewidth = 0.9,
    color = "#3B6C8E"
  ) +
  geom_text(
    data = label_positions,
    aes(x = Water, y = label_y, label = .group),
    inherit.aes = FALSE,
    fontface = "bold",
    size = 5,
    color = "black"
  ) +
  geom_text(
    data = stats_positions,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    size = 4,
    color = "black"
  ) +
  facet_wrap(~ Response, scales = "free_y") +
  scale_y_continuous(expand = expansion(mult = c(0.12, 0.35))) +
  labs(
    title = "Drought Impact on Total Productivity and Yield",
    x = "Drought treatment",
    y = "Biomass (g per plant)",
    caption = "Points show means ± SE. Different letters indicate Sidak-adjusted differences among drought treatments within each panel."
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    strip.text = element_text(face = "bold", size = 13),
    strip.background = element_rect(fill = "#EAF1F5", color = NA),
    panel.grid = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.8),
    axis.ticks = element_line(color = "black"),
    panel.spacing = unit(1.2, "lines"),
    plot.caption = element_text(hjust = 0, size = 10),
    plot.margin = margin(10, 20, 15, 10)
  )

biomass_water_facet.png

ggsave(
  "PresentationFigs/Plabels/biomass_water_facet.png",
  plot = biomass_water_facet.png,
  width = 9,
  height = 4.8,
  dpi = 300
)
```

## FIGURE S2: drought does not impact hyphal colonization

```{r}
# Fit model for p-value and R²
mod <- lm(hyphal_colonization ~ Water, data = mono_only)
mod_summary <- summary(mod)

p_val <- anova(mod)[["Pr(>F)"]][1]
r2_val <- mod_summary$r.squared

p_label <- ifelse(p_val < 0.001, "italic(p) < 0.001",
                  paste0("italic(p) == ", signif(p_val, 2)))
r2_label <- paste0("italic(R)^2 == ", round(r2_val, 2))

label_text <- paste(r2_label, "*','~~", p_label)

# Get label position
label_pos <- mono_only %>%
  summarise(
    x = 1,
    y = max(hyphal_colonization, na.rm = TRUE)
  )

root_hyphal_drought.png <- ggplot(
  mono_only,
  aes(x = Water, y = hyphal_colonization)
) +
  stat_summary(
    fun = mean,
    geom = "bar",
    fill = "#3B6C8E",
    color = "black",
    alpha = 0.85,
    width = 0.7
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    linewidth = 0.9,
    color = "black"
  ) +
  geom_text(
    data = label_pos,
    aes(x = x, y = y, label = label_text),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    parse = TRUE,
    size = 5
  ) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15))) +
  labs(
    title = "Impact of Precipitation on AMF Hyphal Colonization",
    x = "",
    y = "Root hyphal colonization (%)"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),

    panel.grid = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.8),
    axis.ticks = element_line(color = "black"),

    plot.margin = margin(10, 20, 15, 10)
  )

ggsave(
  "PresentationFigs/Plabels/root_hyphal_drought.png",
  plot = root_hyphal_drought.png,
  width = 9,
  height = 4.8,
  dpi = 300
)
```


## FIGURE S3: IS CANCELLED BC TUKEY IS NOT VALID

```{r}
se <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))

format_p <- function(p) {
  case_when(
    is.na(p) ~ "p = NA",
    p < 0.001 ~ "p < 0.001",
    TRUE ~ paste0("p = ", sprintf("%.3f", p))
  )
}

# Long dataframe: 3 allocation responses
alloc_water_long <- mono_only %>%
  dplyr::select(
    Water,
    percent_ANPP_of_wheat,
    percent_BNPP_of_wheat,
    percent_berries_of_wheat
  ) %>%
  pivot_longer(
    cols = c(
      percent_ANPP_of_wheat,
      percent_BNPP_of_wheat,
      percent_berries_of_wheat
    ),
    names_to = "Response",
    values_to = "Value"
  ) %>%
  mutate(
    Response = case_when(
      Response == "percent_ANPP_of_wheat" ~ "(A) Aboveground Biomass",
      Response == "percent_BNPP_of_wheat" ~ "(B) Belowground Biomass",
      Response == "percent_berries_of_wheat" ~ "(C) Yield"
    ),
    Response = factor(
      Response,
      levels = c(
        "(A) Aboveground Biomass",
        "(B) Belowground Biomass",
        "(C) Yield"
      )
    ),
    Water = factor(Water, levels = c("Control", "Moderate", "Severe"))
  )

# Summary stats
summary_alloc_water <- alloc_water_long %>%
  group_by(Response, Water) %>%
  summarise(
    mean = mean(Value, na.rm = TRUE),
    se = se(Value),
    ymin = mean - se,
    ymax = mean + se,
    .groups = "drop"
  )

# Compact letters for Water differences within each response
letters_alloc_water <- alloc_water_long %>%
  group_by(Response) %>%
  group_modify(~ {

    mod <- lm(Value ~ Water, data = .x)

    em <- emmeans(mod, ~ Water)

    cld_out <- multcomp::cld(
      em,
      adjust = "tukey",
      Letters = letters
    )

    as.data.frame(cld_out)

  }) %>%
  mutate(.group = stringr::str_trim(.group)) %>%
  dplyr::select(Response, Water, .group)

# Label positions
label_positions_alloc_water <- summary_alloc_water %>%
  left_join(letters_alloc_water, by = c("Response", "Water")) %>%
  group_by(Response) %>%
  mutate(
    panel_range = max(ymax, na.rm = TRUE) - min(ymin, na.rm = TRUE),
    label_y = ymax + 0.10 * panel_range
  ) %>%
  ungroup()

# Model stats: Water p-value + model R²
model_stats_alloc_water <- alloc_water_long %>%
  group_by(Response) %>%
  group_modify(~ {

    mod <- lm(Value ~ Water, data = .x)
    anova_out <- anova(mod)

    tibble(
      r2 = summary(mod)$r.squared,
      p_water = anova_out["Water", "Pr(>F)"]
    )

  }) %>%
  mutate(
    label = paste0(
      "Water: ",
      format_p(p_water),
      "\nR² = ",
      sprintf("%.2f", r2)
    )
  )

# Position model-stat labels
stats_positions_alloc_water <- summary_alloc_water %>%
  group_by(Response) %>%
  summarise(
    x = 1,
    y = max(ymax, na.rm = TRUE) +
      0.28 * (max(ymax, na.rm = TRUE) - min(ymin, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  left_join(model_stats_alloc_water, by = "Response")

# Plot
alloc_ppt_water_3panel.png <- ggplot(
  summary_alloc_water,
  aes(x = Water, y = mean, group = 1)
) +
  geom_line(
    linewidth = 1.2,
    color = "#3B6C8E"
  ) +
  geom_point(
    size = 3.8,
    color = "#3B6C8E"
  ) +
  geom_errorbar(
    aes(ymin = ymin, ymax = ymax),
    width = 0.12,
    linewidth = 0.8,
    color = "#3B6C8E"
  ) +
  geom_text(
    data = label_positions_alloc_water,
    aes(
      x = Water,
      y = label_y,
      label = .group
    ),
    inherit.aes = FALSE,
    fontface = "bold",
    size = 5,
    color = "black"
  ) +
  geom_text(
    data = stats_positions_alloc_water,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    size = 3.6,
    color = "black"
  ) +
  facet_wrap(~ Response, scales = "free_y", nrow = 1) +
  scale_y_continuous(expand = expansion(mult = c(0.12, 0.32))) +
  labs(
    title = "Drought Impact on Biomass Allocation",
    x = "Drought treatment",
    y = "Allocation (%)",
    caption = "Points show means ± SE. Letters indicate Sidak-adjusted pairwise differences among drought treatments within each panel."
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    strip.text = element_text(face = "bold", size = 13),
    strip.background = element_rect(fill = "#EAF1F5", color = NA),
    panel.grid = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.8),
    axis.ticks = element_line(color = "black"),
    panel.spacing = unit(1.5, "lines"),
    plot.caption = element_text(hjust = 0, size = 10),
    plot.margin = margin(10, 20, 15, 10)
  )

ggsave(
  "PresentationFigs/Plabels/alloc_ppt_3panel.png",
  plot = alloc_ppt_water_3panel.png,
  width = 12,
  height = 6,
  dpi = 300
)
```

## FIGURE S2 (b version with arb vesc and hyphal)

```{r}
# Helper for standard error
mean_se <- function(x) {
  m <- mean(x, na.rm = TRUE)
  se <- sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))
  data.frame(y = m, ymin = m - se, ymax = m + se)
}

# Long format for the two colonization metrics
colonization_long <- mono_only %>%
  dplyr::select(Water, hyphal_colonization, arb_vesc_colonization) %>%
  pivot_longer(
    cols = c(hyphal_colonization, arb_vesc_colonization),
    names_to = "Response",
    values_to = "Value"
  ) %>%
  mutate(
    Response = case_when(
      Response == "hyphal_colonization" ~ "(A) Hyphal Colonization",
      Response == "arb_vesc_colonization" ~ "(B) Arbuscular + Vesicular Colonization"
    )
  )

# Model-level p-values and R² for each panel
model_labels <- colonization_long %>%
  group_by(Response) %>%
  group_modify(~ {
    mod <- lm(Value ~ Water, data = .x)
    p_val <- anova(mod)[["Pr(>F)"]][1]
    r2_val <- summary(mod)$r.squared
    
    p_label <- ifelse(
      p_val < 0.001,
      "italic(p) < 0.001",
      paste0("italic(p) == ", signif(p_val, 2))
    )
    
    r2_label <- paste0("italic(R)^2 == ", round(r2_val, 2))
    
    tibble(
      x = 1,
      y = max(.x$Value, na.rm = TRUE),
      label = paste(r2_label, "*','~~", p_label)
    )
  }) %>%
  ungroup()

# Figure
root_colonization_drought.png <- ggplot(
  colonization_long,
  aes(x = Water, y = Value)
) +
  stat_summary(
    fun = mean,
    geom = "bar",
    fill = "#3B6C8E",
    color = "black",
    alpha = 0.85,
    width = 0.7
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    linewidth = 0.9,
    color = "black"
  ) +
  geom_text(
    data = model_labels,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = 0,
    vjust = 1,
    parse = TRUE,
    size = 4.5
  ) +
  facet_wrap(~ Response, nrow = 1) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.18))) +
  labs(
    title = "Drought Impact on AMF Root Colonization",
    x = "",
    y = "Root Colonization (%)"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    
    strip.text = element_text(face = "bold", size = 14),
    
    panel.grid = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.8),
    axis.ticks = element_line(color = "black"),
    
    plot.margin = margin(10, 20, 15, 10)
  )

root_colonization_drought.png

ggsave(
  "PresentationFigs/Plabels/root_colonization_drought.png",
  plot = root_colonization_drought.png,
  width = 10,
  height = 4.8,
  dpi = 300
)
```


# ONLY FOR THE SLIDESHOW

## Thank you slide -- NPP, Berries, Berries %

```{r}
se <- function(x) sd(x, na.rm = TRUE) / sqrt(sum(!is.na(x)))

dir.create("PresentationFigs/Plabels", recursive = TRUE, showWarnings = FALSE)

# Make long dataframe with 3 response variables
inoc_response_long <- mono_only %>%
  dplyr::select(
    Water, Inoculation,
    NPP_g_pp,
    Berries_g_pp,
    percent_berries_of_wheat
  ) %>%
  pivot_longer(
    cols = c(NPP_g_pp, Berries_g_pp, percent_berries_of_wheat),
    names_to = "Response",
    values_to = "Value"
  ) %>%
  mutate(
    Response = case_when(
      Response == "NPP_g_pp" ~ "Total Biomass",
      Response == "Berries_g_pp" ~ "Yield",
      Response == "percent_berries_of_wheat" ~ "Yield Allocation"
    ),
    Response = factor(Response, levels = c("Total Biomass", "Yield", "Yield Allocation")),
    Water = factor(Water, levels = c("Control", "Moderate", "Severe"))
  )

# Summary statistics
summary_inoc <- inoc_response_long %>%
  group_by(Response, Water, Inoculation) %>%
  summarise(
    mean = mean(Value, na.rm = TRUE),
    se = se(Value),
    ymin = mean - se,
    ymax = mean + se,
    .groups = "drop"
  )

# Pairwise letters: comparing Inoculation within each Water level, within each Response
letters_inoc <- inoc_response_long %>%
  group_by(Response, Water) %>%
  group_modify(~ {
    mod <- lm(Value ~ Inoculation, data = .x)
    em <- emmeans::emmeans(mod, ~ Inoculation)
    cld <- multcomp::cld(em, adjust = "sidak", Letters = letters)
    as.data.frame(cld)
  }) %>%
  mutate(.group = stringr::str_trim(.group)) %>%
  dplyr::select(Response, Water, Inoculation, .group)

# Label positions
label_positions_inoc <- summary_inoc %>%
  left_join(letters_inoc, by = c("Response", "Water", "Inoculation")) %>%
  group_by(Response) %>%
  mutate(
    panel_range = max(ymax, na.rm = TRUE) - min(ymin, na.rm = TRUE),
    label_y = ymax + 0.10 * panel_range
  ) %>%
  ungroup()

pd <- position_dodge(width = 0.3)

inoc_ppt_interact_3panel.png <- ggplot(
  summary_inoc,
  aes(x = Water, y = mean, color = Inoculation, group = Inoculation)
) +
  geom_line(linewidth = 1.2, position = pd) +
  geom_point(size = 3.8, position = pd) +
  geom_errorbar(
    aes(ymin = ymin, ymax = ymax),
    width = 0.12,
    linewidth = 0.8,
    position = pd
  ) +
  geom_text(
    data = label_positions_inoc,
    aes(x = Water, y = label_y, label = .group, color = Inoculation),
    position = pd,
    inherit.aes = FALSE,
    fontface = "bold",
    size = 5
  ) +
  facet_wrap(~ Response, scales = "free_y", nrow = 1) +
  scale_color_manual(values = c("#3B6C8E", "#C2A46D")) +
  scale_y_continuous(expand = expansion(mult = c(0.12, 0.25))) +
  labs(
    title = "Interacting Effects of Precipitation and AMF Inoculation",
    x = "Drought treatment",
    y = NULL,
    color = "Inoculation",
    caption = "Points show means ± SE. Different letters indicate adjusted pairwise differences between inoculation treatments within each drought level (p < 0.05)."
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),

    strip.text = element_text(face = "bold", size = 13),
    strip.background = element_rect(fill = "#EAF1F5", color = NA),

    panel.grid = element_blank(),
    axis.line = element_line(color = "black", linewidth = 0.8),
    axis.ticks = element_line(color = "black"),

    panel.spacing = unit(1.5, "lines"),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    plot.caption = element_text(hjust = 0, size = 10),
    plot.margin = margin(10, 20, 15, 10)
  )

ggsave(
  "PresentationFigs/Plabels/inoc_ppt_interact_3panel.png",
  plot = inoc_ppt_interact_3panel.png,
  width = 12,
  height = 5,
  dpi = 300
)


```




# FAVA

# Fava as mitgator? BUT NOT EVEN PLABELED

```{r}
# Berry allocation
fava_berry_allocation.png <- ggplot(
  combined,
  aes(
    x = Water,
    y = percent_berries_of_wheat,
    color = Inoculation,
    shape = Polyculture,
    linetype = Polyculture,
    group = interaction(Inoculation, Polyculture)
  )
) +
  stat_summary(
    fun = mean,
    geom = "line",
    linewidth = 1.1,
    position = position_dodge(width = 0.15)
  ) +
  stat_summary(
    fun = mean,
    geom = "point",
    size = 3,
    alpha = 0.9,
    position = position_dodge(width = 0.15)
  ) +
  scale_color_manual(values = c("Inoc" = "#3B6C8E", "Strl" = "#C2A46D")) +
  labs(
    title = "Fava Intercropping May Ameliorate Decline in Yield \n Allocation Incurred by Inoculation During Drought",
    x = "",
    y = "Allocation to Yield (%)",
    color = "Inoculation",
    shape = "Polyculture",
    linetype = "Polyculture"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 17),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(color = "black"),
    legend.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank()
  )

ggsave("PresentationFigs/fava_berry_allocation.png", plot = fava_berry_allocation.png, width = 9, height = 4.8, dpi = 300)
```


