---
title: "statistically_rigorous"
author: "Celeste Basken"
date: "2026-04-04"
output: html_document
---

This is Yet Another doc to look at the significance of relationships! Wow! Yay! This one is using better stats though hopefully

It's Mono Only because that's my more limited paper scope


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}

library(tidyverse)
library(broom)
library(knitr)
library(dplyr)
library(ggplot2)

# specific new ones
library(car)
library(emmeans)
library(effectsize)
library(WRS2)
```

## Load data

```{r}
# Combined data
combined <- read.csv("../Cross_compare/NPP_AMF_combined.csv")

# need to remove the fourth row bc I'm missing data (the bnpp)
combined <- combined[-4,]

combined <- combined %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Group = interaction(Inoculation, Polyculture, sep = "_")
  )

# Speciality subsetted ones
mono_only <- combined[combined$Polyculture != "Inter", ]

inter_only <- combined[combined$Polyculture != "Mono", ]

```

## Part 1: Water

### NPP (g) per plant

```{r}
# ANOVA
an_biomass2_mono_only <- aov(NPP_g_pp~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)

# Assumption checking
mod <- aov(NPP_g_pp ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  #LOOKS GOOD

plot(mod, 1)  # residuals vs fitted
leveneTest(NPP_g_pp ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```

### ANPP pp g

```{r}
# ANOVA
an_biomass2_mono_only <- aov(ANPP_g_pp~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(ANPP_g_pp ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  #LOOKS GOOD

plot(mod, 1)  # residuals vs fitted
  # it seems random but not SUPER random?? idk
leveneTest(ANPP_g_pp ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```

### BNPP pp g

```{r}
# ANOVA
an_biomass2_mono_only <- aov(BNPP_g_pp~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(BNPP_g_pp ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  #LOOKS GOOD

plot(mod, 1)  # residuals vs fitted
  # looks okay
leveneTest(BNPP_g_pp ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```


### Berries pp g

```{r}
# ANOVA
an_biomass2_mono_only <- aov(Berries_g_pp~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(Berries_g_pp ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  #I think this looks okay?

plot(mod, 1)  # residuals vs fitted
  # looks okay
leveneTest(Berries_g_pp ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```

### ANPP %

```{r}
# ANOVA
an_biomass2_mono_only <- aov(percent_ANPP_of_wheat~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(percent_ANPP_of_wheat ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  # Looks good!

plot(mod, 1)  # residuals vs fitted
  # looks good!
leveneTest(percent_ANPP_of_wheat ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```

### BNPP %

```{r}
# ANOVA
an_biomass2_mono_only <- aov(percent_BNPP_of_wheat~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(percent_BNPP_of_wheat ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  # Looks good!

plot(mod, 1)  # residuals vs fitted
  # looks okay?
leveneTest(percent_BNPP_of_wheat ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```

### Berries %

```{r}
# ANOVA
an_biomass2_mono_only <- aov(percent_berries_of_wheat~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(percent_berries_of_wheat ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  # Looks potentially problematic?? Additionally testing:

plot(mod, 1)  # residuals vs fitted
  # looks okay?
leveneTest(percent_berries_of_wheat ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```

Oh now things are getting funky with assumptions due to the nonnormal QQ
```{r}
kruskal.test(percent_berries_of_wheat ~ interaction(Inoculation, Water), data = mono_only)

# check range first
range(mono_only$percent_berries_of_wheat, na.rm = TRUE)

# if values are proportions on 0–1 scale:
mono_only$percent_berries_logit <- qlogis(
  pmin(pmax(mono_only$percent_berries_of_wheat, 0.001), 0.999)
)

mod_logit <- aov(percent_berries_logit ~ Inoculation * Water, data = mono_only)
summary(mod_logit)

plot(mod_logit, 2)
shapiro.test(residuals(mod_logit))
plot(mod_logit, 1)

library(car)
leveneTest(percent_berries_logit ~ interaction(Inoculation, Water), data = mono_only)
```

This isn't actually needed but:
```{r}
#library(ARTool)

#mono_only$Water <- factor(mono_only$Water)
#mono_only$Inoculation <- factor(mono_only$Inoculation)

#mod_art <- art(percent_berries_of_wheat ~ Inoculation * Water, data = mono_only)
#anova(mod_art)
```

### AMF colonization

```{r}
inoc_only <- mono_only[mono_only$Inoculation == "Inoc", ]

aov(hyphal_colonization ~ Water, data = inoc_only)
aov(arb_vesc_colonization ~ Water, data = inoc_only)

# GLM
inoc_only$hyphae_present <- round(inoc_only$hyphal_colonization * 100)
inoc_only$hyphae_absent  <- 100 - inoc_only$hyphae_present

glm_hyphae <- glm(cbind(hyphae_present, hyphae_absent) ~ Water,
                  family = binomial, data = inoc_only)

summary(glm_hyphae)

# lm
lm(percent_berries_of_wheat ~ hyphal_colonization, data = inoc_only)
lm(percent_ANPP_of_wheat ~ hyphal_colonization, data = inoc_only)
lm(percent_BNPP_of_wheat ~ hyphal_colonization, data = inoc_only)
lm(percent_berries_of_wheat ~ hyphal_colonization * Water, data = inoc_only)

summary(lm(percent_ANPP_of_wheat ~ hyphal_colonization, data = inoc_only))
summary(lm(percent_BNPP_of_wheat ~ hyphal_colonization, data = inoc_only))
summary(lm(percent_berries_of_wheat ~ hyphal_colonization, data = inoc_only))

```

