---
title: "statistically_rigorous"
author: "Celeste Basken"
date: "2026-04-04"
output: html_document
---

This is Yet Another doc to look at the significance of relationships! Wow! Yay! This one is using better stats though hopefully

It's Mono Only because that's my more limited paper scope


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}

library(tidyverse)
library(broom)
library(knitr)
library(dplyr)
library(ggplot2)

# specific new ones
library(car)
library(emmeans)
library(effectsize)
library(WRS2)
```

## Load data

```{r}
# Combined data
combined <- read.csv("../Cross_compare/NPP_AMF_combined.csv")

# need to remove the fourth row bc I'm missing data (the bnpp)
combined <- combined[-4,]

combined <- combined %>%
  mutate(
    Inoculation = factor(Inoculation),
    Polyculture = factor(Polyculture),
    Group = interaction(Inoculation, Polyculture, sep = "_")
  )

# percent vegetative of wheat
combined <- combined %>%
  mutate(
    percent_vegetative_of_wheat = percent_ANPP_of_wheat + percent_BNPP_of_wheat
  )

# Speciality subsetted ones
mono_only <- combined[combined$Polyculture != "Inter", ]

inter_only <- combined[combined$Polyculture != "Mono", ]
```

## Part 1: More rigorous stats

### NPP (g) per plant

```{r}
# ANOVA
an_biomass2_mono_only <- aov(NPP_g_pp~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)

# Assumption checking
mod <- aov(NPP_g_pp ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  #LOOKS GOOD

plot(mod, 1)  # residuals vs fitted
leveneTest(NPP_g_pp ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water, adjust = "tukey")

# Effect sizes
eta_squared(mod)

# r squared
mod_lm <- lm(NPP_g_pp ~ Inoculation * Water, data = mono_only)
summary(mod_lm)


# cleanly together
mod <- lm(NPP_g_pp ~ Inoculation * Water, data = mono_only)

summary(mod)        # gives R²
anova(mod)  

# okay type 2 anova??
mod <- lm(NPP_g_pp ~ Inoculation * Water, data = mono_only)

Anova(mod, type = 2)   # correct ANOVA
summary(mod)$r.squared # R²
```

### ANPP pp g

```{r}
# ANOVA
an_biomass2_mono_only <- aov(ANPP_g_pp~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(ANPP_g_pp ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  #LOOKS GOOD

plot(mod, 1)  # residuals vs fitted
  # it seems random but not SUPER random?? idk
leveneTest(ANPP_g_pp ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```

### BNPP pp g

```{r}
# ANOVA
an_biomass2_mono_only <- aov(BNPP_g_pp~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(BNPP_g_pp ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  #LOOKS GOOD

plot(mod, 1)  # residuals vs fitted
  # looks okay
leveneTest(BNPP_g_pp ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```


### Berries pp g

```{r}
# ANOVA
an_biomass2_mono_only <- aov(Berries_g_pp~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(Berries_g_pp ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  #I think this looks okay?

plot(mod, 1)  # residuals vs fitted
  # looks okay
leveneTest(Berries_g_pp ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)

# cleanly together
mod <- lm(Berries_g_pp ~ Inoculation * Water, data = mono_only)

summary(mod)        # gives R²
anova(mod)  
```

### ANPP %

```{r}
# ANOVA
an_biomass2_mono_only <- aov(percent_ANPP_of_wheat~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(percent_ANPP_of_wheat ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  # Looks good!

plot(mod, 1)  # residuals vs fitted
  # looks good!
leveneTest(percent_ANPP_of_wheat ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```

### BNPP %

```{r}
# ANOVA
an_biomass2_mono_only <- aov(percent_BNPP_of_wheat~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(percent_BNPP_of_wheat ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  # Looks good!

plot(mod, 1)  # residuals vs fitted
  # looks okay?
leveneTest(percent_BNPP_of_wheat ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)
```

### Berries %

```{r}
# ANOVA
an_biomass2_mono_only <- aov(percent_berries_of_wheat~Inoculation*Water, data= mono_only)
summary(an_biomass2_mono_only)
# THERE is a interaction inoc:water

# Assumption checking
mod <- aov(percent_berries_of_wheat ~ Inoculation * Water, data = mono_only)
plot(mod, 2)  # QQ plot normality of residuals
shapiro.test(residuals(mod))
  # Looks potentially problematic?? Additionally testing:

plot(mod, 1)  # residuals vs fitted
  # looks okay?
leveneTest(percent_berries_of_wheat ~ interaction(Inoculation, Water), data = mono_only)

# Post-hoc tests
emmeans(mod, pairwise ~ Water | Inoculation, adjust = "tukey")

# Effect sizes
eta_squared(mod)

# cleanly together
mod <- lm(percent_berries_of_wheat ~ Inoculation * Water, data = mono_only)

summary(mod)        # gives R²
anova(mod) 
```

```{r}
# allocation to vegetative

mod <- lm(percent_vegetative_of_wheat ~ Inoculation * Water, data = mono_only)

summary(mod)        # gives R²
anova(mod) 
```


Oh now things are getting funky with assumptions due to the nonnormal QQ
```{r}
kruskal.test(percent_berries_of_wheat ~ interaction(Inoculation, Water), data = mono_only)

# check range first
range(mono_only$percent_berries_of_wheat, na.rm = TRUE)

# if values are proportions on 0–1 scale:
mono_only$percent_berries_logit <- qlogis(
  pmin(pmax(mono_only$percent_berries_of_wheat, 0.001), 0.999)
)

mod_logit <- aov(percent_berries_logit ~ Inoculation * Water, data = mono_only)
summary(mod_logit)

plot(mod_logit, 2)
shapiro.test(residuals(mod_logit))
plot(mod_logit, 1)

library(car)
leveneTest(percent_berries_logit ~ interaction(Inoculation, Water), data = mono_only)
```

This isn't actually needed but:
```{r}
#library(ARTool)

#mono_only$Water <- factor(mono_only$Water)
#mono_only$Inoculation <- factor(mono_only$Inoculation)

#mod_art <- art(percent_berries_of_wheat ~ Inoculation * Water, data = mono_only)
#anova(mod_art)
```

### AMF colonization INOC ONLY (Lau is anti this)

```{r}
aov(hyphal_colonization ~ Water, data = inoc_only)
aov(arb_vesc_colonization ~ Water, data = inoc_only)

# GLM
inoc_only$hyphae_present <- round(inoc_only$hyphal_colonization * 100)
inoc_only$hyphae_absent  <- 100 - inoc_only$hyphae_present

glm_hyphae <- glm(cbind(hyphae_present, hyphae_absent) ~ Water,
                  family = binomial, data = inoc_only)

summary(glm_hyphae)

# lm
lm(percent_berries_of_wheat ~ hyphal_colonization, data = inoc_only)
lm(percent_ANPP_of_wheat ~ hyphal_colonization, data = inoc_only)
lm(percent_BNPP_of_wheat ~ hyphal_colonization, data = inoc_only)
lm(percent_berries_of_wheat ~ hyphal_colonization * Water, data = inoc_only)

summary(lm(percent_ANPP_of_wheat ~ hyphal_colonization, data = inoc_only))
summary(lm(percent_BNPP_of_wheat ~ hyphal_colonization, data = inoc_only))
summary(lm(percent_berries_of_wheat ~ hyphal_colonization, data = inoc_only))

```

# AMF hyphal but with all mono data

```{r}
aov(hyphal_colonization ~ Water, data = mono_only)
aov(arb_vesc_colonization ~ Water, data = mono_only)

# GLM
mono_only$hyphae_present <- round(mono_only$hyphal_colonization * 100)
mono_only$hyphae_absent  <- 100 - mono_only$hyphae_present

glm_hyphae <- glm(cbind(hyphae_present, hyphae_absent) ~ Water,
                  family = binomial, data = mono_only)

summary(glm_hyphae)

# lm
lm(percent_berries_of_wheat ~ hyphal_colonization, data = mono_only)
lm(percent_ANPP_of_wheat ~ hyphal_colonization, data = mono_only)
lm(percent_BNPP_of_wheat ~ hyphal_colonization, data = mono_only)
lm(percent_berries_of_wheat ~ hyphal_colonization * Water, data = mono_only)

summary(lm(percent_ANPP_of_wheat ~ hyphal_colonization, data = mono_only))
summary(lm(percent_BNPP_of_wheat ~ hyphal_colonization, data = mono_only))
summary(lm(percent_berries_of_wheat ~ hyphal_colonization, data = mono_only))

```

#NPP

```{r}
mod <- lm(NPP_g_pp ~ hyphal_colonization * Water, data = mono_only)

summary(mod)        # gives R²
anova(mod) 
```

# Berries

```{r}
mod <- lm(Berries_g_pp ~ hyphal_colonization * Water, data = mono_only)

summary(mod)        # gives R²
anova(mod) 
```

# Berries %

```{r}
mod <- lm(percent_berries_of_wheat ~ hyphal_colonization * Water, data = mono_only)

summary(mod)        # gives R²
anova(mod) 
```

Kind of out of scope but can we visualize this rq

```{r}
ggplot(mono_only, aes(x = Water, y = hyphal_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on hyphal colonization", x = "", y = "Hyphal colonization")

ggplot(mono_only, aes(x = Water, y = arb_vesc_colonization)) +
  stat_summary(fun = mean, geom = "bar") +
  labs(title = "Water impact on Arb/Vesc colonization", x = "", y = "Arb/Vesc colonization")


ggplot(mono_only, aes(x = hyphal_colonization, y = percent_berries_of_wheat)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Hyphal colonization (%)",
    y = "Berry allocation (%)",
    title = "AMF colonization & berry allocation"
  )

ggplot(mono_only, aes(x = hyphal_colonization, y = percent_ANPP_of_wheat)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Hyphal colonization (%)",
    y = "ANPP allocation (%)",
    title = "AMF colonization & ANPP allocation"
  )

ggplot(mono_only, aes(x = hyphal_colonization, y = NPP_g_pp)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Hyphal colonization (%)",
    y = "NPP per plant (g)",
    title = "AMF colonization & NPP"
  )
```


# Most interesting data for abstract

```{r}
# If mono_only is already made, start here:
berry_summary <- mono_only %>%
  group_by(Water, Inoculation) %>%
  summarise(
    mean_berry_alloc = mean(percent_berries_of_wheat, na.rm = TRUE),
    sd_berry_alloc = sd(percent_berries_of_wheat, na.rm = TRUE),
    n = n(),
    se_berry_alloc = sd_berry_alloc / sqrt(n),
    .groups = "drop"
  )

berry_summary

severe_berry_change <- berry_summary %>%
  filter(Water == "Severe") %>%
  select(Inoculation, mean_berry_alloc) %>%
  tidyr::pivot_wider(
    names_from = Inoculation,
    values_from = mean_berry_alloc
  ) %>%
  mutate(
    difference_proportion = Inoc - Strl,
    difference_percentage_points = difference_proportion * 100,
    percent_change_relative_to_sterile = ((Inoc - Strl) / Strl) * 100
  )

severe_berry_change

severe_berry_change %>%
  mutate(
    Inoc_percent = Inoc * 100,
    Strl_percent = Strl * 100
  ) %>%
  select(
    Inoc_percent,
    Strl_percent,
    difference_percentage_points,
    percent_change_relative_to_sterile
  )
```

```{r}
# Summarize overall means by inoculation treatment
overall_summary <- mono_only %>%
  group_by(Inoculation) %>%
  summarise(
    mean_NPP = mean(NPP_g_pp, na.rm = TRUE),
    mean_berries = mean(Berries_g_pp, na.rm = TRUE),
    
    sd_NPP = sd(NPP_g_pp, na.rm = TRUE),
    sd_berries = sd(Berries_g_pp, na.rm = TRUE),
    
    n = n(),
    .groups = "drop"
  )

overall_summary

overall_changes <- overall_summary %>%
  select(Inoculation, mean_NPP, mean_berries) %>%
  pivot_wider(
    names_from = Inoculation,
    values_from = c(mean_NPP, mean_berries)
  ) %>%
  mutate(
    
    # NPP change with inoculation
    NPP_difference = mean_NPP_Inoc - mean_NPP_Strl,
    NPP_percent_change = 
      ((mean_NPP_Inoc - mean_NPP_Strl) / mean_NPP_Strl) * 100,
    
    # Berry biomass change with inoculation
    berry_difference = mean_berries_Inoc - mean_berries_Strl,
    berry_percent_change =
      ((mean_berries_Inoc - mean_berries_Strl) / mean_berries_Strl) * 100
  )

overall_changes
```

# and now more summaries for discussion

```{r}
# Summarize means across water treatments
water_summary <- mono_only %>%
  group_by(Water) %>%
  summarise(
    mean_NPP = mean(NPP_g_pp, na.rm = TRUE),
    mean_ANPP = mean(ANPP_g_pp, na.rm = TRUE),
    mean_BNPP = mean(BNPP_g_pp, na.rm = TRUE),
    mean_berries = mean(Berries_g_pp, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

water_summary

drought_changes <- water_summary %>%
  filter(Water %in% c("Control", "Severe")) %>%
  pivot_wider(
    names_from = Water,
    values_from = c(
      mean_NPP,
      mean_ANPP,
      mean_BNPP,
      mean_berries
    )
  ) %>%
  mutate(

    # NPP
    NPP_difference = mean_NPP_Severe - mean_NPP_Control,
    NPP_percent_change =
      ((mean_NPP_Severe - mean_NPP_Control) /
         mean_NPP_Control) * 100,

    # ANPP
    ANPP_difference = mean_ANPP_Severe - mean_ANPP_Control,
    ANPP_percent_change =
      ((mean_ANPP_Severe - mean_ANPP_Control) /
         mean_ANPP_Control) * 100,

    # BNPP
    BNPP_difference = mean_BNPP_Severe - mean_BNPP_Control,
    BNPP_percent_change =
      ((mean_BNPP_Severe - mean_BNPP_Control) /
         mean_BNPP_Control) * 100,

    # Reproductive output
    berry_difference = mean_berries_Severe - mean_berries_Control,
    berry_percent_change =
      ((mean_berries_Severe - mean_berries_Control) /
         mean_berries_Control) * 100
  )

drought_changes

drought_changes %>%
  select(
    starts_with("mean_"),
    ends_with("difference"),
    ends_with("percent_change")
  )
```

```{r}
allocation_summary <- mono_only %>%
  group_by(Water) %>%
  summarise(
    mean_berry_alloc =
      mean(percent_berries_of_wheat, na.rm = TRUE),

    mean_ANPP_alloc =
      mean(percent_ANPP_of_wheat, na.rm = TRUE),

    mean_BNPP_alloc =
      mean(percent_BNPP_of_wheat, na.rm = TRUE),

    mean_vegetative_alloc =
      mean(percent_vegetative_of_wheat, na.rm = TRUE),

    .groups = "drop"
  )

allocation_summary

allocation_changes <- allocation_summary %>%
  filter(Water %in% c("Control", "Severe")) %>%
  pivot_wider(
    names_from = Water,
    values_from = c(
      mean_berry_alloc,
      mean_ANPP_alloc,
      mean_BNPP_alloc,
      mean_vegetative_alloc
    )
  ) %>%
  mutate(

    # Berry allocation
    berry_alloc_difference =
      mean_berry_alloc_Severe -
      mean_berry_alloc_Control,

    berry_alloc_percent_change =
      ((mean_berry_alloc_Severe -
          mean_berry_alloc_Control) /
          mean_berry_alloc_Control) * 100,

    # ANPP allocation
    ANPP_alloc_difference =
      mean_ANPP_alloc_Severe -
      mean_ANPP_alloc_Control,

    ANPP_alloc_percent_change =
      ((mean_ANPP_alloc_Severe -
          mean_ANPP_alloc_Control) /
          mean_ANPP_alloc_Control) * 100,

    # BNPP allocation
    BNPP_alloc_difference =
      mean_BNPP_alloc_Severe -
      mean_BNPP_alloc_Control,

    BNPP_alloc_percent_change =
      ((mean_BNPP_alloc_Severe -
          mean_BNPP_alloc_Control) /
          mean_BNPP_alloc_Control) * 100,

    # Vegetative allocation
    vegetative_alloc_difference =
      mean_vegetative_alloc_Severe -
      mean_vegetative_alloc_Control,

    vegetative_alloc_percent_change =
      ((mean_vegetative_alloc_Severe -
          mean_vegetative_alloc_Control) /
          mean_vegetative_alloc_Control) * 100
  )

allocation_changes

allocation_changes %>%
  select(
    ends_with("difference"),
    ends_with("percent_change")
  )
```

