---
title: "Significance grid - Hyphal colonization (continuous) - Intercropped only - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-08-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# READ ME FIRST

Same as `Significance_Grid_Hyphal_Full_DRAFT.Rmd` (see that file's READ ME for the full
explanation of what's different about a continuous predictor - ANCOVA models, `emtrends()`
post-hoc, scatter+regression thumbnails, the adapted Levene check), restricted to **Inter
(intercropped) pots that were also inoculated** (`Polyculture == "Inter" & Inoculation ==
"Inoc"`). That's only **n=15 pots** - half of the categorical Inter-only grid's 30, since
only Inoc pots have a real `hyphal_colonization` value to use as a predictor. Models here
are `response ~ hyphal_colonization * Water` (Polyculture is constant within this subset,
same as the categorical Inter-only grid).

**Power caveat, more acute here than in the Full-dataset version:** n=15 total means only
~5 pots per Water level - thin for any slope estimate, thinner still for the
`hyphal_colonization:Water` interaction row (comparing 3 slopes, each drawn from ~5
points). Read `assumption_summary` below before trusting any single cell here.

Fava columns are included (every Inter pot has a fava plant, so "does colonization level
predict fava biomass" is still a meaningful question); no AMF-family rows (colonization
can't predict itself).

```{r, message=FALSE}
source("battery_core.R")              # combined/inoc_only/inter_only/mono_only, run_battery()
source("battery_core_continuous.R")   # build_battery_spec_continuous(), run_ancova_check()
source("grid_helpers.R")              # sig_style(), apply_significance_subtitle(), build_cell_data(), render_grid()
source("grid_helpers_continuous.R")   # direction_label_continuous(), plot_cell_continuous(), term_labels_for_scope_continuous()
```

```{r}
battery_spec_hyphal_inter <- build_battery_spec_continuous("inter")
batt_hyphal_inter          <- run_battery(battery_spec_hyphal_inter, model_fn = run_ancova_check)
```

```{r}
knitr::kable(batt_hyphal_inter$assumption_summary,
             caption = "Assumption checks - Hyphal colonization (continuous), Inter only")
```

```{r}
plot_dir <- "significance_grid_plots_hyphal_inter"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. Build the cell data (one row per response variable x ANOVA term)

```{r}
cell_data <- build_cell_data(
  batt_hyphal_inter$term_summary, batt_hyphal_inter$spec_by_label, batt_hyphal_inter$results_list,
  plot_dir,
  direction_fn = direction_label_continuous,
  plot_fn      = plot_cell_continuous
)
```

# 2. Render the two grids (mass / percent)

```{r}
term_labels <- term_labels_for_scope_continuous("inter")
```

## 2.1 Mass variables (grams, per plant)

```{r}
render_grid("mass", cell_data, battery_spec_hyphal_inter, term_labels)
```

## 2.2 Percent / allocation variables

```{r}
render_grid("percent", cell_data, battery_spec_hyphal_inter, term_labels)
```
