---
title: "Significance grid - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-08-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# READ ME FIRST

This recreates the shape of your old PDF significance table (rows = ANOVA terms, columns
= response variables, one grid for raw-gram "mass" variables and one for "percent"
variables) - but generated from the `battery_core.R` models instead of hand-typed, with:

- **cell label** (`Yes` / `Kinda (<.1)` / `No` / blank-`NA`) based on the **BH-adjusted**
  p-value (`p_adj`) from `term_summary`, not a raw per-test p-value - per your call.
- **direction text** (e.g. `Inter > Mono`, or per-slice text for interactions like
  `Inter: Inoc > Strl<br>Mono: Strl > Inoc`) computed automatically from group means,
  instead of typed by hand.
- **a small thumbnail plot** embedded directly in the cell, generated from the same data
  used to fit the model - now with the p-value and significance stars (`***`/`**`/`*`/`.`/`ns`,
  the standard convention) printed as a subtitle directly on the plot itself, bolded and
  darkened when p_adj < .05 so significant panels visually pop out of the grid. This uses
  the exact same `p_adj` as the `Yes`/`Kinda`/`No` text below it, so the two can never
  disagree with each other.

A blank/`NA` cell here means the term literally isn't in that variable's model (e.g. no
"AMF inoculation" row for the AMF-colonization columns, since those are fit on the
Inoc-only subset) - this happens automatically from the model structure in
`battery_core.R`, nothing is being special-cased here to produce it. That's also why the
old table's "Yes (duh)" cells don't reappear: there's no tautological "does inoculation
predict colonization" test being run in the first place anymore.

```{r, message=FALSE}
# kableExtra isn't used elsewhere in your scripts yet - install if needed:
# install.packages("kableExtra")
library(kableExtra)

source("battery_core.R")
```

```{r}
# Where the per-cell PNG thumbnails get saved. Created fresh each knit.
plot_dir <- "significance_grid_plots"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. Helper: turn "Inoculation:Water" into a direction sentence

```{r}
direction_label <- function(data, response, term_vars) {
  # term_vars is the term string split on ":", e.g. c("Water") for a main
  # effect, or c("Inoculation","Water") for a 2-way interaction.

  if (length(term_vars) == 1) {
    # ---- Main effect: just rank the group means, high to low ----
    form  <- stats::as.formula(paste(response, "~", term_vars[1]))
    means <- stats::aggregate(form, data = data, FUN = mean, na.rm = TRUE)
    means <- means[order(-means[[response]]), ]
    paste(means[[term_vars[1]]], collapse = " > ")

  } else if (length(term_vars) == 2) {
    # ---- 2-way interaction: rank factor A's levels SEPARATELY within each
    # level of factor B, e.g. "Inter: Inoc > Strl" / "Mono: Strl > Inoc" -
    # this is exactly the by-hand format you used in the old PDF table.
    fA <- term_vars[1]
    fB <- term_vars[2]
    form  <- stats::as.formula(paste(response, "~", fA, "+", fB))
    means <- stats::aggregate(form, data = data, FUN = mean, na.rm = TRUE)

    splits <- split(means, means[[fB]])
    lines <- vapply(names(splits), function(lvl) {
      sub <- splits[[lvl]]
      sub <- sub[order(-sub[[response]]), ]
      paste0(lvl, ": ", paste(sub[[fA]], collapse = " > "))
    }, character(1))
    paste(lines, collapse = "<br>")

  } else {
    # ---- 3-way interaction: too many combinations for a one-line summary -
    # leave a pointer to the thumbnail instead of guessing at a caption.
    # (Your old table hand-wrote this row too, e.g. the "But! In mono,
    # strl > inoc..." note - that kind of nuance still needs a human once
    # you're looking at the actual faceted plot below.)
    "see facets ->"
  }
}
```

# 2. Helper: format a p-value into stars + text

```{r}
# Standard significance-star convention. Deliberately keyed off p_adj (the
# same BH-adjusted value that decides the Yes/Kinda(<.1)/No label below the
# plot) - not raw p - so the star on the figure can never disagree with the
# text label next to it.
p_to_stars <- function(p) {
  dplyr::case_when(
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    p < 0.10  ~ ".",     # conventional "marginal" marker
    TRUE      ~ "ns"
  )
}

format_p <- function(p) {
  if (p < 0.001) "p < .001" else paste0("p = ", formatC(p, digits = 3, format = "f"))
}
```

# 3. Helper: build + save one cell's thumbnail plot

```{r}
plot_cell <- function(data, response, term_vars, out_dir, cell_id, p_val) {

  # ---- Step 1: mean + standard error per group, however many factors ----
  summary_df <- data %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(term_vars))) %>%
    dplyr::summarise(
      mean_y = mean(.data[[response]], na.rm = TRUE),
      se_y   = sd(.data[[response]], na.rm = TRUE) / sqrt(sum(!is.na(.data[[response]]))),
      .groups = "drop"
    )

  # ---- Step 2: decide which factor goes on x / color / facet ----
  # Whichever factor has the MOST levels goes on the x-axis (in our design
  # that's always Water, when Water is involved, since it has 3 levels vs.
  # 2 for Inoculation/Polyculture). Second-most goes to color. Third (only
  # for the 3-way term) goes to facets.
  n_levels     <- sapply(term_vars, function(v) dplyr::n_distinct(data[[v]]))
  ordered_vars <- term_vars[order(-n_levels)]

  x_var     <- ordered_vars[1]
  color_var <- if (length(ordered_vars) >= 2) ordered_vars[2] else NA
  facet_var <- if (length(ordered_vars) == 3) ordered_vars[3] else NA

  # ---- Step 3: build the plot ----
  # IMPORTANT ORDERING NOTE: theme_minimal() is a COMPLETE theme, not an
  # incremental tweak - adding it AFTER a theme(legend.position = "bottom")
  # call (as an earlier version of this function did, and as is easy to do
  # by accident) silently wipes that setting back to theme_minimal()'s own
  # default (legend on the right), which is exactly what was cramming the
  # legend into the corner of the interaction-plot thumbnails. Fix: apply
  # theme_minimal() FIRST, immediately on creation, and only ever use plain
  # theme(...) calls (which merge into the existing theme rather than
  # replacing it) after that point.
  if (length(term_vars) == 1) {
    p <- ggplot(summary_df, aes(x = .data[[x_var]], y = mean_y)) +
      theme_minimal(base_size = 6) +
      geom_col(fill = "grey30", width = 0.6) +
      geom_errorbar(aes(ymin = mean_y - se_y, ymax = mean_y + se_y), width = 0.15)
  } else {
    p <- ggplot(summary_df, aes(x = .data[[x_var]], y = mean_y,
                                 color = .data[[color_var]], group = .data[[color_var]])) +
      theme_minimal(base_size = 6) +
      geom_line(linewidth = 0.5) +
      geom_point(size = 1) +
      geom_errorbar(aes(ymin = mean_y - se_y, ymax = mean_y + se_y), width = 0.15) +
      theme(legend.position = "bottom",
            legend.title    = element_blank(),
            legend.text     = element_text(size = 5),
            legend.key.size = unit(3, "mm"),
            legend.margin   = margin(t = -4))

    if (!is.na(facet_var)) {
      p <- p + facet_wrap(stats::as.formula(paste("~", facet_var)))
    }
  }

  # ---- Step 4: p-value + stars as a subtitle directly on the plot ----
  # This is the new part: rather than only stating "Yes"/"Kinda"/"No" in the
  # table cell text below the image, the actual number and stars are baked
  # into the figure itself - so if this thumbnail is ever pulled out and
  # used on its own (e.g. dropped into the thesis as a standalone figure),
  # it's still self-labeled and doesn't depend on the surrounding table.
  subtitle_txt <- paste0(format_p(p_val), "  ", p_to_stars(p_val))

  # `expand_limits(y = 0)` is the actual fix for "0 should always be at the
  # bottom": geom_col() already anchors bars at 0 by default, but the
  # interaction plots (geom_line/geom_point) don't - their y-axis just
  # ranges over whatever the group means happen to span, which does NOT
  # necessarily reach down to 0. This forces every panel's axis to include
  # 0 as its floor, bar chart or line chart alike, so the baseline is
  # consistent across the whole grid. (It only ever EXTENDS the range to
  # reach 0 - it can't clip data - so nothing gets cut off.)
  #
  # NOTE: this is a real fix on top of the plot_response change in
  # battery_core.R (drawing the natural 0-1 proportion scale instead of the
  # logit scale used for the model) - that change is what stopped the
  # percent-family panels from going negative in the first place, which was
  # the actual root cause of "0 at the top" for some of them. This
  # expand_limits() call is a belt-and-suspenders guarantee on top of that.
  p <- p +
    ggplot2::expand_limits(y = 0) +
    labs(subtitle = subtitle_txt) +
    theme(
      axis.title      = element_blank(),
      plot.margin     = margin(1, 1, 1, 1),
      # Bold + darker when the BH-adjusted result is significant (p < .05),
      # so a significant panel visually pops out of the grid even before
      # you read the number - grey/plain otherwise so "ns" doesn't compete
      # for attention.
      plot.subtitle   = element_text(
        size   = 5.5,
        face   = ifelse(p_val < 0.05, "bold", "plain"),
        colour = ifelse(p_val < 0.05, "black", "grey40"),
        hjust  = 0.5,
        margin = margin(b = 1)
      )
    )

  # ---- Step 5: save + return the file path ----
  # Height bumped up slightly from the first version (1.0 -> 1.2 in) to
  # leave room for the new subtitle line without it overlapping the plot.
  out_path <- file.path(out_dir, paste0(cell_id, ".png"))
  ggplot2::ggsave(out_path, p, width = 1.3, height = 1.2, dpi = 150)
  out_path
}
```

# 4. Build one row per (response variable, ANOVA term) cell

```{r}
# term_summary (from battery_core.R) already has one row per label x term,
# with the BH-adjusted p_adj we agreed to key the labels off. We add the
# significance code, direction text, and thumbnail path here.
cell_data <- term_summary %>%
  dplyr::mutate(
    term_vars = strsplit(term, ":"),
    sig = dplyr::case_when(
      p_adj < 0.05 ~ "Yes",
      p_adj < 0.10 ~ "Kinda (&lt;.1)",
      TRUE         ~ "No"
    )
  )

# purrr::pmap_chr walks label + term_vars together, row by row, looking up
# that label's actual data/response from spec_by_label (built in
# battery_core.R) to compute the direction text and render the plot.
cell_data$direction <- purrr::map2_chr(
  cell_data$label, cell_data$term_vars,
  function(lbl, tv) {
    spec <- spec_by_label[[lbl]]
    direction_label(spec$data, spec$response, tv)
  }
)

# p_adj is now passed straight into plot_cell() too, so the number/stars
# baked into the image are guaranteed to be the same p_adj driving `sig`
# above - never a second, silently-different p-value.
#
# A figure is generated for EVERY cell, "No" included - per your call, you'd
# rather always have the picture available to glance at than save the
# render time. (An earlier draft skipped "No" cells here; reverted.)
#
# spec$plot_response is used instead of spec$response when it's set - that's
# the natural 0-1 scale for the logit-transformed percent/AMF variables (see
# battery_core.R comment on plot_response for why: plotting the raw logit
# values put "0" at the top of the panel for some variables and the bottom
# for others, entirely depending on which side of 0.5 that variable's mean
# happened to fall on).
cell_data$plot_path <- purrr::pmap_chr(
  list(cell_data$label, cell_data$term, cell_data$term_vars, cell_data$p_adj),
  function(lbl, trm, tv, p_val) {
    spec           <- spec_by_label[[lbl]]
    plot_response  <- if (is.null(spec$plot_response)) spec$response else spec$plot_response
    cell_id        <- paste(gsub("[^A-Za-z0-9]", "_", lbl), gsub(":", "_", trm), sep = "__")
    plot_cell(spec$data, plot_response, tv, out_dir = plot_dir, cell_id = cell_id, p_val = p_val)
  }
)

# ---- Styling per significance level ----
# Yes/Kinda get a tinted cell background + colored bold label so they pop
# out of the grid at a glance; No stays plain/muted so it visually recedes
# instead of competing for attention. This is a plain if/else (not
# dplyr::case_when) because it runs once per row inside the pmap below, on
# a single scalar `sig` value each time - case_when is for vectorized use.
sig_style <- function(sig) {
  if (sig == "Yes") {
    list(bg = "#e9f7ec", fg = "#1a7431", weight = "bold")     # green tint
  } else if (sig == "Kinda (&lt;.1)") {
    list(bg = "#fff8e1", fg = "#8a6100", weight = "bold")     # amber tint
  } else {
    list(bg = "transparent", fg = "#999999", weight = "normal")  # No: muted, no tint
  }
}

# ---- Assemble the actual HTML that will go inside each table cell ----
# knitr::image_uri() base64-encodes the PNG straight into the <img> tag, so
# the knitted HTML file is self-contained (no separate image files it
# depends on to display correctly if you move/share the .html).
cell_data$cell_html <- purrr::pmap_chr(
  list(cell_data$sig, cell_data$direction, cell_data$plot_path),
  function(sig, dir, path) {
    sty <- sig_style(sig)

    # Every cell gets its image + direction text now, "No" included -
    # only the background tint/label color (sig_style) distinguishes them.
    img_html <- paste0('<img src="', knitr::image_uri(path), '" width="70"><br>')
    dir_html <- paste0("<br><span style='font-size:85%;color:#555'>", dir, "</span>")

    paste0(
      "<div style='background-color:", sty$bg, "; padding:4px; border-radius:4px;'>",
      img_html,
      "<b style='color:", sty$fg, "; font-weight:", sty$weight, ";'>", sig, "</b>",
      dir_html,
      "</div>"
    )
  }
)
```

# 5. Render the two grids (mass / percent), matching the old table's layout

```{r}
# Row order + friendlier row labels than the raw "Inoculation:Water" term
# strings - matches the order your old PDF table used.
term_labels <- c(
  "Polyculture"                   = "Polyculture",
  "Inoculation"                   = "AMF inoculation",
  "Water"                         = "Water (PPT)",
  "Inoculation:Polyculture"       = "Poly:Inoc",
  "Polyculture:Water"             = "Poly:Water",
  "Inoculation:Water"             = "Inoc:Water",
  "Inoculation:Polyculture:Water" = "Poly:Inoc:Water (all 3)"
)
term_order <- names(term_labels)

render_grid <- function(group) {
  # Which labels (response variables) belong in this grid, and in what
  # column order - taken straight from battery_spec so this file never
  # has to repeat that list itself.
  labels_in_group <- battery_spec %>%
    purrr::keep(~ .x$display_group == group) %>%
    purrr::map_chr("label")

  grid_df <- cell_data %>%
    dplyr::filter(label %in% labels_in_group) %>%
    dplyr::mutate(term = factor(term, levels = term_order)) %>%
    dplyr::select(term, label, cell_html) %>%
    tidyr::pivot_wider(names_from = label, values_from = cell_html) %>%
    dplyr::arrange(term) %>%
    # Blank out any term row this response family never fits (NA from the
    # pivot = "this term isn't in that variable's model") with a muted
    # "NA" tag instead of a literal R "NA" string.
    dplyr::mutate(dplyr::across(
      -term,
      ~ ifelse(is.na(.x), "<span style='color:#aaa'><i>NA</i></span>", .x)
    )) %>%
    dplyr::mutate(term = term_labels[as.character(term)]) %>%
    dplyr::select(term, dplyr::all_of(labels_in_group))

  grid_df %>%
    knitr::kable(format = "html", escape = FALSE, col.names = c("", labels_in_group)) %>%
    kableExtra::kable_styling(full_width = FALSE, font_size = 11) %>%
    kableExtra::column_spec(1, bold = TRUE, background = "#f0f0f0")
}
```

## 5.1 Mass variables (grams, per plant)

```{r}
render_grid("mass")
```

## 5.2 Percent / allocation variables

```{r}
render_grid("percent")
```

# 6. What this doesn't do (yet)

This only rebuilds grids 1+2 from your PDF (the significance codes + figures-with-
direction). It does **not** rebuild grid 3 (the literature-comparison table with the
yellow "significant in my data" highlighting) - that one is much more a matter of you
reading papers and writing summaries than something to generate from the model output,
though the yellow highlighting itself could be driven off `cell_data$sig == "Yes"` if you
want that layered on top of a literature table later.
