---
title: "Significance grid - Hyphal colonization (continuous) - Full dataset - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-08-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# READ ME FIRST

Same grid shape as `Significance_Grid_DRAFT.Rmd`, but with the binary `Inoculation`
(Inoc/Strl) factor replaced by the **continuous** `hyphal_colonization` measurement -
i.e. instead of "did you inoculate this pot," the predictor is "how much AMF colonization
actually resulted." This is a different statistical question, not just a different
column, so several things about the table are genuinely different:

- **Only Inoc pots are used** (n=30 of the full 60), same reasoning already applied to the
  AMF-colonization rows in the categorical grid: `hyphal_colonization` is ~0 for every
  Strl pot by construction (nothing was inoculated), not a real measured value on the same
  continuous scale - including Strl pots would just be re-detecting the Inoc/Strl split in
  disguise, with noise added.
- **Models are ANCOVA-style**, `response ~ hyphal_colonization * Polyculture * Water`, not
  the categorical 3-way factorial. Type II sums-of-squares (`car::Anova`) work the same way
  for a mix of continuous and categorical terms as for all-categorical ones.
- **No "AMF family" rows.** `hyphal_colonization`/`arb_vesc_colonization` were response
  variables in the categorical grid; here `hyphal_colonization` IS the predictor, so it
  can't also be an outcome being predicted by itself.
- **Post-hoc is `emtrends()`, not `emmeans()` pairwise comparisons.** There's no "group
  mean" to compare for a continuous predictor - the equivalent question is "does the
  SLOPE of hyphal_colonization on the response differ across Water/Polyculture levels,"
  answered by comparing estimated slopes (trends) instead of estimated means.
- **Direction text describes a slope**, not a ranking - e.g. `↑ hyphal colonization →
  ↑ response (slope = 2.13)` for a main effect, or one slope per group for an interaction
  (e.g. `Control: slope +1.8<br>Severe: slope -0.4`).
- **Thumbnails are scatter + linear-regression-line plots**, not bar/line-of-means charts -
  color/facet still follow the same "more levels → facet, fewer → color" convention as the
  categorical grid's interaction plots.
- **Levene's test only checks variance homogeneity across the categorical predictors**
  (Water, Polyculture) - it can't group observations by a continuous variable, so it does
  NOT check for a "fan-shaped" residual pattern across the range of hyphal_colonization
  itself. Flagged rather than silently treated as equivalent to the categorical grid's
  Levene check.

**A real power caveat, not just a formality:** restricting to Inoc-only pots means every
model here starts from n=30 - already half of what the categorical "full" grid uses -
before even splitting into the 2×3 = 6 cells of Polyculture×Water (~5 pots/cell). Treat
the 3-way interaction row's per-facet slopes here as exploratory. `assumption_summary`
below shows the actual `n` per model so this isn't hidden.

```{r, message=FALSE}
source("battery_core.R")              # combined/inoc_only/inter_only/mono_only, run_battery()
source("battery_core_continuous.R")   # build_battery_spec_continuous(), run_ancova_check()
source("grid_helpers.R")              # sig_style(), apply_significance_subtitle(), build_cell_data(), render_grid()
source("grid_helpers_continuous.R")   # direction_label_continuous(), plot_cell_continuous(), term_labels_for_scope_continuous()
```

```{r}
battery_spec_hyphal_full <- build_battery_spec_continuous("full")
batt_hyphal_full          <- run_battery(battery_spec_hyphal_full, model_fn = run_ancova_check)
```

```{r}
# Assumption checks first, before the grid - small-n models are the main
# thing worth actually reading here (see the power caveat above).
knitr::kable(batt_hyphal_full$assumption_summary,
             caption = "Assumption checks - Hyphal colonization (continuous), Full dataset")
```

```{r}
# Own plot directory, separate from every other grid doc's.
plot_dir <- "significance_grid_plots_hyphal_full"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. Build the cell data (one row per response variable x ANOVA term)

```{r}
cell_data <- build_cell_data(
  batt_hyphal_full$term_summary, batt_hyphal_full$spec_by_label, batt_hyphal_full$results_list,
  plot_dir,
  direction_fn = direction_label_continuous,
  plot_fn      = plot_cell_continuous
)
```

# 2. Render the two grids (mass / percent)

```{r}
term_labels <- term_labels_for_scope_continuous("full")
```

## 2.1 Mass variables (grams, per plant)

```{r}
render_grid("mass", cell_data, battery_spec_hyphal_full, term_labels)
```

## 2.2 Percent / allocation variables

```{r}
render_grid("percent", cell_data, battery_spec_hyphal_full, term_labels)
```
