---
title: "Significance grid - Full dataset - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-08-05"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# READ ME FIRST

This recreates the shape of your old PDF significance table (rows = ANOVA terms, columns
= response variables, one grid for raw-gram "mass" variables and one for "percent"
variables) for the **full dataset** (Inter + Mono pots together, Polyculture as a 3-way
factor) - generated from the `battery_core.R` models instead of hand-typed, with:

- **cell label** (`Yes` / `Kinda (<.1)` / `No` / blank-`NA`) based on the **BH-adjusted**
  p-value (`p_adj`) from `term_summary`, not a raw per-test p-value - per your call.
- **direction text** (e.g. `Inter > Mono`, or per-slice text for interactions like
  `Inter: Inoc > Strl<br>Mono: Strl > Inoc`) computed automatically from group means,
  instead of typed by hand.
- **a small thumbnail plot** embedded directly in every cell (`No` included), generated
  from the same data used to fit the model, with the p-value and significance stars
  (`***`/`**`/`*`/`.`/`ns`) printed as a subtitle directly on the plot - bolded and
  darkened when p_adj < .05 so significant panels visually pop out of the grid. This uses
  the exact same `p_adj` as the `Yes`/`Kinda`/`No` text below it, so the two can never
  disagree with each other.

A blank/`NA` cell here means the term literally isn't in that variable's model (e.g. no
"AMF inoculation" row for the AMF-colonization columns, since those are fit on the
Inoc-only subset) - this happens automatically from the model structure in
`battery_core.R`, nothing is being special-cased here to produce it. That's also why the
old table's "Yes (duh)" cells don't reappear: there's no tautological "does inoculation
predict colonization" test being run in the first place anymore.

This is one of **three** near-identical grid docs - see also
[Significance_Grid_Inter_DRAFT.Rmd] (Inter pots only) and
[Significance_Grid_Mono_DRAFT.Rmd] (Mono pots only). All three source the same
`battery_core.R` (model-fitting) and `grid_helpers.R` (plot/table-rendering) files, so
there's one place to fix anything rather than three.

```{r, message=FALSE}
source("battery_core.R")   # combined/inoc_only/inter_only/mono_only, battery_spec,
                             # spec_by_label, results_list, term_summary (scope = "full")
source("grid_helpers.R")    # plot_cell(), direction_label(), build_cell_data(), render_grid()
```

```{r}
# Where the per-cell PNG thumbnails get saved. Created fresh each knit. Each
# of the three grid docs uses its own directory (this one has no suffix,
# the Inter/Mono docs use "_inter"/"_mono") so their same-named cells (e.g.
# "Wheat_total_NPP_per_plant__Water.png" occurs in all three) don't collide
# or overwrite each other on disk.
plot_dir <- "significance_grid_plots"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. Build the cell data (one row per response variable x ANOVA term)

```{r}
cell_data <- build_cell_data(term_summary, spec_by_label, results_list, plot_dir)
```

# 2. Render the two grids (mass / percent), matching the old table's layout

```{r}
term_labels <- term_labels_for_scope("full")
```

## 2.1 Mass variables (grams, per plant)

```{r}
render_grid("mass", cell_data, battery_spec, term_labels)
```

## 2.2 Percent / allocation variables

```{r}
render_grid("percent", cell_data, battery_spec, term_labels)
```

# 3. What this doesn't do (yet)

This only rebuilds grids 1+2 from your PDF (the significance codes + figures-with-
direction). It does **not** rebuild grid 3 (the literature-comparison table with the
yellow "significant in my data" highlighting) - that one is much more a matter of you
reading papers and writing summaries than something to generate from the model output,
though the yellow highlighting itself could be driven off `cell_data$sig == "Yes"` if you
want that layered on top of a literature table later.
