---
title: "AMF root colonization x intercropping (Polyculture) - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-08-17"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(scales)   # percent-format axis labels below
```

# READ ME FIRST

This is a focused look at one question buried inside the AMF rows of
`../Significance_Grid_DRAFT.Rmd`: **does intercropping (Polyculture: Inter vs. Mono)
affect AMF root colonization** - `hyphal_colonization` and `arb_vesc_colonization`?

It doesn't fit any new models - it reuses the exact same `Polyculture * Water` models
already fit in the shared `../battery_core.R`'s "amf" family (on `inoc_only`, the
Inoc-treatment subset - colonization is a structural zero for Strl pots, so Strl pots
aren't part of this question in the first place), and just pulls out + re-presents the
`Polyculture` piece of that with dedicated plots, instead of it being two rows spread
across a 9-column grid.

Lives in `PerPlant/AMF/` (not `PerPot/`) alongside the other AMF/hyphal-colonization
analyses - `hyphal_colonization`/`arb_vesc_colonization` have no "per plant" (`_pp`)
counterpart in the first place (see `Data/Data_key_column_definitions.csv`), so these
results are identical whether they're run against the per-plant or per-pot core; this
file sources `../battery_core.R`, the per-plant one, purely because that's the core
this folder's other docs already share.

```{r, message=FALSE}
source("../battery_core.R")   # combined/inoc_only, battery_spec, spec_by_label,
                             # results_list, term_summary (scope = "full")
source("../grid_helpers.R")    # p_to_stars(), format_p(), direction_label(), build_cell_data()
```

```{r}
plot_dir <- "amf_intercropping_plots"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. The two models, side by side

Both variables are logit-transformed for the model (`hyphal_colonization_logit`,
`arb_vesc_colonization_logit`) but plotted here on their natural 0-100% scale. P-values
are **BH-adjusted within the "amf" family** (these two variables' three terms each, six
p-values total) - the same correction used in the grid doc, so these numbers match it
exactly.

```{r}
amf_terms    <- term_summary %>% dplyr::filter(family == "amf")
cell_data_amf <- build_cell_data(amf_terms, spec_by_label, results_list, plot_dir)
```

```{r}
term_order <- c("Polyculture", "Water", "Polyculture:Water")
term_labels_amf <- c(
  "Polyculture"       = "Intercropping (Poly)",
  "Water"             = "Water (PPT)",
  "Polyculture:Water" = "Poly:Water"
)

grid_df <- cell_data_amf %>%
  dplyr::mutate(term = factor(term, levels = term_order)) %>%
  dplyr::select(term, label, cell_html) %>%
  tidyr::pivot_wider(names_from = label, values_from = cell_html) %>%
  dplyr::arrange(term) %>%
  dplyr::mutate(term = term_labels_amf[as.character(term)])

grid_df %>%
  knitr::kable(format = "html", escape = FALSE, col.names = c("", names(grid_df)[-1])) %>%
  kableExtra::kable_styling(full_width = FALSE, font_size = 11) %>%
  kableExtra::column_spec(1, bold = TRUE, background = "#f0f0f0")
```

# 2. The intercropping (Polyculture) main effect, specifically

```{r}
poly_rows <- amf_terms %>% dplyr::filter(term == "Polyculture")

hyphal_row   <- poly_rows %>% dplyr::filter(label == "Hyphal colonization (%)")
arbvesc_row  <- poly_rows %>% dplyr::filter(label == "Arbuscular/vesicular colonization (%)")

means_tbl <- inoc_only %>%
  dplyr::group_by(Polyculture) %>%
  dplyr::summarise(
    hyphal   = mean(hyphal_colonization,   na.rm = TRUE),
    arb_vesc = mean(arb_vesc_colonization, na.rm = TRUE),
    .groups  = "drop"
  )
```

- **Hyphal colonization**: F(`r hyphal_row$df`) = `r round(hyphal_row$statistic, 2)`,
  `r format_p(hyphal_row$p.value)` raw / `r format_p(hyphal_row$p_adj)` BH-adjusted
  (`r p_to_stars(hyphal_row$p_adj)`), partial eta2 = `r round(hyphal_row$eta2_partial, 2)`.
  Mono mean = `r scales::percent(means_tbl$hyphal[means_tbl$Polyculture == "Mono"], accuracy = 0.1)`,
  Inter mean = `r scales::percent(means_tbl$hyphal[means_tbl$Polyculture == "Inter"], accuracy = 0.1)`.

- **Arbuscular/vesicular colonization**: F(`r arbvesc_row$df`) = `r round(arbvesc_row$statistic, 2)`,
  `r format_p(arbvesc_row$p.value)` raw / `r format_p(arbvesc_row$p_adj)` BH-adjusted
  (`r p_to_stars(arbvesc_row$p_adj)`), partial eta2 = `r round(arbvesc_row$eta2_partial, 2)`.
  Mono mean = `r scales::percent(means_tbl$arb_vesc[means_tbl$Polyculture == "Mono"], accuracy = 0.1)`,
  Inter mean = `r scales::percent(means_tbl$arb_vesc[means_tbl$Polyculture == "Inter"], accuracy = 0.1)`.

Both point the same direction (Mono > Inter) and both land in "Kinda (<.1)" territory
after correction, not "Yes" - worth flagging, not worth calling a confirmed effect yet.

# 3. Raw data: colonization by Polyculture (Water pooled)

```{r}
amf_long <- inoc_only %>%
  dplyr::select(Pot, Polyculture, Water, hyphal_colonization, arb_vesc_colonization) %>%
  tidyr::pivot_longer(
    cols      = c(hyphal_colonization, arb_vesc_colonization),
    names_to  = "colonization_type",
    values_to = "pct"
  ) %>%
  dplyr::mutate(colonization_type = dplyr::recode(colonization_type,
    hyphal_colonization   = "Hyphal (%)",
    arb_vesc_colonization = "Arbuscular/vesicular (%)"
  ))

ggplot(amf_long, aes(x = Polyculture, y = pct)) +
  geom_boxplot(outlier.shape = NA, width = 0.5, fill = "grey90") +
  geom_jitter(width = 0.08, height = 0, alpha = 0.6, size = 1.8) +
  facet_wrap(~ colonization_type) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(
    title = "AMF root colonization by intercropping treatment",
    x     = "Polyculture",
    y     = "Root colonization (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(strip.text = element_text(face = "bold"))
```

# 4. Does Water change the intercropping effect? (Polyculture x Water)

The grid's "Kinda"/"No" for `Polyculture:Water` (Section 1) says: only weakly for hyphal,
not really for arbuscular/vesicular. Plotted out:

```{r}
amf_summary <- amf_long %>%
  dplyr::group_by(colonization_type, Polyculture, Water) %>%
  dplyr::summarise(
    mean_pct = mean(pct, na.rm = TRUE),
    se_pct   = sd(pct, na.rm = TRUE) / sqrt(dplyr::n()),
    .groups  = "drop"
  )

ggplot(amf_summary, aes(x = Water, y = mean_pct, color = Polyculture, group = Polyculture)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2.5) +
  geom_errorbar(aes(ymin = mean_pct - se_pct, ymax = mean_pct + se_pct), width = 0.1) +
  facet_wrap(~ colonization_type) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(
    title = "Polyculture x Water interaction",
    x     = "Water (PPT)",
    y     = "Root colonization (%)",
    color = "Polyculture"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", strip.text = element_text(face = "bold"))
```

# 5. What this doesn't do (yet)

This only isolates + re-plots the AMF-family Polyculture results that
`Significance_Grid_DRAFT.Rmd` already computes - it doesn't add pairwise post-hoc tests
(not needed here: Polyculture only has 2 levels, so the ANOVA F-test above already IS the
Inter-vs-Mono comparison) and it doesn't touch the Inter-only/Mono-only scoped grids,
since Polyculture is constant within either of those and this question doesn't apply
there.
