---
title: "Significance grid - Full dataset, per-pot variables - DRAFT (for review)"
author: "CLAUDE + Celeste Basken"
date: "2026-08-17"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# READ ME FIRST

This is the **per-pot** counterpart of `../PerPlant/Significance_Grid_DRAFT.Rmd`: same grid
layout (rows = ANOVA terms, columns = response variables, one grid for raw-gram "mass"
variables and one for "percent" variables), same **full dataset** scope (Inter + Mono
pots together, Polyculture as a 3-way factor), same everything - except the mass-family
variables are fit on the **pot-total** biomass columns (`W_NPP`, `W_ANPP_g`, `W_BNPP_g`,
`W_Berries_g`, `Pot_NPP_g`) instead of the density-adjusted, "per plant" (`_pp`) columns
(`W_NPP_pp`, `W_ANPP_g_pp`, `W_BNPP_g_pp`, `W_Berries_g_pp`, `Pot_NPP_g_pp`) that the
original doc uses. Percent/allocation and AMF-colonization variables have no `_pp`
counterpart in the first place (see `Data/Data_key_column_definitions.csv`), so those
columns and results are unchanged from the original.

As with the original:

- **cell label** (`Yes` / `Kinda (<.1)` / `No` / blank-`NA`) based on the **BH-adjusted**
  p-value (`p_adj`) from `term_summary`, not a raw per-test p-value.
- **direction text** (e.g. `Inter > Mono`, or per-slice text for interactions like
  `Inter: Inoc > Strl<br>Mono: Strl > Inoc`) computed automatically from group means.
- **a small thumbnail plot** embedded directly in every cell (`No` included), generated
  from the same data used to fit the model, with the p-value and significance stars
  (`***`/`**`/`*`/`.`/`ns`) printed as a subtitle directly on the plot - bolded and
  darkened when p_adj < .05.

A blank/`NA` cell here means the term literally isn't in that variable's model, same as
the original - this happens automatically from the model structure in `battery_core.R`,
nothing is being special-cased here to produce it.

This doc sources this folder's own `battery_core.R` and `grid_helpers.R` (not the ones in
`CleanCode/`) - see the header comments in each for exactly what differs (only
`battery_core.R`'s mass-family response columns; `grid_helpers.R` is an unchanged copy).

```{r, message=FALSE}
source("battery_core.R")   # combined/inoc_only/inter_only/mono_only, battery_spec,
                             # spec_by_label, results_list, term_summary (scope = "full",
                             # per-pot mass variables)
source("grid_helpers.R")    # plot_cell(), direction_label(), build_cell_data(), render_grid()
```

```{r}
# Where the per-cell PNG thumbnails get saved. Created fresh each knit. This
# folder's own directory, separate from CleanCode/PerPlant/significance_grid_plots/,
# so same-named cells (e.g. this doc's "Total_pot_productivity__Water.png")
# never collide with the per-plant version's files.
plot_dir <- "significance_grid_plots"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
```

# 1. Build the cell data (one row per response variable x ANOVA term)

```{r}
cell_data <- build_cell_data(term_summary, spec_by_label, results_list, plot_dir)
```

# 2. Render the two grids (mass / percent), matching the old table's layout

```{r}
term_labels <- term_labels_for_scope("full")
```

## 2.1 Mass variables (grams, per pot)

```{r}
render_grid("mass", cell_data, battery_spec, term_labels)
```

## 2.2 Percent / allocation variables

```{r}
render_grid("percent", cell_data, battery_spec, term_labels)
```

